---
title: "Filtering for blocks with TSCA facilities"
author: "Paige Varner"
date: "2025-09-04"
output: html_document
---


#load libraries
```{r}
library("dplyr")
library("tidyr")
library("bigrquery")
library("DBI")
library("tidycensus")
library("sf")
library("purrr")
library("tigris")
library("progress")
options(tigris_use_cache = TRUE)

naics = read.csv("~/TSCA/CRA/metdb_tsca/data/naics_tri_filter.csv")
naics_codes <- naics$Code %>%
  as.character()

valid_states <- c(state.abb, "DC")
```


#database connection
```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "airtoxscreen",)

bq_auth(email = NA)
```


### BTEXSN blocks ###

#pull df of TRI facility IDs, their lat/longs, and NAICS code for facilities releasing BTEXSN, then filter for NAICS codes of interest
```{r}
# ------------------------------------------------------------
# 1. Pull TRI facilities releasing BTEXSN chemicals
# ------------------------------------------------------------

SQL_BTEXSN <- "
SELECT DISTINCT
  trifd,
  latitude,
  longitude,
  primary_NAICS,
  NAICS2,
  NAICS3,
  NAICS4,
  NAICS5,
  NAICS6
FROM `edf-aq-data.tri.tri_basic_2017-2022`
WHERE chemical IN (
  'Benzene',
  'Toluene',
  'Ethylbenzene',
  'm-Xylene',
  'p-Xylene',
  'o-Xylene',
  'Xylene (mixed isomers)',
  'Styrene',
  'Naphthalene'
)
"

TRI_BTEXSN <- dbGetQuery(
  con,
  SQL_BTEXSN
)


# ------------------------------------------------------------
# 2. Filter TRI facilities for NAICS codes of interest
# ------------------------------------------------------------

TRI_BTEXSN <- TRI_BTEXSN %>%
  filter(
    primary_NAICS %in% naics_codes |
      NAICS2 %in% naics_codes |
      NAICS3 %in% naics_codes |
      NAICS4 %in% naics_codes |
      NAICS5 %in% naics_codes |
      NAICS6 %in% naics_codes
  )


# Check number of unique filtered TRI facilities
cat(
  "BTEXSN unique TRI facilities:",
  n_distinct(TRI_BTEXSN$trifd),
  "\n"
)
```


#get list of blocks overlapping within 10km of these facilities
```{r}
# ------------------------------------------------------------
# 3. Convert BTEXSN TRI facilities to sf
# ------------------------------------------------------------

tri_sf_BTEXSN <- TRI_BTEXSN %>%
  filter(
    !is.na(latitude),
    !is.na(longitude)
  ) %>%
  st_as_sf(
    coords = c("longitude", "latitude"),
    crs = 4326
  )


# Project to CRS in meters
tri_sf_proj_BTEXSN <- tri_sf_BTEXSN %>%
  st_transform(5070)


# ------------------------------------------------------------
# 4. Create UNION of 10-km BTEXSN facility buffers
# ------------------------------------------------------------

tri_buffer_union_BTEXSN <- tri_sf_proj_BTEXSN %>%
  st_buffer(dist = 10000) %>%
  st_union()


# ------------------------------------------------------------
# 5. Create BTEXSN-specific intermediate block directory
# ------------------------------------------------------------

BTEXSN_block_dir <- 
  "~/TSCA/CRA/metdb_tsca/data/filtblocks_states_BTEXSN"

dir.create(
  BTEXSN_block_dir,
  recursive = TRUE,
  showWarnings = FALSE
)


# ------------------------------------------------------------
# 6. Find census blocks within 10 km of BTEXSN facilities
# ------------------------------------------------------------

pb_BTEXSN <- progress_bar$new(
  format = "BTEXSN :state [:bar] :percent | Found :n blocks",
  total = length(valid_states),
  clear = FALSE,
  width = 80
)


blocks_overlap_BTEXSN <- map_dfr(
  valid_states,
  function(st) {
    
    outfile <- file.path(
      BTEXSN_block_dir,
      paste0("blocks_", st, ".csv")
    )
    
    
    # Use previously calculated BTEXSN result if it exists
    if (file.exists(outfile)) {
      
      existing <- read.csv(
        outfile,
        colClasses = "character"
      )
      
      pb_BTEXSN$tick(
        tokens = list(
          state = st,
          n = paste0(nrow(existing), " cached")
        )
      )
      
      return(existing)
    }
    
    
    # Load 2020 census blocks for state
    blk <- blocks(
      state = st,
      year = 2020
    ) %>%
      st_transform(5070)
    
    
    # Find blocks intersecting any BTEXSN 10-km buffer
    blk_overlap <- blk %>%
      filter(
        st_intersects(
          .,
          tri_buffer_union_BTEXSN,
          sparse = FALSE
        )
      )
    
    
    n_found <- nrow(blk_overlap)
    
    
    pb_BTEXSN$tick(
      tokens = list(
        state = st,
        n = n_found
      )
    )
    
    
    # Keep only unique block GEOIDs
    df_out <- blk_overlap %>%
      st_drop_geometry() %>%
      select(GEOID20) %>%
      distinct() %>%
      rename(
        GEOID = GEOID20
      )
    
    
    # Save BTEXSN-specific state result
    write.csv(
      df_out,
      outfile,
      row.names = FALSE
    )
    
    
    df_out
  }
)


# ------------------------------------------------------------
# 7. Combine and deduplicate BTEXSN blocks
# ------------------------------------------------------------

filtblocks_BTEXSN <- blocks_overlap_BTEXSN %>%
  distinct(GEOID)


# Save final BTEXSN block list
write.csv(
  filtblocks_BTEXSN,
  "~/TSCA/CRA/metdb_tsca/data/filtblocks_BTEXSN.csv",
  row.names = FALSE
)


cat(
  "BTEXSN total unique blocks:",
  n_distinct(filtblocks_BTEXSN$GEOID),
  "\n"
)
```



#get facility-block pairs into one dataframe
```{r}
# ------------------------------------------------------------
# 8. Create individual 10-km buffers for each facility
# ------------------------------------------------------------

tri_buffers_BTEXSN <- tri_sf_proj_BTEXSN %>%
  st_buffer(dist = 10000)


# ------------------------------------------------------------
# 9. Create BTEXSN-specific facility-block directory
# ------------------------------------------------------------

BTEXSN_facility_block_dir <- 
  "~/TSCA/CRA/metdb_tsca/data/facility_blocks_states_BTEXSN"

dir.create(
  BTEXSN_facility_block_dir,
  recursive = TRUE,
  showWarnings = FALSE
)


# ------------------------------------------------------------
# 10. Find BTEXSN facility-block pairs
# ------------------------------------------------------------

pb_BTEXSN_pairs <- progress_bar$new(
  format = "BTEXSN pairs :state [:bar] :percent | :n pairs",
  total = length(valid_states),
  clear = FALSE,
  width = 80
)


facility_block_pairs_BTEXSN <- map_dfr(
  valid_states,
  function(st) {
    
    outfile <- file.path(
      BTEXSN_facility_block_dir,
      paste0("facility_blocks_", st, ".csv")
    )
    
    
    if (file.exists(outfile)) {
      
      existing <- read.csv(
        outfile,
        colClasses = "character"
      )
      
      pb_BTEXSN_pairs$tick(
        tokens = list(
          state = st,
          n = paste0(nrow(existing), " cached")
        )
      )
      
      return(existing)
    }
    
    
    # Load census blocks
    blk <- blocks(
      state = st,
      year = 2020
    ) %>%
      st_transform(5070)
    
    
    # Find facility-block intersections
    fb_pairs <- st_intersection(
      tri_buffers_BTEXSN %>%
        select(trifd),
      blk %>%
        select(GEOID20)
    ) %>%
      st_drop_geometry() %>%
      distinct() %>%
      rename(
        GEOID = GEOID20
      )
    
    
    pb_BTEXSN_pairs$tick(
      tokens = list(
        state = st,
        n = nrow(fb_pairs)
      )
    )
    
    
    write.csv(
      fb_pairs,
      outfile,
      row.names = FALSE
    )
    
    
    fb_pairs
  }
)


# ------------------------------------------------------------
# 11. Combine BTEXSN facility-block pairs
# ------------------------------------------------------------

facility_block_pairs_BTEXSN_all <- 
  facility_block_pairs_BTEXSN %>%
  distinct(
    trifd,
    GEOID
  )


write.csv(
  facility_block_pairs_BTEXSN_all,
  "~/TSCA/CRA/metdb_tsca/data/BTEXSN_facility_block_pairs.csv",
  row.names = FALSE
)


cat(
  "BTEXSN facility-block pairs:",
  nrow(facility_block_pairs_BTEXSN_all),
  "\n"
)

```



### chlor blocks ### 


#pull df of TRI facility IDs, their lat/longs, and NAICS code for facilities releasing chlor, then filter for NAICS codes of interest
```{r}

# ------------------------------------------------------------
# 12. Pull TRI facilities releasing chlorinated chemicals
# ------------------------------------------------------------

SQL_chlor <- "
SELECT DISTINCT
  trifd,
  latitude,
  longitude,
  primary_NAICS,
  NAICS2,
  NAICS3,
  NAICS4,
  NAICS5,
  NAICS6
FROM `edf-aq-data.tri.tri_basic_2017-2022`
WHERE chemical IN (
  'Trichloroethylene',
  'Tetrachloroethylene',
  'Carbon tetrachloride',
  'Dichloromethane',
  'Ethylidene dichloride',
  '1,2-Dichloroethane'
)
"

TRI_chlor <- dbGetQuery(
  con,
  SQL_chlor
)


# ------------------------------------------------------------
# 13. Filter chlor TRI facilities for NAICS codes
# ------------------------------------------------------------

TRI_chlor <- TRI_chlor %>%
  filter(
    primary_NAICS %in% naics_codes |
      NAICS2 %in% naics_codes |
      NAICS3 %in% naics_codes |
      NAICS4 %in% naics_codes |
      NAICS5 %in% naics_codes |
      NAICS6 %in% naics_codes
  )


cat(
  "Chlor unique TRI facilities:",
  n_distinct(TRI_chlor$trifd),
  "\n"
)
```



#get list of blocks overlapping within 10km of these facilities
```{r}
# ------------------------------------------------------------
# 14. Convert chlor TRI facilities to sf
# ------------------------------------------------------------

tri_sf_chlor <- TRI_chlor %>%
  filter(
    !is.na(latitude),
    !is.na(longitude)
  ) %>%
  st_as_sf(
    coords = c("longitude", "latitude"),
    crs = 4326
  )


tri_sf_proj_chlor <- tri_sf_chlor %>%
  st_transform(5070)


# ------------------------------------------------------------
# 15. Create UNION of chlor 10-km facility buffers
# ------------------------------------------------------------

tri_buffer_union_chlor <- tri_sf_proj_chlor %>%
  st_buffer(dist = 10000) %>%
  st_union()


# ------------------------------------------------------------
# 16. Create CHLOR-specific intermediate block directory
# ------------------------------------------------------------

chlor_block_dir <- 
  "~/TSCA/CRA/metdb_tsca/data/filtblocks_states_chlor"

dir.create(
  chlor_block_dir,
  recursive = TRUE,
  showWarnings = FALSE
)


# ------------------------------------------------------------
# 17. Find census blocks within 10 km of chlor facilities
# ------------------------------------------------------------

pb_chlor <- progress_bar$new(
  format = "Chlor :state [:bar] :percent | Found :n blocks",
  total = length(valid_states),
  clear = FALSE,
  width = 80
)


blocks_overlap_chlor <- map_dfr(
  valid_states,
  function(st) {
    
    outfile <- file.path(
      chlor_block_dir,
      paste0("blocks_", st, ".csv")
    )
    
    
    # Only reuse CHLOR-specific cached results
    if (file.exists(outfile)) {
      
      existing <- read.csv(
        outfile,
        colClasses = "character"
      )
      
      pb_chlor$tick(
        tokens = list(
          state = st,
          n = paste0(nrow(existing), " cached")
        )
      )
      
      return(existing)
    }
    
    
    blk <- blocks(
      state = st,
      year = 2020
    ) %>%
      st_transform(5070)
    
    
    # Find blocks intersecting chlor-specific buffers
    blk_overlap <- blk %>%
      filter(
        st_intersects(
          .,
          tri_buffer_union_chlor,
          sparse = FALSE
        )
      )
    
    
    n_found <- nrow(blk_overlap)
    
    
    pb_chlor$tick(
      tokens = list(
        state = st,
        n = n_found
      )
    )
    
    
    df_out <- blk_overlap %>%
      st_drop_geometry() %>%
      select(GEOID20) %>%
      distinct() %>%
      rename(
        GEOID = GEOID20
      )
    
    
    write.csv(
      df_out,
      outfile,
      row.names = FALSE
    )
    
    
    df_out
  }
)


# ------------------------------------------------------------
# 18. Combine and deduplicate chlor blocks
# ------------------------------------------------------------

filtblocks_chlor <- blocks_overlap_chlor %>%
  distinct(GEOID)


write.csv(
  filtblocks_chlor,
  "~/TSCA/CRA/metdb_tsca/data/filtblocks_chlor.csv",
  row.names = FALSE
)


cat(
  "Chlor total unique blocks:",
  n_distinct(filtblocks_chlor$GEOID),
  "\n"
)
```



#get facility-block pairs into one dataframe
```{r}
# ------------------------------------------------------------
# 19. Create individual chlor facility buffers
# ------------------------------------------------------------

tri_buffers_chlor <- tri_sf_proj_chlor %>%
  st_buffer(dist = 10000)


# ------------------------------------------------------------
# 20. Create chlor-specific facility-block directory
# ------------------------------------------------------------

chlor_facility_block_dir <- 
  "~/TSCA/CRA/metdb_tsca/data/facility_blocks_states_chlor"

dir.create(
  chlor_facility_block_dir,
  recursive = TRUE,
  showWarnings = FALSE
)


# ------------------------------------------------------------
# 21. Find chlor facility-block pairs
# ------------------------------------------------------------

pb_chlor_pairs <- progress_bar$new(
  format = "Chlor pairs :state [:bar] :percent | :n pairs",
  total = length(valid_states),
  clear = FALSE,
  width = 80
)


facility_block_pairs_chlor <- map_dfr(
  valid_states,
  function(st) {
    
    outfile <- file.path(
      chlor_facility_block_dir,
      paste0("facility_blocks_", st, ".csv")
    )
    
    
    if (file.exists(outfile)) {
      
      existing <- read.csv(
        outfile,
        colClasses = "character"
      )
      
      pb_chlor_pairs$tick(
        tokens = list(
          state = st,
          n = paste0(nrow(existing), " cached")
        )
      )
      
      return(existing)
    }
    
    
    blk <- blocks(
      state = st,
      year = 2020
    ) %>%
      st_transform(5070)
    
    
    fb_pairs <- st_intersection(
      tri_buffers_chlor %>%
        select(trifd),
      blk %>%
        select(GEOID20)
    ) %>%
      st_drop_geometry() %>%
      distinct() %>%
      rename(
        GEOID = GEOID20
      )
    
    
    pb_chlor_pairs$tick(
      tokens = list(
        state = st,
        n = nrow(fb_pairs)
      )
    )
    
    
    write.csv(
      fb_pairs,
      outfile,
      row.names = FALSE
    )
    
    
    fb_pairs
  }
)


# ------------------------------------------------------------
# 22. Combine chlor facility-block pairs
# ------------------------------------------------------------

facility_block_pairs_chlor_all <- 
  facility_block_pairs_chlor %>%
  distinct(
    trifd,
    GEOID
  )


write.csv(
  facility_block_pairs_chlor_all,
  "~/TSCA/CRA/metdb_tsca/data/chlor_facility_block_pairs.csv",
  row.names = FALSE
)


cat(
  "Chlor facility-block pairs:",
  nrow(facility_block_pairs_chlor_all),
  "\n"
)

```



#number of TRI facilities and blocks for each chemical group
##BTEXSN: 4,802 TRI facilities; 3,416,151 blocks
##chlor: 620 facilities; 1,118,013 blocks
```{r}
filter_summary <- data.frame(
  chemical_group = c(
    "BTEXSN",
    "Chlor"
  ),
  
  n_TRI_facilities = c(
    n_distinct(TRI_BTEXSN$trifd),
    n_distinct(TRI_chlor$trifd)
  ),
  
  n_blocks = c(
    n_distinct(filtblocks_BTEXSN$GEOID),
    n_distinct(filtblocks_chlor$GEOID)
  ),
  
  n_facility_block_pairs = c(
    nrow(facility_block_pairs_BTEXSN_all),
    nrow(facility_block_pairs_chlor_all)
  )
)


filter_summary
```




##pulling chlor HQs/HIs for blocks within 10km of TRI facility 47620GPLSTLEXAN for MC analysis##



#get buffer
```{r}
#load 

#---------------------------------------------------------
# 1. Select the facility of interest
#---------------------------------------------------------

facility_id <- "47620GPLSTLEXAN"

facility_sf <- TRI_chlor %>%
  filter(
    trifd == facility_id,
    !is.na(latitude),
    !is.na(longitude)
  ) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(5070)

#---------------------------------------------------------
# 2. Create 1 km buffer
#---------------------------------------------------------

facility_buffer <- st_buffer(facility_sf, dist = 1000)

#---------------------------------------------------------
# 3. Find intersecting census blocks
#---------------------------------------------------------

valid_states <- c(state.abb, "DC")

blocks_1km <- map_dfr(valid_states, function(st){

  blk <- blocks(state = st, year = 2020) %>%
    st_transform(5070)

  blk %>%
    filter(st_intersects(., facility_buffer, sparse = FALSE)) %>%
    st_drop_geometry() %>%
    transmute(GEOID = GEOID20)

}) %>%
  distinct()

cat("Number of census blocks within 1 km:", nrow(blocks_1km), "\n")

write.csv(
  blocks_1km,
  "~/TSCA/CRA/metdb_tsca/data/blocks_1km_47620GPLSTLEXAN.csv",
  row.names = FALSE
)
```



#pull HQs and HIs
```{r}
#---------------------------------------------------------
# Pull HQs for those blocks
#---------------------------------------------------------

geoid_string <- paste0("'", blocks_1km$GEOID, "'", collapse = ",")

SQL <- paste0("
SELECT *
FROM `edf-aq-data.TSCA.chlor_ATS_HQs`
WHERE block IN (", geoid_string, ")
")

chlor_HQs_1km <- dbGetQuery(con, SQL)

write.csv(
  chlor_HQs_1km,
  "~/TSCA/CRA/metdb_tsca/data/chlor_HQs_1km_47620GPLSTLEXAN.csv",
  row.names = FALSE
)

#---------------------------------------------------------
# Pull HIs for those blocks
#---------------------------------------------------------

SQL <- paste0("
SELECT *
FROM `edf-aq-data.TSCA.chlor_ATS_HIs`
WHERE block IN (", geoid_string, ")
")

chlor_HIs_1km <- dbGetQuery(con, SQL)

write.csv(
  chlor_HIs_1km,
  "~/TSCA/CRA/metdb_tsca/data/chlor_HIs_1km_47620GPLSTLEXAN.csv",
  row.names = FALSE
)
```


#troubleshooting (looks like the blocks aren't in the ATS database before I even made the HQs and HIs)
```{r}
facility_id <- "47620GPLSTLEXAN"

TRI_chlor %>%
  filter(trifd == facility_id)

facility_block_pairs_chlor_all %>%
  filter(trifd == facility_id)

# Compare the independently calculated 1-km blocks with
# the facility's corrected 10-km chlor block list
intersect(
  blocks_1km$GEOID,
  facility_block_pairs_chlor_all %>%
    filter(trifd == facility_id) %>%
    pull(GEOID)
)

# Check whether any of the 1-km blocks exist in ATS HQ data
geoid_string <- paste0(
  "'",
  blocks_1km$GEOID,
  "'",
  collapse = ","
)

SQL <- paste0("
SELECT DISTINCT block
FROM `edf-aq-data.TSCA.chlor_ATS_HQs`
WHERE block IN (", geoid_string, ")
")

dbGetQuery(con, SQL)
```



#checking if the population in these 9 blocks is 0 to see if that's why ATS didn't include (yep, this is the case)
```{r}
# Extract state and county FIPS
block_lookup <- blocks_1km %>%
  mutate(
    state = substr(GEOID, 1, 2),
    county = substr(GEOID, 3, 5)
  )

# Unique state-county combinations
state_counties <- block_lookup %>%
  distinct(state, county)

# Download block populations for each county
pop_data <- map_dfr(seq_len(nrow(state_counties)), function(i){

  st <- state_counties$state[i]
  co <- state_counties$county[i]

  get_decennial(
    geography = "block",
    variables = "P1_001N",
    year = 2020,
    sumfile = "pl",
    state = st,
    county = co,
    geometry = FALSE
  )
})

# Join to your blocks
pop_check <- block_lookup %>%
  left_join(
    pop_data %>%
      select(GEOID, population = value),
    by = "GEOID"
  )

pop_check
```


#pull HQs and HIs for blocks within 10km
```{r}

facility_id <- "47620GPLSTLEXAN"


# ------------------------------------------------------------
# 1. Get corrected chlor blocks within 10 km of facility
# ------------------------------------------------------------

blocks_10km <- facility_block_pairs_chlor_all %>%
  filter(
    trifd == facility_id
  ) %>%
  distinct(
    GEOID
  )


# Check number of blocks
cat(
  "Number of census blocks within 10 km:",
  nrow(blocks_10km),
  "\n"
)

head(blocks_10km)


# ------------------------------------------------------------
# 2. Create GEOID string for BigQuery
# ------------------------------------------------------------

geoid_string <- paste0(
  "'",
  blocks_10km$GEOID,
  "'",
  collapse = ","
)


# ------------------------------------------------------------
# 3. Pull chlor HQs
# ------------------------------------------------------------

SQL_hq <- paste0("
SELECT *
FROM `edf-aq-data.TSCA.chlor_ATS_HQs`
WHERE block IN (", geoid_string, ")
")


chlor_HQs_10km <- dbGetQuery(
  con,
  SQL_hq
)


# ------------------------------------------------------------
# 4. Keep methylene chloride HQs
# ------------------------------------------------------------

chlor_HQs_10km_mc <- chlor_HQs_10km %>%
  filter(
    pollutant_name == "METHYLENE CHLORIDE"
  )


write.csv(
  chlor_HQs_10km_mc,
  "~/TSCA/CRA/metdb_tsca/output/chlor/chlor_HQs_10km_mc_47620GPLSTLEXAN.csv",
  row.names = FALSE
)


# ------------------------------------------------------------
# 5. Pull chlor HIs
# ------------------------------------------------------------

SQL_hi <- paste0("
SELECT *
FROM `edf-aq-data.TSCA.chlor_ATS_HIs`
WHERE block IN (", geoid_string, ")
")


chlor_HIs_10km <- dbGetQuery(
  con,
  SQL_hi
)


write.csv(
  chlor_HIs_10km,
  "~/TSCA/CRA/metdb_tsca/output/chlor/chlor_HIs_10km_47620GPLSTLEXAN.csv",
  row.names = FALSE
)


# ------------------------------------------------------------
# 6. Checks
# ------------------------------------------------------------

cat(
  "Blocks geometrically within 10 km:",
  nrow(blocks_10km),
  "\n"
)

cat(
  "Total chlor HQ rows returned:",
  nrow(chlor_HQs_10km),
  "\n"
)

cat(
  "Unique blocks represented in chlor HQ data:",
  n_distinct(chlor_HQs_10km$block),
  "\n"
)

cat(
  "Unique blocks represented in methylene chloride HQ data:",
  n_distinct(chlor_HQs_10km_mc$block),
  "\n"
)
```



### formatting and uploading relevant info to BQ for downstream analysis ###



#database connection
```{r}
bq_auth(email = NA)

con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "airtoxscreen",)

#project_id <- "edf-aq-data"
```



#format rfc data to add back to BQ for calculations 
```{r}
rfcs <- readr::read_csv("~/TSCA/CRA/metdb_tsca/data/rfcs_final.csv")
str(rfcs)

#filter RfCs for our analysis
rfcs <- rfcs %>%
  select(casrn, name, average_mass, endpoint, new_rfc) %>%
  filter(casrn %in% c("71-43-2", "100-41-4", "108-88-3", "1330-20-7", 
                      "100-42-5", "91-20-3", "79-01-6", "127-18-4", 
                      "56-23-5", "75-09-2", "75-34-3", "107-06-2"))

#RfCs to ug/m3 to match ATS data
rfcs$rfcs_ugm <- (rfcs$new_rfc * rfcs$average_mass) / 24.45

#add ATS names to RfC df
rfcs <- rfcs %>%
  mutate(ATS_name = case_when(
    name == "Benzene" ~ "BENZENE",
    name == "Ethylbenzene" ~ "ETHYLBENZENE",
    name == "Toluene" ~ "TOLUENE",
    name == "Styrene" ~ "STYRENE",
    name == "Naphthalene" ~ "NAPHTHALENE",
    name == "Xylenes" ~ "XYLENES (MIXED ISOMERS)",
    name == "Trichloroethylene" ~ "TRICHLOROETHYLENE",
    name == "Tetrachloroethylene" ~ "TETRACHLOROETHYLENE",
    name == "Carbon tetrachloride" ~ "CARBON TETRACHLORIDE",
    name == "Dichloromethane (methylene chloride)" ~ "METHYLENE CHLORIDE",
    name == "1,1-Dichloroethane" ~ "ETHYLIDENE DICHLORIDE (1,1-DICHLOROETHANE)",
    name == "1,2-Dichloroethane" ~ "ETHYLENE DICHLORIDE (1,2-DICHLOROETHANE)",
    TRUE ~ NA_character_
  ))

#save new rfc csv to do joins and calculations in BQ
write.csv(rfcs, file = "~/TSCA/CRA/metdb_tsca/data/rfcs_final.csv")

```


#upload a BTEXSN filtblocks table to BQ to filter downstream analysis for those blocks of interest 
```{r}
filtblocks = read.csv("~/TSCA/CRA/metdb_tsca/data/filtblocks_BTEXSN.csv")

project <- "edf-aq-data"
dataset <- "TSCA"  
table   <- "filtblocks_BTEXSN"

tbl <- bq_table(project, dataset, table)

# Pad GEOIDs and upload filtblocks_BTEXSN
filtblocks_BTEXSN_upload <- filtblocks %>%
  mutate(GEOID = str_pad(as.character(GEOID), width = 15, side = "left", pad = "0")) %>%
  select(GEOID)

stopifnot(all(nchar(filtblocks_BTEXSN_upload$GEOID) == 15))

bq_table_upload(
  x = tbl,
  values = filtblocks_BTEXSN_upload,
  create_disposition = "CREATE_IF_NEEDED",
  write_disposition = "WRITE_TRUNCATE",
  fields = list(
    bq_field("GEOID", "STRING")
  )
)

```


#upload a chlor filtblocks table to BQ to filter downstream analysis for those blocks of interest
```{r}
project <- "edf-aq-data"
dataset <- "TSCA"  
table   <- "filtblocks_chlor"

tbl <- bq_table(project, dataset, table)

# Pad GEOIDs and upload filtblocks_chlor
filtblocks_chlor_upload <- filtblocks %>%
  mutate(GEOID = str_pad(as.character(GEOID), width = 15, side = "left", pad = "0")) %>%
  select(GEOID)

stopifnot(all(nchar(filtblocks_chlor_upload$GEOID) == 15))

bq_table_upload(
  x = tbl,
  values = filtblocks_chlor_upload,
  create_disposition = "CREATE_IF_NEEDED",
  write_disposition = "WRITE_TRUNCATE",
  fields = list(
    bq_field("GEOID", "STRING")
  )
)

```





