---
title: "Filtering for blocks with TSCA facilities"
author: "Paige Varner"
date: "2025-09-04"
output: html_document
---


#load libraries
```{r}
library("dplyr")
library("tidyr")
library("bigrquery")
library("DBI")
library("tidycensus")
library("sf")
library("purrr")
library("tigris")
library("progress")
options(tigris_use_cache = TRUE)

naics = read.csv("~/TSCA/CRA/metdb_tsca/data/naics_tri_filter.csv")
```


#database connection
```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "airtoxscreen",)

bq_auth(email = NA)
```


### BTEXSN blocks ###

#pull df of TRI facility IDs, their lat/longs, and NAICS code for facilities releasing BTEXSN, then filter for NAICS codes of interest
```{r}
#pull df from BQ
SQL <- "
SELECT DISTINCT
  trifd,
  latitude,
  longitude,
  primary_NAICS,
  NAICS2,
  NAICS3,
  NAICS4,
  NAICS5,
  NAICS6
FROM `edf-aq-data.tri.tri_basic_2017-2022`
WHERE chemical IN (
  'Benzene',
  'Toluene',
  'Ethylbenzene',
  'm-Xylene',
  'p-Xylene',
  'o-Xylene',
  'Xylene (mixed isomers)',
  'Styrene',
  'Naphthalene'
)
"

TRI_BTEXSN = dbGetQuery(con, SQL)

#filter for NAICS codes
naics_codes <- naics$Code %>% as.character()

TRI_BTEXSN <- TRI_BTEXSN %>%
  filter(
    primary_NAICS %in% naics_codes |
    NAICS2 %in% naics_codes |
    NAICS3 %in% naics_codes |
    NAICS4 %in% naics_codes |
    NAICS5 %in% naics_codes |
    NAICS6 %in% naics_codes
  )

```


#get list of blocks overlapping within 10km of any of these facilities
```{r}
# 1. TRI facilities into sf
tri_sf <- TRI_BTEXSN %>%
  filter(!is.na(latitude), !is.na(longitude)) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# 2. Project to CRS in meters (good for buffering)
tri_sf_proj <- st_transform(tri_sf, 5070)

# 3. 10 km buffer, union into one geometry (for block-only overlap check)
tri_buffer_union <- st_buffer(tri_sf_proj, dist = 10000) %>%
  st_union()

# 4. Valid states: 50 states + DC only
valid_states <- c(state.abb, "DC")

# 5. Setup progress bar
pb <- progress_bar$new(
  format = "Processing :state [:bar] :percent | Found :n blocks",
  total = length(valid_states), clear = FALSE, width = 80
)

# Directory to store intermediate results
dir.create("~/TSCA/CRA/metdb_tsca/data/filtblocks_states", showWarnings = FALSE)

# 6. Loop state-by-state with intermediate save (blocks only)
map_dfr(valid_states, function(st) {
  # Skip if already processed
  outfile <- file.path("~/TSCA/CRA/metdb_tsca/data/filtblocks_states", paste0("blocks_", st, ".csv"))
  if (file.exists(outfile)) {
    pb$tick(tokens = list(state = st, n = "skipped"))
    return(read.csv(outfile, colClasses = "character"))
  }
  
  # Load blocks for one state
  blk <- blocks(state = st, year = 2020) %>%
    st_transform(5070)
  
  # Find overlaps with union of all buffers
  blk_overlap <- blk %>%
    filter(st_intersects(., tri_buffer_union, sparse = FALSE))
  
  n_found <- nrow(blk_overlap)
  
  # Update progress bar
  pb$tick(tokens = list(state = st, n = n_found))
  
  # Prepare output
  df_out <- blk_overlap %>%
    st_drop_geometry() %>%
    select(GEOID20) %>%
    distinct() %>%
    rename(GEOID = GEOID20)
  
  # Save intermediate
  write.csv(df_out, outfile, row.names = FALSE)
  
  df_out
}) -> blocks_overlap

# 7. Combine all states + deduplicate
filtblocks_BTEXSN <- blocks_overlap %>% distinct()

# 8. Save final combined file (block-only)
write.csv(filtblocks_BTEXSN, 
          file = "~/TSCA/CRA/metdb_tsca/data/filtblocks_BTEXSN.csv", 
          row.names = FALSE)

# 9. Optional sanity check
cat("Total unique blocks found:", nrow(filtblocks_BTEXSN), "\n")

# ------------------------------------------------------------------
# 10. NEW STEP: Facility-block pairs
# ------------------------------------------------------------------

# 10a. Create buffers per facility (no union)
tri_buffers <- st_buffer(tri_sf_proj, dist = 10000)

# Directory for facility-block state outputs
dir.create("~/TSCA/CRA/metdb_tsca/data/facility_blocks_states", showWarnings = FALSE)

# 10b. Loop state by state, save facility-block pairs
map_dfr(valid_states, function(st) {
  outfile <- file.path("~/TSCA/CRA/metdb_tsca/data/facility_blocks_states", 
                       paste0("facility_blocks_", st, ".csv"))
  
  if (file.exists(outfile)) {
    return(read.csv(outfile, colClasses = "character"))
  }
  
  # Load state blocks
  blk <- blocks(state = st, year = 2020) %>%
    st_transform(5070)
  
  # Intersections: gives facility–block matches
  fb_pairs <- st_intersection(
    tri_buffers %>% select(trifd),
    blk %>% select(GEOID20)
  ) %>%
    st_drop_geometry() %>%
    distinct() %>%
    rename(GEOID = GEOID20)
  
  # Save state-level CSV
  write.csv(fb_pairs, outfile, row.names = FALSE)
  
  fb_pairs
}) -> facility_block_pairs

# 10c. Combine all states
facility_block_pairs_all <- facility_block_pairs %>% distinct()

# 10d. Save final combined facility-block file
write.csv(facility_block_pairs_all, 
          file = "~/TSCA/CRA/metdb_tsca/data/BTEXSN_facility_block_pairs.csv", 
          row.names = FALSE)

cat("Total facility-block pairs found:", nrow(facility_block_pairs_all), "\n")

```


### chlor blocks ### (this needs to be updated to reflect what was done above for BTEXSN)


#pull df of TRI facility IDs, their lat/longs, and NAICS code for facilities releasing BTEXSN, then filter for NAICS codes of interest
```{r}
#pull df from BQ
SQL <- "
SELECT DISTINCT
  trifd,
  latitude,
  longitude,
  primary_NAICS,
  NAICS2,
  NAICS3,
  NAICS4,
  NAICS5,
  NAICS6
FROM `edf-aq-data.tri.tri_basic_2017-2022`
WHERE chemical IN (
  'Trichloroethylene',
  'Tetrachloroethylene',
  'Carbon tetrachloride',
  'Dichloromethane',
  'Ethylidene dichloride',
  '1,2-Dichloroethane'
)
"

TRI_chlor = dbGetQuery(con, SQL)

#filter for NAICS codes
naics_codes <- naics$Code %>% as.character()

TRI_chlor <- TRI_chlor %>%
  filter(
    primary_NAICS %in% naics_codes |
    NAICS2 %in% naics_codes |
    NAICS3 %in% naics_codes |
    NAICS4 %in% naics_codes |
    NAICS5 %in% naics_codes |
    NAICS6 %in% naics_codes
  )
```



#get list of blocks overlapping within 10km of any of these facilities
```{r}
# 1. TRI facilities into sf
tri_sf <- TRI_chlor %>%
  filter(!is.na(latitude), !is.na(longitude)) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# 2. Project to CRS in meters (good for buffering)
tri_sf_proj <- st_transform(tri_sf, 5070)

# 3. 10 km buffer, union into one geometry
tri_buffer_union <- st_buffer(tri_sf_proj, dist = 10000) %>%
  st_union()

# 4. Valid states: 50 states + DC only
valid_states <- c(state.abb, "DC")

# 5. Setup progress bar
pb <- progress_bar$new(
  format = "Processing :state [:bar] :percent | Found :n blocks",
  total = length(valid_states), clear = FALSE, width = 80
)

# Directory to store intermediate results
dir.create("~/TSCA/CRA/metdb_tsca/data/filtblocks_states_chlor", showWarnings = FALSE)

# 6. Loop state-by-state with intermediate save
map_dfr(valid_states, function(st) {
  # Skip if already processed
  outfile <- file.path("~/TSCA/CRA/metdb_tsca/data/filtblocks_states_chlor", paste0("blocks_", st, ".csv"))
  if (file.exists(outfile)) {
    pb$tick(tokens = list(state = st, n = "skipped"))
    return(read.csv(outfile, colClasses = "character"))
  }
  
  # Load blocks for one state
  blk <- blocks(state = st, year = 2020) %>%
    st_transform(5070)
  
  # Find overlaps
  blk_overlap <- blk %>%
    filter(st_intersects(., tri_buffer_union, sparse = FALSE))
  
  n_found <- nrow(blk_overlap)
  
  # Update progress bar
  pb$tick(tokens = list(state = st, n = n_found))
  
  # Prepare output
  df_out <- blk_overlap %>%
    st_drop_geometry() %>%
    select(GEOID20) %>%
    distinct() %>%
    rename(GEOID = GEOID20)
  
  # Save intermediate
  write.csv(df_out, outfile, row.names = FALSE)
  
  df_out
}) -> blocks_overlap

# 7. Combine all states + deduplicate
filtblocks_chlor <- blocks_overlap %>% distinct()

# 8. Save final combined file
write.csv(filtblocks_chlor, 
          file = "~/TSCA/CRA/metdb_tsca/data/filtblocks_chlor.csv", 
          row.names = FALSE)

# 9. Optional sanity check
cat("Total unique blocks found:", nrow(filtblocks_chlor), "\n")
```









