---
title: "HI calculations for BTEXSN"
author: "Paige Varner"
date: "2025-04-18"
output: html_document
---


#load libraries
```{r}
library("dplyr")
library("tidyr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("car")
library("arrow")
library("progressr")
library("purrr")
library("readr")
library("haven")
library("stringr")
library("ggokabeito")
library("magick")
library("cowplot")
detach(package:Rmisc)
detach(package:plyr)
```


#load data, if needed
```{r}
rfcs <- readr::read_csv("~/TSCA/CRA/metdb_tsca/data/rfcs_final.csv")
str(rfcs)

filtblocks = read.csv("~/TSCA/CRA/metdb_tsca/data/filtblocks_BTEXSN.csv")
```


#format rfc data to add back to BQ for calculations (don't have to do if already done)
```{r}
#filter RfCs for our analysis
rfcs <- rfcs %>%
  select(casrn, name, average_mass, endpoint, new_rfc) %>%
  filter(casrn %in% c("71-43-2", "100-41-4", "108-88-3", "1330-20-7", 
                      "100-42-5", "91-20-3", "79-01-6", "127-18-4", 
                      "56-23-5", "75-09-2", "75-34-3", "107-06-2"))

#RfCs to ug/m3 to match ATS data
rfcs$rfcs_ugm <- (rfcs$new_rfc * rfcs$average_mass) / 24.45

#add ATS names to RfC df
rfcs <- rfcs %>%
  mutate(ATS_name = case_when(
    name == "Benzene" ~ "BENZENE",
    name == "Ethylbenzene" ~ "ETHYLBENZENE",
    name == "Toluene" ~ "TOLUENE",
    name == "Styrene" ~ "STYRENE",
    name == "Naphthalene" ~ "NAPHTHALENE",
    name == "Xylenes" ~ "XYLENES (MIXED ISOMERS)",
    name == "Trichloroethylene" ~ "TRICHLOROETHYLENE",
    name == "Tetrachloroethylene" ~ "TETRACHLOROETHYLENE",
    name == "Carbon tetrachloride" ~ "CARBON TETRACHLORIDE",
    name == "Dichloromethane (methylene chloride)" ~ "METHYLENE CHLORIDE",
    name == "1,1-Dichloroethane" ~ "ETHYLIDENE DICHLORIDE (1,1-DICHLOROETHANE)",
    name == "1,2-Dichloroethane" ~ "ETHYLENE DICHLORIDE (1,2-DICHLOROETHANE)",
    TRUE ~ NA_character_
  ))

#save new rfc csv to do joins and calculations in BQ
write.csv(rfcs, file = "~/TSCA/CRA/metdb_tsca/data/rfcs_final.csv")

```


#database connection
```{r}
bq_auth(email = NA)

con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "airtoxscreen",)

#project_id <- "edf-aq-data"
```


#upload a filtblocks table to BQ to filter downstream analysis for those blocks of interest (don't have to run if already done)
```{r}
project <- "edf-aq-data"
dataset <- "TSCA"  
table   <- "filtblocks_BTEXSN"

tbl <- bq_table(project, dataset, table)

# Pad GEOIDs and upload filtblocks_BTEXSN
filtblocks_BTEXSN_upload <- filtblocks %>%
  mutate(GEOID = str_pad(as.character(GEOID), width = 15, side = "left", pad = "0")) %>%
  select(GEOID)

stopifnot(all(nchar(filtblocks_BTEXSN_upload$GEOID) == 15))

bq_table_upload(
  x = tbl,
  values = filtblocks_BTEXSN_upload,
  create_disposition = "CREATE_IF_NEEDED",
  write_disposition = "WRITE_TRUNCATE",
  fields = list(
    bq_field("GEOID", "STRING")
  )
)

```


#pull summary information on HIs (note: the N in these summaries is for block + endpoint combinations)
```{r}
# ---  Overall summary table ---
SQL <- "
SELECT 
  'HI_tot' AS HI_type,
  MIN(HI_tot) AS min,
  MAX(HI_tot) AS max,
  AVG(HI_tot) AS mean,
  APPROX_QUANTILES(HI_tot, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot) AS std_dev,
  STDDEV(HI_tot) / SQRT(COUNT(HI_tot)) AS std_error,
  COUNT(HI_tot) AS n
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID

UNION ALL

SELECT
  'HI_pt',
  MIN(HI_pt),
  MAX(HI_pt),
  AVG(HI_pt),
  APPROX_QUANTILES(HI_pt, 2)[OFFSET(1)],
  STDDEV(HI_pt),
  STDDEV(HI_pt) / SQRT(COUNT(HI_pt)),
  COUNT(HI_pt)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID

UNION ALL

SELECT
  'HI_other',
  MIN(HI_other),
  MAX(HI_other),
  AVG(HI_other),
  APPROX_QUANTILES(HI_other, 2)[OFFSET(1)],
  STDDEV(HI_other),
  STDDEV(HI_other) / SQRT(COUNT(HI_other)),
  COUNT(HI_other)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID;
"

BTEXSN_HI_summary <- dbGetQuery(con, SQL)
write.csv(BTEXSN_HI_summary, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HIsummary.csv")


# ---  EPA region summary table ---
SQL <- "
SELECT
  epa_region AS group_value,
  'HI_tot' AS HI_type,
  MIN(HI_tot) AS min,
  MAX(HI_tot) AS max,
  AVG(HI_tot) AS mean,
  APPROX_QUANTILES(HI_tot, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot) AS std_dev,
  STDDEV(HI_tot) / SQRT(COUNT(HI_tot)) AS std_error,
  COUNT(HI_tot) AS n
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY epa_region

UNION ALL

SELECT
  epa_region,
  'HI_pt',
  MIN(HI_pt),
  MAX(HI_pt),
  AVG(HI_pt),
  APPROX_QUANTILES(HI_pt, 2)[OFFSET(1)],
  STDDEV(HI_pt),
  STDDEV(HI_pt) / SQRT(COUNT(HI_pt)),
  COUNT(HI_pt)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY epa_region

UNION ALL

SELECT
  epa_region,
  'HI_other',
  MIN(HI_other),
  MAX(HI_other),
  AVG(HI_other),
  APPROX_QUANTILES(HI_other, 2)[OFFSET(1)],
  STDDEV(HI_other),
  STDDEV(HI_other) / SQRT(COUNT(HI_other)),
  COUNT(HI_other)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY epa_region
ORDER BY group_value, HI_type;
"

BTEXSN_HI_summary_region <- dbGetQuery(con, SQL)
write.csv(BTEXSN_HI_summary_region, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HIsummary_region.csv")


# ---  State summary table ---
SQL <- "
SELECT
  state AS group_value,
  'HI_tot' AS HI_type,
  MIN(HI_tot) AS min,
  MAX(HI_tot) AS max,
  AVG(HI_tot) AS mean,
  APPROX_QUANTILES(HI_tot, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot) AS std_dev,
  STDDEV(HI_tot) / SQRT(COUNT(HI_tot)) AS std_error,
  COUNT(HI_tot) AS n
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY state

UNION ALL

SELECT
  state,
  'HI_pt',
  MIN(HI_pt),
  MAX(HI_pt),
  AVG(HI_pt),
  APPROX_QUANTILES(HI_pt, 2)[OFFSET(1)],
  STDDEV(HI_pt),
  STDDEV(HI_pt) / SQRT(COUNT(HI_pt)),
  COUNT(HI_pt)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY state

UNION ALL

SELECT
  state,
  'HI_other',
  MIN(HI_other),
  MAX(HI_other),
  AVG(HI_other),
  APPROX_QUANTILES(HI_other, 2)[OFFSET(1)],
  STDDEV(HI_other),
  STDDEV(HI_other) / SQRT(COUNT(HI_other)),
  COUNT(HI_other)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY state
ORDER BY group_value, HI_type;
"

BTEXSN_HI_summary_state <- dbGetQuery(con, SQL)
write.csv(BTEXSN_HI_summary_state, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HIsummary_state.csv")


# ---  Endpoint summary table ---
SQL <- "
SELECT
  endpoint AS group_value,
  'HI_tot' AS HI_type,
  MIN(HI_tot) AS min,
  MAX(HI_tot) AS max,
  AVG(HI_tot) AS mean,
  APPROX_QUANTILES(HI_tot, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot) AS std_dev,
  STDDEV(HI_tot) / SQRT(COUNT(HI_tot)) AS std_error,
  COUNT(HI_tot) AS n
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY endpoint

UNION ALL

SELECT
  endpoint,
  'HI_pt',
  MIN(HI_pt),
  MAX(HI_pt),
  AVG(HI_pt),
  APPROX_QUANTILES(HI_pt, 2)[OFFSET(1)],
  STDDEV(HI_pt),
  STDDEV(HI_pt) / SQRT(COUNT(HI_pt)),
  COUNT(HI_pt)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY endpoint

UNION ALL

SELECT
  endpoint,
  'HI_other',
  MIN(HI_other),
  MAX(HI_other),
  AVG(HI_other),
  APPROX_QUANTILES(HI_other, 2)[OFFSET(1)],
  STDDEV(HI_other),
  STDDEV(HI_other) / SQRT(COUNT(HI_other)),
  COUNT(HI_other)
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f ON t.block = f.GEOID
GROUP BY endpoint
ORDER BY group_value, HI_type;
"

BTEXSN_HI_summary_endpoint <- dbGetQuery(con, SQL)
write.csv(BTEXSN_HI_summary_endpoint, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HIsummary_endpoint.csv")

```


#delta for each pollutant
```{r}
# overall deltas
SQL <- "
SELECT
  hq.pollutant_name,

  AVG(hq.HQ_tot) AS avg_HQ_tot,
  AVG(hi.HI_tot) AS avg_HI_tot,
  SAFE_DIVIDE(AVG(hi.HI_tot) - AVG(hq.HQ_tot), 1) AS delta_avg_tot,

  AVG(hq.HQ_pt) AS avg_HQ_pt,
  AVG(hi.HI_pt) AS avg_HI_pt,
  SAFE_DIVIDE(AVG(hi.HI_pt) - AVG(hq.HQ_pt), 1) AS delta_avg_pt,

  AVG(hq.HQ_other) AS avg_HQ_other,
  AVG(hi.HI_other) AS avg_HI_other,
  SAFE_DIVIDE(AVG(hi.HI_other) - AVG(hq.HQ_other), 1) AS delta_avg_other,

  COUNT(*) AS n

FROM `edf-aq-data.TSCA.BTEXSN_ATS_HQs` AS hq
JOIN `edf-aq-data.TSCA.BTEXSN_ATS_HIs` AS hi
  ON hq.block = hi.block
     AND hq.endpoint = hi.endpoint
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON hq.block = f.GEOID

GROUP BY hq.pollutant_name
ORDER BY hq.pollutant_name
"

BTEXSN_pollutant_avg_deltas <- dbGetQuery(con, SQL)
write.csv(BTEXSN_pollutant_avg_deltas, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/avg_deltas.csv")


# deltas for blocks with HIs > 1
SQL <- "
WITH blocks_with_high_HI AS (
  SELECT DISTINCT block, endpoint
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.block = f.GEOID
  WHERE HI_tot > 1 OR HI_pt > 1 OR HI_other > 1
)

SELECT
  hq.pollutant_name,

  AVG(hq.HQ_tot) AS avg_HQ_tot,
  AVG(hi.HI_tot) AS avg_HI_tot,
  SAFE_DIVIDE(AVG(hi.HI_tot) - AVG(hq.HQ_tot), 1) AS delta_avg_tot,

  AVG(hq.HQ_pt) AS avg_HQ_pt,
  AVG(hi.HI_pt) AS avg_HI_pt,
  SAFE_DIVIDE(AVG(hi.HI_pt) - AVG(hq.HQ_pt), 1) AS delta_avg_pt,

  AVG(hq.HQ_other) AS avg_HQ_other,
  AVG(hi.HI_other) AS avg_HI_other,
  SAFE_DIVIDE(AVG(hi.HI_other) - AVG(hq.HQ_other), 1) AS delta_avg_other,

  COUNT(*) AS n

FROM `edf-aq-data.TSCA.BTEXSN_ATS_HQs` AS hq
JOIN `edf-aq-data.TSCA.BTEXSN_ATS_HIs` AS hi
  ON hq.block = hi.block
     AND hq.endpoint = hi.endpoint
JOIN blocks_with_high_HI b
  ON hq.block = b.block
     AND hq.endpoint = b.endpoint
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON hq.block = f.GEOID

GROUP BY hq.pollutant_name
ORDER BY hq.pollutant_name
"

BTEXSN_pollutant_avg_deltas_1 <- dbGetQuery(con, SQL)
write.csv(BTEXSN_pollutant_avg_deltas_1, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI1/avg_deltas_HI1.csv")


# deltas for blocks with HIs > 0.1
SQL <- "
WITH blocks_with_high_HI AS (
  SELECT DISTINCT block, endpoint
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.block = f.GEOID
  WHERE HI_tot > 0.1 OR HI_pt > 0.1 OR HI_other > 0.1
)

SELECT
  hq.pollutant_name,

  AVG(hq.HQ_tot) AS avg_HQ_tot,
  AVG(hi.HI_tot) AS avg_HI_tot,
  SAFE_DIVIDE(AVG(hi.HI_tot) - AVG(hq.HQ_tot), 1) AS delta_avg_tot,

  AVG(hq.HQ_pt) AS avg_HQ_pt,
  AVG(hi.HI_pt) AS avg_HI_pt,
  SAFE_DIVIDE(AVG(hi.HI_pt) - AVG(hq.HQ_pt), 1) AS delta_avg_pt,

  AVG(hq.HQ_other) AS avg_HQ_other,
  AVG(hi.HI_other) AS avg_HI_other,
  SAFE_DIVIDE(AVG(hi.HI_other) - AVG(hq.HQ_other), 1) AS delta_avg_other,

  COUNT(*) AS n

FROM `edf-aq-data.TSCA.BTEXSN_ATS_HQs` AS hq
JOIN `edf-aq-data.TSCA.BTEXSN_ATS_HIs` AS hi
  ON hq.block = hi.block
     AND hq.endpoint = hi.endpoint
JOIN blocks_with_high_HI b
  ON hq.block = b.block
     AND hq.endpoint = b.endpoint
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON hq.block = f.GEOID

GROUP BY hq.pollutant_name
ORDER BY hq.pollutant_name
"

BTEXSN_pollutant_avg_deltas_01 <- dbGetQuery(con, SQL)
write.csv(BTEXSN_pollutant_avg_deltas_01, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/avg_deltas_HI01.csv")

```


#CDFs for HI_tot and HI_pt
```{r}
# --- 1. Pull the data ---
project_id <- "edf-aq-data"

SQL <- "
SELECT
  t.block,
  t.endpoint,
  t.HI_tot,
  t.HI_pt
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON t.block = f.GEOID
"

BTEXSN_ATS_HIs_filtered <- bq_project_query(project_id, SQL) %>%
  bq_table_download()

# --- 2. Reshape for plotting ---
BTEXSN_long <- BTEXSN_ATS_HIs_filtered %>%
  pivot_longer(
    cols = c(HI_tot, HI_pt),
    names_to = "HI_type",
    values_to = "HI_value"
  )

# --- 3. Pre-compute CDF ---
BTEXSN_cdf <- BTEXSN_long %>%
  group_by(endpoint, HI_type) %>%
  arrange(HI_value) %>%
  mutate(cdf = row_number() / n()) %>%
  ungroup()

#Optional: thin points for plotting (e.g., keep every 10th point)
BTEXSN_cdf_sample <- BTEXSN_cdf %>%
  group_by(endpoint, HI_type) %>%
  slice(seq(1, n(), by = 10)) %>%
  ungroup()

# --- 4. Plot ---
cdf_plot <- ggplot(
  BTEXSN_cdf_sample,
  aes(x = HI_value, y = cdf, group = endpoint, color = endpoint)

) +
  geom_line() +
  geom_vline(xintercept = c(0.1, 1), linetype = "dotted", color = "black") +
  scale_x_log10() +
  labs(
    title = "CDF of HI_tot and HI_pt",
    x = "HI value (log scale)",
    y = "P(HI ≤ x)"
  ) +
  facet_wrap(~HI_type, scales = "free_y") +
  theme_minimal(base_size = 14)


cdf_plot

ggsave(filename = "BTEXSN_cdf.png", device = "png", plot = cdf_plot, path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/", width = 7, height = 5)

```



#######################################

### analysis with HI threshold of 1

#######################################




#get count of blocks with various HIs above 1
```{r}
SQL <- "
WITH filtered_HIs AS (
  SELECT *
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.block = f.GEOID
)

SELECT
  'HI_tot' AS HI_type,
  COUNT(DISTINCT CASE WHEN HI_tot > 1 THEN block END) AS unique_blocks_above_1,
  COUNT(DISTINCT block) AS total_unique_blocks,
  COUNT(CASE WHEN HI_tot > 1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs

UNION ALL

SELECT
  'HI_pt',
  COUNT(DISTINCT CASE WHEN HI_pt > 1 THEN block END),
  COUNT(DISTINCT block),
  COUNT(CASE WHEN HI_pt > 1 THEN 1 END),
  COUNT(*)
FROM filtered_HIs

UNION ALL

SELECT
  'HI_other',
  COUNT(DISTINCT CASE WHEN HI_other > 1 THEN block END),
  COUNT(DISTINCT block),
  COUNT(CASE WHEN HI_other > 1 THEN 1 END),
  COUNT(*)
FROM filtered_HIs;
"

BTEXSN_blocks_HIsabove1 <- dbGetQuery(con, SQL)

write.csv(BTEXSN_blocks_HIsabove1, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI1/blocks_HIsabove1_summary.csv",
          row.names = FALSE)

```


#pull df of HIs >1 for further analysis and HQs for those blocks
```{r}
#pull blocks with HIs > 1 filtered by filtblocks_BTEXSN
SQL = "
SELECT t.*
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON t.block = f.GEOID
WHERE HI_tot > 1 OR HI_pt > 1 OR HI_other > 1;
"

BTEXSN_exceedances = dbGetQuery(con, SQL)

#pull HQ data for those blocks
block_ids = unique(BTEXSN_exceedances$block)

block_list_sql = paste0("'", block_ids, "'", collapse = ", ")

SQL = paste0("
  SELECT *
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HQs`
  WHERE block IN (", block_list_sql, ")
")

BTEXSN_HQs_HIexceedances = dbGetQuery(con, SQL)
```


# how many blocks exceed more than one health endpoint threshold
```{r}
blocks_mult_endpoints <- BTEXSN_exceedances %>%
  group_by(block) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(unique(endpoint), collapse = ", ")
  ) %>%
  filter(n_endpoints_exceeding > 1)

BTEXSN_mult_endpoints <- BTEXSN_exceedances %>%
  semi_join(blocks_mult_endpoints, by = "block") %>%
  left_join(blocks_mult_endpoints, by = "block")

write.csv(blocks_mult_endpoints, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI1/blocks_mult_endpoints.csv")
```



#what chemicals are contributing to HI exceedances - and would these exceedances exist without considering cumulative risk?
```{r}
# join HQ and HI datasets together
BTEXSN_exceedances_with_HQs <- BTEXSN_exceedances %>%
  left_join(
    BTEXSN_HQs_HIexceedances %>% select(block, endpoint, HQ_tot, HQ_pt, HQ_other, pollutant_name),
    by = c("block", "endpoint")
  )

# filter for blocks where HIs are greater than 1 but HQs are not (no blocks of this type - no cumulative risk consideration)
BTEXSN_exceedances_HI_only <- BTEXSN_exceedances_with_HQs %>%
  group_by(block, endpoint) %>%
  filter(
    (any(HI_tot > 1) & all(is.na(HQ_tot) | HQ_tot <= 1)) |
    (any(HI_pt > 1) & all(is.na(HQ_pt) | HQ_pt <= 1)) |
    (any(HI_other > 1) & all(is.na(HQ_other) | HQ_other <= 1))
  ) %>%
  mutate(
    HI_exceeding_type = case_when(
      HI_tot > 1 & (is.na(HQ_tot) | HQ_tot <= 1) ~ "HI_tot",
      HI_pt  > 1 & (is.na(HQ_pt)  | HQ_pt  <= 1) ~ "HI_pt",
      HI_other > 1 & (is.na(HQ_other) | HQ_other <= 1) ~ "HI_other",
      TRUE ~ NA_character_
    ),
    max_HI = pmax(HI_tot, HI_pt, HI_other, na.rm = TRUE)
  ) %>%
  ungroup()

```


#bar graph average HQs for each endpoint that was exceeded in any block group to see which chemicals contribute 
```{r}
# get averages of HQs for each chemical/endpoint combination for graphing
HQ_averages <- BTEXSN_exceedances_with_HQs %>%
  group_by(endpoint, pollutant_name) %>%
  summarise(
    avg_HQ_tot   = mean(HQ_tot, na.rm = TRUE),
    avg_HQ_pt    = mean(HQ_pt, na.rm = TRUE),
    avg_HQ_other = mean(HQ_other, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

HQ_averages <- HQ_averages %>%
  dplyr::mutate(
    pollutant_name = dplyr::recode(
      pollutant_name,
      "BENZENE" = "Benzene",
      "ETHYLBENZENE" = "Ethylbenzene",
      "STYRENE" = "Styrene",
      "TOLUENE" = "Toluene",
      "XYLENES (MIXED ISOMERS)" = "Xylenes"
    ),
    pollutant_name = factor(
      pollutant_name,
      levels = c(
        "Benzene",
        "Ethylbenzene",
        "Styrene",
        "Toluene",
        "Xylenes"
      )
    )
  )

write.csv(HQ_averages, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI1/HQ_averages.csv")

# stacked bar plot for HQ_tot
BTEXSN_HQtot_bar <- ggplot(
  HQ_averages,
  aes(x = endpoint, y = avg_HQ_tot, fill = pollutant_name)
) +
  geom_bar(stat = "identity") +
  geom_hline(
    yintercept = 1,
    linetype = "dotted",
    linewidth = 0.8
  ) +
  scale_fill_okabe_ito() +
  labs(
    x = NULL,
    y = "Average Hazard Quotient (HQ)",
    fill = "Chemical"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

BTEXSN_HQtot_bar

ggsave(
  "bar_avgHQtot_endpoints.png",
  BTEXSN_HQtot_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI1/",
  width = 7,
  height = 5,
  dpi = 600
)

# stacked bar plot for HQ_pt
BTEXSN_HQpt_bar <- ggplot(
  HQ_averages,
  aes(x = endpoint, y = avg_HQ_pt, fill = pollutant_name)
) +
  geom_bar(stat = "identity") +
  geom_hline(
    yintercept = 1,
    linetype = "dotted",
    linewidth = 0.8
  ) +
  scale_fill_okabe_ito() +
  labs(
    x = NULL,
    y = "Average Hazard Quotient (HQ)",
    fill = "Chemical"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

BTEXSN_HQpt_bar

ggsave(
  "bar_avgHQpt_endpoints.png",
  BTEXSN_HQpt_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI1/",
  width = 7,
  height = 5,
  dpi = 600
)

#for downstream paneling 
BTEXSN_HQpt_bar_HI1  <- BTEXSN_HQpt_bar
BTEXSN_HQtot_bar_HI1 <- BTEXSN_HQtot_bar
```


#population counts for exceedances
```{r}
# Get total population of all assessed ATS blocks (filtered)
SQL <- "
SELECT
  SUM(population) AS total_assessed_population
FROM (
  SELECT DISTINCT t.block, t.population
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.block = f.GEOID
)
"

ats_all_blocks <- dbGetQuery(con, SQL)  # total pop of filtered blocks

### Population counts for exceeding blocks (HI > 1)

# Ensure GEOID in filtblocks is character with 15 digits
filtblocks <- filtblocks %>%
  mutate(GEOID = str_pad(as.character(GEOID), width = 15, side = "left", pad = "0"))

# Filter BTEXSN_exceedances to only blocks in filtblocks
BTEXSN_exceedances_filtered <- BTEXSN_exceedances %>%
  semi_join(filtblocks, by = c("block" = "GEOID"))

# Summarize population for blocks with HI > 1 (unique blocks only)
pop_summary_by_source <- BTEXSN_exceedances_filtered %>%
  pivot_longer(
    cols = c(HI_tot, HI_pt, HI_other),
    names_to = "HI_type",
    values_to = "HI_value"
  ) %>%
  filter(HI_value > 1) %>%
  group_by(HI_type, block) %>%                               # ensure uniqueness
  summarise(pop_block = first(population), .groups = "drop") %>%
  group_by(HI_type) %>%
  summarise(
    total_pop = sum(pop_block, na.rm = TRUE),
    avg_pop   = mean(pop_block, na.rm = TRUE),
    n_blocks  = n(),   # count unique blocks per HI type
    .groups = "drop"
  )

write.csv(
  pop_summary_by_source, 
  "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI1/population_summary.csv",
  row.names = FALSE
)

```



#######################################

### analysis with HI threshold of 0.1 

#######################################


# Count of blocks with various HIs above 0.1 for further analysis and HQs for those blocks
```{r}
SQL <- "
WITH filtered_HIs AS (
  SELECT *
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.block = f.GEOID
)

SELECT
  'HI_tot' AS HI_type,
  COUNT(DISTINCT CASE WHEN HI_tot > 0.1 THEN block END) AS unique_blocks_above_01,
  COUNT(DISTINCT block) AS total_unique_blocks,
  COUNT(CASE WHEN HI_tot > 0.1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs

UNION ALL

SELECT
  'HI_pt',
  COUNT(DISTINCT CASE WHEN HI_pt > 0.1 THEN block END),
  COUNT(DISTINCT block),
  COUNT(CASE WHEN HI_pt > 0.1 THEN 1 END),
  COUNT(*)
FROM filtered_HIs

UNION ALL

SELECT
  'HI_other',
  COUNT(DISTINCT CASE WHEN HI_other > 0.1 THEN block END),
  COUNT(DISTINCT block),
  COUNT(CASE WHEN HI_other > 0.1 THEN 1 END),
  COUNT(*)
FROM filtered_HIs;
"

BTEXSN_blocks_HIsabove01 <- dbGetQuery(con, SQL)

write.csv(BTEXSN_blocks_HIsabove01, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/blocks_HIsabove01_summary.csv",
          row.names = FALSE)

```


# Pull df of HIs > 0.1
```{r}
# --- pull blocks with HIs > 0.1 filtered by filtblocks_BTEXSN ---
SQL = "
SELECT t.*
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON t.block = f.GEOID
WHERE HI_tot > 0.1 OR HI_pt > 0.1 OR HI_other > 0.1;
"

BTEXSN_exceedances = dbGetQuery(con, SQL)

# --- pull HQ data for those blocks by joining to filtblocks in BQ ---
SQL = "
SELECT hq.*
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HQs` hq
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON hq.block = f.GEOID
JOIN (
  SELECT block, endpoint
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs`
  WHERE HI_tot > 0.1 OR HI_pt > 0.1 OR HI_other > 0.1
) t
  ON hq.block = t.block AND hq.endpoint = t.endpoint;
"

BTEXSN_HQs_HIexceedances = dbGetQuery(con, SQL)

```


# Blocks exceeding more than one endpoint
```{r}
blocks_mult_endpoints <- BTEXSN_exceedances %>%
  group_by(block) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(unique(endpoint), collapse = ", ")
  ) %>%
  filter(n_endpoints_exceeding > 1)  

BTEXSN_mult_endpoints <- BTEXSN_exceedances %>%
  semi_join(blocks_mult_endpoints, by = "block") %>%
  left_join(blocks_mult_endpoints, by = "block")

write.csv(blocks_mult_endpoints, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/blocks_mult_endpoints.csv")

```


# Chemicals contributing to HI exceedances
```{r}
# join HQ and HI datasets together
BTEXSN_exceedances_with_HQs <- BTEXSN_exceedances %>%
  left_join(
    BTEXSN_HQs_HIexceedances %>%
      select(block, endpoint, HQ_tot, pollutant_name),
    by = c("block", "endpoint")
  )

# keep only block/endpoint combinations where:
#   HI_tot > 0.1
#   AND every individual HQ_tot <= 0.1
BTEXSN_exceedances_HI_only <- BTEXSN_exceedances_with_HQs %>%
  group_by(block, endpoint) %>%
  filter(
    any(HI_tot > 0.1) &
      all(is.na(HQ_tot) | HQ_tot <= 0.1)
  ) %>%
  mutate(
    max_HI = max(HI_tot, na.rm = TRUE)
  ) %>%
  ungroup()

# select top 10 blocks per endpoint by HI_tot
top10_blocks_per_endpoint <- BTEXSN_exceedances_HI_only %>%
  group_by(endpoint, block) %>%
  summarise(
    max_HI_block = max(max_HI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(endpoint) %>%
  arrange(desc(max_HI_block), .by_group = TRUE) %>%
  slice_head(n = 10) %>%
  ungroup()

# keep only the selected endpoint-block combinations
BTEXSN_exceedances_HI_only_top10 <- BTEXSN_exceedances_HI_only %>%
  inner_join(
    top10_blocks_per_endpoint %>%
      select(endpoint, block),
    by = c("endpoint", "block")
  )

# create generic block labels within each endpoint
BTEXSN_exceedances_HI_only_top10 <- BTEXSN_exceedances_HI_only_top10 %>%
  group_by(endpoint) %>%
  mutate(block_label = paste0("Block ", dense_rank(desc(max_HI)))) %>%
  ungroup()

# clean chemical naming for legend
BTEXSN_exceedances_HI_only_top10 <- BTEXSN_exceedances_HI_only_top10 %>%
  mutate(
    pollutant_name = dplyr::case_match(
      pollutant_name,
      "BENZENE" ~ "Benzene",
      "ETHYLBENZENE" ~ "Ethylbenzene",
      "STYRENE" ~ "Styrene",
      "TOLUENE" ~ "Toluene",
      "XYLENES (MIXED ISOMERS)" ~ "Xylenes",
      "NAPHTHALENE" ~ "Naphthalene",
      .default = pollutant_name
    ),
    pollutant_name = factor(
      pollutant_name,
      levels = c(
        "Benzene",
        "Ethylbenzene",
        "Styrene",
        "Toluene",
        "Xylenes",
        "Naphthalene"
      )
    )
  )

# graph
BTEXSN_HQblocks_bar <- ggplot(
  BTEXSN_exceedances_HI_only_top10,
  aes(x = block_label, y = HQ_tot, fill = pollutant_name)
) +
  geom_bar(stat = "identity", color = NA) +
  facet_wrap(~ endpoint, scales = "free_x") +
  scale_fill_manual(values = ggokabeito::palette_okabe_ito()) +
  labs(
    x = "Blocks",
    y = "Hazard Quotient (HQ)",
    fill = "Chemical"
  ) +
    geom_hline(
    yintercept = 0.1,
    linetype = "dotted",
    linewidth = 0.8
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(size = 10, face = "bold")
  )

BTEXSN_HQblocks_bar

ggsave(
  filename = "bar_HQtot_blocks.png",
  plot = BTEXSN_HQblocks_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/",
  width = 7,
  height = 5,
  dpi = 600
)

#save ggplot object for later paneling in the BTEXSN_dw.Rmd
saveRDS(BTEXSN_HQblocks_bar, "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/plot_A.rds")


##troubleshooting (returns 0 rows, so we're good)
n_HI_exceedances = BTEXSN_exceedances_HI_only_top10 %>%
  group_by(block, endpoint) %>%
  summarise(
    max_single_HQ = max(HQ_tot, na.rm = TRUE),
    HI_tot = max(HI_tot, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(max_single_HQ > 0.1)

# all HI exceedances = 370,874
n_cumulative_only = BTEXSN_exceedances %>%
  group_by(block, endpoint) %>%
  summarise(HI_tot = max(HI_tot, na.rm = TRUE), .groups = "drop") %>%
  filter(HI_tot > 0.1) %>%
  nrow()

# HI exceedances with no individual HQ exceedance = 4,639
BTEXSN_exceedances_HI_only %>%
  distinct(block, endpoint) %>%
  nrow()

# percentage = 1.25%
100 * n_cumulative_only / n_HI_exceedances
```



#graph average HQs for each endpoint that was exceeded in any block group to see which chemicals contribute 
```{r}
# HQ averages
HQ_averages <- BTEXSN_exceedances_with_HQs %>%
  group_by(endpoint, pollutant_name) %>%
  summarise(
    avg_HQ_tot   = mean(HQ_tot, na.rm = TRUE),
    avg_HQ_pt    = mean(HQ_pt, na.rm = TRUE),
    avg_HQ_other = mean(HQ_other, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    pollutant_name = dplyr::case_match(
      pollutant_name,
      "BENZENE" ~ "Benzene",
      "ETHYLBENZENE" ~ "Ethylbenzene",
      "STYRENE" ~ "Styrene",
      "TOLUENE" ~ "Toluene",
      "XYLENES (MIXED ISOMERS)" ~ "Xylenes",
      "NAPHTHALENE" ~ "Naphthalene",
      .default = pollutant_name
    ),
    pollutant_name = factor(
      pollutant_name,
      levels = c(
        "Benzene",
        "Ethylbenzene",
        "Styrene",
        "Toluene",
        "Xylenes",
        "Naphthalene"
      )
    )
  )

write.csv(HQ_averages, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/HQ_averages.csv")

# stacked bar plot for HQ_tot
BTEXSN_HQtot_bar <- ggplot(
  HQ_averages,
  aes(x = endpoint, y = avg_HQ_tot, fill = pollutant_name)
) +
  geom_bar(stat = "identity") +
  geom_hline(
    yintercept = 0.1,
    linetype = "dotted",
    linewidth = 0.8
  ) +
  scale_fill_okabe_ito() +
  labs(
    x = "Endpoint",
    y = "Average Hazard Quotient (HQ)",
    fill = "Chemical"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    legend.direction = "horizontal"
  )

BTEXSN_HQtot_bar

ggsave(
  filename = "bar_avgHQtot_endpoints.png",
  plot = BTEXSN_HQtot_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/",
  width = 7,
  height = 5,
  dpi = 600
)

# stacked bar plot for HQ_pt
BTEXSN_HQpt_bar <- ggplot(
  HQ_averages,
  aes(x = endpoint, y = avg_HQ_pt, fill = pollutant_name)
) +
  geom_bar(stat = "identity") +
  geom_hline(
    yintercept = 0.1,
    linetype = "dotted",
    linewidth = 0.8
  ) +
  scale_fill_okabe_ito() +
  labs(
    x = "Endpoint",
    y = "Average Hazard Quotient (HQ)",
    fill = "Chemical"
  ) +
  theme_bw(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

BTEXSN_HQpt_bar

ggsave(
  filename = "bar_avgHQpt_endpoints.png",
  plot = BTEXSN_HQpt_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/",
  width = 7,
  height = 5,
  dpi = 600
)

#for downstream paneling
BTEXSN_HQpt_bar_HI01  <- BTEXSN_HQpt_bar
BTEXSN_HQtot_bar_HI01 <- BTEXSN_HQtot_bar
```


#population counts for exceedances
```{r}
# Get total population of all assessed ATS blocks (filtered)
SQL <- "
SELECT
  SUM(population) AS total_assessed_population
FROM (
  SELECT DISTINCT t.block, t.population
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.block = f.GEOID
)
"

ats_all_blocks <- dbGetQuery(con, SQL)  # total pop of filtered blocks

### Population counts for exceeding blocks (HI > 0.1)

# Ensure GEOID in filtblocks is character with 15 digits
filtblocks <- filtblocks %>%
  mutate(GEOID = str_pad(as.character(GEOID), width = 15, side = "left", pad = "0"))

# Filter BTEXSN_exceedances to only blocks in filtblocks
BTEXSN_exceedances_filtered <- BTEXSN_exceedances %>%
  semi_join(filtblocks, by = c("block" = "GEOID"))

# Summarize population for blocks with HI > 0.1 (unique blocks only)
pop_summary_by_source <- BTEXSN_exceedances_filtered %>%
  pivot_longer(
    cols = c(HI_tot, HI_pt, HI_other),
    names_to = "HI_type",
    values_to = "HI_value"
  ) %>%
  filter(HI_value > 0.1) %>%
  group_by(HI_type, block) %>%                               # ensure uniqueness
  summarise(pop_block = first(population), .groups = "drop") %>%
  group_by(HI_type) %>%
  summarise(
    total_pop = sum(pop_block, na.rm = TRUE),
    avg_pop   = mean(pop_block, na.rm = TRUE),
    n_blocks  = n(),   # count unique blocks per HI type
    .groups = "drop"
  )

write.csv(
  pop_summary_by_source, 
  "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/population_summary.csv",
  row.names = FALSE
)

```



################## COMBINING GRAPHS FOR PUBLICATION ###########################


#4 panel figure for average HIs per endpoint colored by chemical for A) blocks exceeding HI of 1, point sources; B) blocks exceeding HI of 1; total sources; C) blocks exceeding HI of 0.1, point sources; D) blocks exceeding HI of 0.1, total sources
```{r}

#--------------------------------------------------
# Extract legend from Panel D
#--------------------------------------------------

legend <- get_legend(
  BTEXSN_HQtot_bar_HI01 +
    theme(
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.box = "horizontal"
    )
)

#--------------------------------------------------
# Remove legends from all panels
#--------------------------------------------------

pA <- BTEXSN_HQpt_bar_HI1 +
  theme(legend.position = "none")

pB <- BTEXSN_HQtot_bar_HI1 +
  theme(legend.position = "none")

pC <- BTEXSN_HQpt_bar_HI01 +
  theme(legend.position = "none")

pD <- BTEXSN_HQtot_bar_HI01 +
  theme(legend.position = "none")

#--------------------------------------------------
# Top row
#--------------------------------------------------

top_row <- plot_grid(
  pA,
  pB,
  labels = c("A", "B"),
  label_size = 18,
  label_fontface = "bold",
  nrow = 1,
  align = "hv"
)

#--------------------------------------------------
# Bottom row
#--------------------------------------------------

bottom_row <- plot_grid(
  pC,
  pD,
  labels = c("C", "D"),
  label_size = 18,
  label_fontface = "bold",
  nrow = 1,
  align = "hv"
)

#--------------------------------------------------
# Combine rows and legend
#--------------------------------------------------

final_fig <- plot_grid(
  top_row,
  bottom_row,
  legend,
  ncol = 1,
  rel_heights = c(1, 1, 0.15)
)

#--------------------------------------------------
# Save outputs
#--------------------------------------------------

outdir <- "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air"

ggsave(
  filename = file.path(outdir,
                       "avgHI_combined_thresholds_byendpoint.png"),
  plot = final_fig,
  width = 12,
  height = 10,
  dpi = 600
)

ggsave(
  filename = file.path(outdir,
                       "avgHI_combined_thresholds_byendpoint.tiff"),
  plot = final_fig,
  width = 12,
  height = 10,
  dpi = 600,
  compression = "lzw"
)

ggsave(
  filename = file.path(outdir,
                       "avgHI_combined_thresholds_byendpoint.pdf"),
  plot = final_fig,
  width = 12,
  height = 10,
  device = cairo_pdf
)
```











