---
title: "Water data analysis for BTEXSN"
author: "Paige Varner"
date: "2025-06-04"
output: html_document
---


#load libraries
```{r}
library("dplyr")
library("tidyr")
library("stringr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("car")
library("arrow")
library("progressr")
library("purrr")
library("readr")
library("haven")
library("tidycensus")
library("ggokabeito")
library("cowplot")
detach(package:Rmisc)
detach(package:plyr)

filtblocks = read.csv("~/TSCA/CRA/metdb_tsca/data/filtblocks_BTEXSN.csv")
```


#database connection
```{r}
bq_auth(email = NA)

con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "TSCA",)

```

## note: running DW analysis after creating a filtblocks_BTEXSN table in BQ during the ATS analysis phase so this table already exists, and subsequent analysis is filtering only for those blocks ##

#summary table for assuming nd is zero (low) and nd is the LOD (high) overall and by child/adult populations (note: the N in these summaries is for block + endpoint combinations)
```{r}
# --- Overall low/high summary for PWS, filtered by filtblocks_BTEXSN ---
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.geoid20 = f.GEOID
)
SELECT 
  low_up,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_HIs
GROUP BY low_up
ORDER BY low_up
"

BTEXSN_HI_summary_low_up <- dbGetQuery(con, SQL)

write.csv(
  BTEXSN_HI_summary_low_up, 
  file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HIsummary_low_up.csv",
  row.names = FALSE
)

# --- Low/high summary by population for PWS ---
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.geoid20 = f.GEOID
)
SELECT 
  low_up,
  pop,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_HIs
GROUP BY low_up, pop
ORDER BY low_up, pop
"

BTEXSN_HI_summary_low_up_pop <- dbGetQuery(con, SQL)

write.csv(
  BTEXSN_HI_summary_low_up_pop, 
  file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HIsummary_low_up_pop.csv",
  row.names = FALSE
)

```


## moving forward - analysis conducted for assuming non-detects are 1/2 the LOD; this pulls from BQ views and tables that end in _half ##


#summary tables (note: the N in these summaries is for block + endpoint combinations, except when broken out by endpoint)
```{r}
# Overall
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
"

BTEXSN_HI_summary <- dbGetQuery(con, SQL)

write.csv(BTEXSN_HI_summary, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HIsummary.csv")


# Population
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  pop,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
GROUP BY pop
ORDER BY pop
"

BTEXSN_HI_summary_pop <- dbGetQuery(con, SQL)

write.csv(BTEXSN_HI_summary_pop, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HIsummary_pop.csv")


# Endpoint
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  endpoint,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
GROUP BY endpoint
ORDER BY endpoint
"

BTEXSN_HI_summary_endpoint <- dbGetQuery(con, SQL)

write.csv(BTEXSN_HI_summary_endpoint, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HIsummary_by_endpoint.csv")


# Population and endpoint
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  pop,
  endpoint,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
GROUP BY pop, endpoint
ORDER BY pop, endpoint
"

BTEXSN_HI_summary_pop_endpoint <- dbGetQuery(con, SQL)

write.csv(BTEXSN_HI_summary_pop_endpoint, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HIsummary_pop_endpoint.csv")

```



#delta for each pollutant
```{r}
# overall deltas
SQL <- "
WITH filtered_blocks AS (
  SELECT hq.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HQs_half` hq
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON CAST(hq.geoid20 AS STRING) = f.GEOID
)
SELECT
  hq.chem,
  COUNT(*) AS n,
  AVG(hq.hq) AS avg_HQ,
  AVG(hi.HI) AS avg_HI,
  SAFE_DIVIDE(AVG(hi.HI) - AVG(hq.hq), 1) AS delta_avg
FROM filtered_blocks hq
JOIN `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` hi
  ON hq.geoid20 = hi.geoid20
     AND hq.endpoint = hi.endpoint
GROUP BY hq.chem
ORDER BY hq.chem
"

BTEXSN_chem_avg_deltas <- dbGetQuery(con, SQL)
write.csv(BTEXSN_chem_avg_deltas, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/avg_deltas.csv")


# Deltas for blocks with HI >= 1
SQL <- "
WITH filtered_blocks AS (
  SELECT hq.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HQs_half` hq
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON CAST(hq.geoid20 AS STRING) = f.GEOID
)
SELECT
  hq.chem,
  COUNT(*) AS n,
  AVG(hq.hq) AS avg_HQ,
  AVG(hi.HI) AS avg_HI,
  SAFE_DIVIDE(AVG(hi.HI) - AVG(hq.hq), 1) AS delta_avg
FROM filtered_blocks hq
JOIN `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` hi
  ON hq.geoid20 = hi.geoid20
     AND hq.endpoint = hi.endpoint
WHERE hi.HI >= 1
GROUP BY hq.chem
ORDER BY hq.chem
"

BTEXSN_chem_avg_deltas_1 <- dbGetQuery(con, SQL)
write.csv(BTEXSN_chem_avg_deltas_1, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI1/avg_deltas_HI1.csv")


# Deltas for blocks with HI >= 0.1
SQL <- "
WITH filtered_blocks AS (
  SELECT hq.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HQs_half` hq
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON CAST(hq.geoid20 AS STRING) = f.GEOID
)
SELECT
  hq.chem,
  COUNT(*) AS n,
  AVG(hq.hq) AS avg_HQ,
  AVG(hi.HI) AS avg_HI,
  SAFE_DIVIDE(AVG(hi.HI) - AVG(hq.hq), 1) AS delta_avg
FROM filtered_blocks hq
JOIN `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` hi
  ON hq.geoid20 = hi.geoid20
     AND hq.endpoint = hi.endpoint
WHERE hi.HI >= 0.1
GROUP BY hq.chem
ORDER BY hq.chem
"

BTEXSN_chem_avg_deltas_01 <- dbGetQuery(con, SQL)
write.csv(BTEXSN_chem_avg_deltas_01, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/avg_deltas_HI01.csv")

```


#CDFs for children and adults
```{r}
# --- 1. Pull the data ---
project_id <- "edf-aq-data"

SQL <- "
SELECT
  t.geoid20,
  t.endpoint,
  t.pop,
  t.HI
FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON t.geoid20 = f.GEOID
"

BTEXS_PWS_HIs_filtered <- bq_project_query(project_id, SQL) %>%
  bq_table_download()

# --- 2. Pre-compute CDF by population AND endpoint ---
BTEXS_cdf <- BTEXS_PWS_HIs_filtered %>%
  group_by(pop, endpoint) %>%
  arrange(HI) %>%
  mutate(cdf = row_number() / n()) %>%
  ungroup()

# Optional: thin points (every 10th to speed up plotting)
BTEXS_cdf_sample <- BTEXS_cdf %>%
  group_by(pop, endpoint) %>%
  slice(seq(1, n(), by = 10)) %>%
  ungroup()

# --- 3. Plot ---
cdf_plot <- ggplot(
  BTEXS_cdf_sample,
  aes(x = HI, y = cdf, group = endpoint, color = endpoint)
) +
  geom_line(linewidth = 1) +
  # vertical reference lines for thresholds
  geom_vline(xintercept = c(0.1, 1), linetype = "dotted", color = "black") +
  scale_x_log10() +
  labs(
    title = "CDF of PWS HI (BTEXS, Children vs Adults)",
    x = "HI value (log scale)",
    y = "P(HI ≤ x)",
    color = "Endpoint"
  ) +
  facet_wrap(~pop, scales = "free_y") +
  theme_minimal(base_size = 14)

cdf_plot

ggsave(
  filename = "BTEXSN_PWS_cdf.png",
  device = "png",
  plot = cdf_plot,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/",
  width = 7,
  height = 5
)

```



#######################################

### analysis with HI threshold of 0.1 (no blocks above 1)

#######################################



# Count of blocks with HIs above 0.1 for further analysis and HQs for those blocks
```{r}
#overall
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON CAST(t.geoid20 AS STRING) = f.GEOID
)

SELECT
  'HI' AS HI_type,
  COUNT(DISTINCT CASE WHEN HI > 0.1 THEN geoid20 END) AS unique_blocks_above_01,
  COUNT(DISTINCT geoid20) AS total_unique_blocks,
  COUNT(CASE WHEN HI > 0.1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs;
"

BTEXS_blocks_HIsabove01 <- dbGetQuery(con, SQL)

write.csv(BTEXS_blocks_HIsabove01, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/blocks_HIsabove01_summary.csv",
          row.names = FALSE)


#for each population
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
    ON CAST(t.geoid20 AS STRING) = f.GEOID
)
SELECT
  pop,
  COUNT(DISTINCT CASE WHEN HI > 0.1 THEN geoid20 END) AS unique_blocks_above_01,
  COUNT(DISTINCT geoid20) AS total_unique_blocks,
  COUNT(CASE WHEN HI > 0.1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs
GROUP BY pop
ORDER BY pop;
"

BTEXS_blocks_HIsabove01_by_pop <- dbGetQuery(con, SQL)

write.csv(BTEXS_blocks_HIsabove01_by_pop, 
          file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/blocks_HIsabove01_bypop_summary.csv",
          row.names = FALSE)
```


# Pull HI and HQ dfs of HIs > 0.1
```{r}
# --- pull blocks with HI > 0.1 filtered by filtblocks_BTEXSN ---
SQL_hi <- "
SELECT t.*
FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half` t
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON CAST(t.geoid20 AS STRING) = f.GEOID
WHERE HI > 0.1
"

BTEXS_exceedances <- dbGetQuery(con, SQL_hi)

# --- pull HQ data for those blocks and endpoints ---
SQL_hq <- "
SELECT hq.*
FROM `edf-aq-data.TSCA.BTEXS_PWS_HQs_half` hq
JOIN `edf-aq-data.TSCA.filtblocks_BTEXSN` f
  ON CAST(hq.geoid20 AS STRING) = f.GEOID
JOIN (
  SELECT geoid20, endpoint
  FROM `edf-aq-data.TSCA.BTEXS_PWS_HIs_half`
  WHERE HI > 0.1
) t
  ON hq.geoid20 = t.geoid20 AND hq.endpoint = t.endpoint
"

BTEXS_HQs_HIexceedances <- dbGetQuery(con, SQL_hq)

```



# Blocks exceeding more than one endpoint
```{r}
#overall
blocks_mult_endpoints <- BTEXS_exceedances %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

BTEXS_mult_endpoints <- BTEXS_exceedances %>%
  semi_join(blocks_mult_endpoints, by = "geoid20") %>%
  left_join(blocks_mult_endpoints, by = "geoid20")

write.csv(blocks_mult_endpoints, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/blocks_mult_endpoints.csv", row.names = FALSE)

#adult (no blocks exceed multiple endpoints in adult populations)
blocks_mult_endpoints_ad <- BTEXS_exceedances %>%
  filter(pop == "ad") %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

BTEXS_mult_endpoints_ad <- BTEXS_exceedances %>%
  filter(pop == "ad") %>%
  semi_join(blocks_mult_endpoints_ad, by = "geoid20") %>%
  left_join(blocks_mult_endpoints_ad, by = "geoid20")

write.csv(blocks_mult_endpoints_ad, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/blocks_mult_endpoints_ad.csv", row.names = FALSE)


#child
blocks_mult_endpoints_ch <- BTEXS_exceedances %>%
  filter(pop == "ch") %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

BTEXS_mult_endpoints_ch <- BTEXS_exceedances %>%
  filter(pop == "ch") %>%
  semi_join(blocks_mult_endpoints, by = "geoid20") %>%
  left_join(blocks_mult_endpoints, by = "geoid20")

write.csv(blocks_mult_endpoints, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/blocks_mult_endpoints_ch.csv", row.names = FALSE)

```


#graph average HQs for each endpoint that was exceeded in any block group to see which chemicals contribute (don't think we need to present this figure for now, as the data will be provided in tables; however, if we want it, will need to clean it)
```{r}
# Join HI and HQ datasets together 
BTEXS_exceedances_with_HQs <- BTEXS_exceedances %>%
  left_join(
    BTEXS_HQs_HIexceedances %>%
      select(geoid20, endpoint, pop, hq, chem),
    by = c("geoid20", "endpoint", "pop")
  )

# HQ averages by endpoint, chemical, and population
HQ_averages <- BTEXS_exceedances_with_HQs %>%
  group_by(endpoint, chem, pop) %>%
  summarise(
    n = sum(!is.na(hq)),
    avg_HQ = mean(hq, na.rm = TRUE),
    .groups = "drop"
  )

write.csv(HQ_averages, file = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/HQ_averages.csv")

# stacked bar plot faceted by pop
HQ_averages <- HQ_averages %>%
  mutate(pop = dplyr::recode(pop,
                             "ad" = "Adult",
                             "ch" = "Child"))

BTEXS_HQtot_bar <- ggplot(HQ_averages, aes(x = endpoint, y = avg_HQ, fill = chem)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average HQ by Endpoint, Pollutant, and Population Group",
    x = "Endpoint",
    y = "Average HQ",
    fill = "Pollutant"
  ) +
  facet_wrap(~ pop) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

BTEXS_HQtot_bar

ggsave(
  filename = "bar_avgHQ_endpoints_by_pop.png",
  device = "png",
  plot = BTEXS_HQtot_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/",
  width = 7,
  height = 5
)

```


# Chemicals contributing to HI exceedances for adults - none where there's an HI above 0.1 where an HQ wasn't already above 0.1, so no graph here
```{r}

# Compute max HQ per block/endpoint and filter blocks where HI > 0.1 but HQ_max ≤ 0.1
blocks_HI_only <- BTEXS_exceedances_with_HQs %>%
  group_by(geoid20, endpoint) %>%
  mutate(max_HQ_block = max(hq, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(pop == "ad" & HI > 0.1 & (is.na(max_HQ_block) | max_HQ_block <= 0.1))

```


# Chemicals contributing to HI exceedances for children
```{r}
# 1. Join HI and HQ datasets (child population only)
BTEXS_exceedances_with_HQs <- BTEXS_exceedances %>%
  filter(pop == "ch") %>%
  left_join(
    BTEXS_HQs_HIexceedances %>%
      filter(pop == "ch") %>%
      select(geoid20, endpoint, hq, chem),
    by = c("geoid20", "endpoint")
  )

# 2. Compute max HI and max HQ per block/endpoint
blocks_HI_only <- BTEXS_exceedances_with_HQs %>%
  group_by(geoid20, endpoint) %>%
  summarise(
    HI_block = max(HI, na.rm = TRUE),
    HQ_block_max = max(hq, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(HI_block > 0.1 & (is.na(HQ_block_max) | HQ_block_max <= 0.1))

# 3. Select top 10 blocks per endpoint by HI
top10_blocks_per_endpoint <- blocks_HI_only %>%
  group_by(endpoint) %>%
  arrange(desc(HI_block)) %>%
  slice_head(n = 10) %>%
  ungroup()

# 4. Keep HQ rows only for the top 10 blocks per endpoint
BTEXS_exceedances_HI_only_top10 <- BTEXS_exceedances_with_HQs %>%
  inner_join(
    top10_blocks_per_endpoint %>%
      select(geoid20, endpoint),
    by = c("geoid20", "endpoint")
  )

BTEXS_exceedances_HI_only_top10 <- BTEXS_exceedances_HI_only_top10 %>%
  mutate(
    chem = dplyr::recode(
      chem,
      benzene = "Benzene",
      ethylbenzene = "Ethylbenzene",
      styrene = "Styrene",
      toluene = "Toluene",
      xylenes = "Xylenes"
    )
  )

# 5. Add block-level HI ranking (use HI_block from selection table)
block_labels <- top10_blocks_per_endpoint %>%
  group_by(endpoint) %>%
  arrange(desc(HI_block), geoid20) %>%
  mutate(block_label = paste0("Block ", row_number())) %>%
  ungroup() %>%
  select(geoid20, endpoint, block_label)

# Join labels back to plotting dataset
BTEXS_exceedances_HI_only_top10 <- BTEXS_exceedances_HI_only_top10 %>%
  left_join(block_labels, by = c("geoid20", "endpoint"))


# 6. Plot ONLY the top-10 dataset
BTEXS_HQblocks_bar <- ggplot(
  BTEXS_exceedances_HI_only_top10,
  aes(x = block_label, y = hq, fill = chem)
) +
  geom_bar(stat = "identity") +
  geom_hline(
    yintercept = 0.1,
    linetype = "dotted",
    linewidth = 0.8
  ) +
  facet_wrap(~ endpoint, scales = "free_x") +
  scale_fill_okabe_ito() +
  labs(
    x = "Blocks",
    y = "Hazard Quotient (HQ)",
    fill = "Chemical"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(size = 10, face = "bold"),
    legend.position = "right"
  )

BTEXS_HQblocks_bar

ggsave(
  filename = "bar_HQ_blocks_child.png",
  plot = BTEXS_HQblocks_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/",
  width = 7,
  height = 5,
  dpi = 600
)

#save ggplot object for later paneling in the BTEXSN_dw.Rmd
saveRDS(BTEXS_HQblocks_bar, "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/plot_B.rds")

# Total child HI exceedances = 1,827
n_HI_exceedances_child <- BTEXS_exceedances_with_HQs %>%
  group_by(geoid20, endpoint) %>%
  summarise(
    HI_block = max(HI, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(HI_block > 0.1) %>%
  nrow()

# Child HI exceedances where no individual HQ exceeds 0.1 = 29
n_cumulative_only_child <- BTEXS_exceedances_with_HQs %>%
  group_by(geoid20, endpoint) %>%
  summarise(
    HI_block = max(HI, na.rm = TRUE),
    HQ_block_max = max(hq, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(
    HI_block > 0.1 &
      (is.na(HQ_block_max) | HQ_block_max <= 0.1)
  ) %>%
  nrow()

# Percent attributable solely to cumulative effects = 1.59%
pct_cumulative_only_child <-
  100 * n_cumulative_only_child / n_HI_exceedances_child

# Display results
n_HI_exceedances_child
n_cumulative_only_child
pct_cumulative_only_child
```


#population counts for exceedances 
```{r}
# Step 1: Get distinct state-county combinations from BTEXS_exceedances (already filtered for blocks of interest)
state_county_pairs <- BTEXS_exceedances %>%
  mutate(
    statefp = substr(geoid20, 1, 2),
    countyfp = substr(geoid20, 3, 5)
  ) %>%
  distinct(statefp, countyfp)

# Step 2: Function to get block-level population for a state-county
get_block_pop <- function(statefp, countyfp) {
  tryCatch({
    get_decennial(
      geography = "block",
      variables = "P1_001N",
      year = 2020,
      sumfile = "pl",
      state = statefp,
      county = countyfp,
      geometry = FALSE,
      output = "wide"
    ) %>%
      rename(census_population = P1_001N)
  }, error = function(e) {
    message(paste("Error in", statefp, countyfp, ":", e$message))
    return(NULL)
  })
}

# Step 3: Loop through all state-county combinations and combine results
census_pop_all <- map2_dfr(state_county_pairs$statefp, state_county_pairs$countyfp, get_block_pop)

# Step 4: Join with exceedances
BTEXS_with_pop <- BTEXS_exceedances %>%
  left_join(census_pop_all, by = c("geoid20" = "GEOID"))

# Step 5: Filter HI > 0.1 and summarise population for unique blocks
pop_summary_blocks <- BTEXS_with_pop %>%
  filter(HI > 0.1) %>%
  group_by(pop, geoid20) %>%   # ensure unique block-level counts
  summarise(block_pop = first(census_population), .groups = "drop") %>%
  group_by(pop) %>%
  summarise(
    total_census_population = sum(block_pop, na.rm = TRUE),
    avg_census_population   = mean(block_pop, na.rm = TRUE),
    n_unique_blocks         = n(),
    .groups = "drop"
  )

# Step 6: Write output to CSV
write_csv(
  pop_summary_blocks,
  "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/population_summary.csv"
)
```



######### COMBINING PLOTS FOR PUBLICATION ###########


#figure of HQs of each chemical for relevant endpoints in example blocks for A) air (all sources), and B) water (children) to show how considering more than one chemical pushes thresholds compared to just one chemical
```{r}
pA <- readRDS("~/TSCA/CRA/metdb_tsca/output/BTEXSN/Air/HI01/plot_A.rds")
pB <- readRDS("~/TSCA/CRA/metdb_tsca/output/BTEXSN/Water/HI01/plot_B.rds")

#remove legends and extract from only plot A
pA <- pA + theme(legend.position = "none")
pB <- pB + theme(legend.position = "none")

legend <- get_legend(
  pA +
    theme(
      legend.position = "bottom",
      legend.direction = "horizontal"
    )
)

#combine 
top_row <- plot_grid(
  pA,
  pB,
  labels = c("A", "B"),
  nrow = 1,
  align = "hv"
)

#add shared legend
final_fig <- plot_grid(
  top_row,
  legend,
  ncol = 1,
  rel_heights = c(1, 0.15)
)

#save
ggsave(
  filename = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Combined/contributions_blocks_combined.png",
  plot = final_fig,
  width = 12,
  height = 6,
  dpi = 600
)

ggsave(
  filename = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Combined/contributions_blocks_combined.tiff",
  plot = final_fig,
  width = 12,
  height = 6,
  dpi = 600,
  compression = "lzw"
)

ggsave(
  filename = "~/TSCA/CRA/metdb_tsca/output/BTEXSN/Combined/contributions_blocks_combined.pdf",
  plot = final_fig,
  width = 12,
  height = 6,
  device = cairo_pdf
)
```






