---
title: "BTEXSN Visualization for MS publication - paneled plots"
output: html_document
date: "2026-07-01"
---


# Load libraries
```{r}
library(bigrquery)
library(DBI)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
library(stringr)
```


# Connect to BigQuery
```{r}
bq_auth(
  email = NA
)

con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "airtoxscreen"
)

project_id <- "edf-aq-data"
```

#important sensitivity analysis on combination statistics: 15,175 blocks are included in the PWS dataset but not the ATS dataset. need to check the population from the Census data on these blocks to see if they weren't included in ATS because ATS doesn't model exposures for blocks with population = 0. It makes sense otherwise that there are lots of blocks that have ATS data and not PWS data because millions are not on PWS.
```{r}
# Step 1: Query unmatched PWS blocks from BigQuery
SQL <- "
SELECT DISTINCT
  LPAD(CAST(pws.geoid20 AS STRING), 15, '0') AS geoid20
FROM `edf-aq-data.TSCA.BTEXS_PWS_HQs_half` AS pws
WHERE LPAD(CAST(pws.geoid20 AS STRING), 15, '0') NOT IN (
  SELECT DISTINCT LPAD(CAST(block AS STRING), 15, '0')
  FROM `edf-aq-data.TSCA.BTEXSN_ATS_HIs`
)
"
unmatched_PWS <- dbGetQuery(con, SQL)

# Step 2: Get state FIPS codes from unmatched blocks
unmatched_PWS <- unmatched_PWS %>%
  mutate(state_fips = substr(geoid20, 1, 2))

states_needed <- unique(unmatched_PWS$state_fips)

# Step 4: Loop through those states to pull 2020 block-level population
pop_blocks_all <- data.frame()

for (fips in states_needed) {
  message("Getting population for state FIPS: ", fips)
  
  tryCatch({
    state_blocks <- get_decennial(
      geography = "block",
      variables = "P1_001N",  # total population 2020 PL data
      year = 2020,
      state = fips,
      geometry = FALSE,
      output = "wide",
      dataset = "pl"
    )
    
    pop_blocks_all <- bind_rows(pop_blocks_all, state_blocks)
  }, error = function(e) {
    message("Error for state ", fips, ": ", e$message)
  })
}


# Step 5: Join population to unmatched PWS blocks
pws_blocks_with_pop <- unmatched_PWS %>%
  left_join(pop_blocks_all, by = c("geoid20" = "GEOID"))


# Step 6: Summarize
summary_table <- pws_blocks_with_pop %>%
  mutate(pop = P1_001N) %>%
  summarise(
    total_blocks = n(),
    blocks_with_population = sum(!is.na(pop) & pop > 0),
    blocks_with_zero_or_na = sum(is.na(pop) | pop == 0),
    total_population = sum(pop, na.rm = TRUE)
  )

# Step 7: View results
print(summary_table)

#all 15,175 blocks included in PWS and not in ATS have a population of zero or NA in the 2020 Census - this is great news, which means that the joining of the ATS and PWS HIs are joined properly and there are no issues with mismatching geoids, for example
```




# Define output directory
```{r}
output_dir <- "~/TSCA/CRA/metdb_tsca/output/BTEXSN/"

dir.create(
  output_dir,
  recursive = TRUE,
  showWarnings = FALSE
)
```


# Define endpoint and scenario settings
```{r}
selected_endpoints <- c(
  "hematological",
  "reproductive"
)

endpoint_levels <- c(
  "Hematological",
  "Reproductive"
)

four_scenario_levels <- c(
  "Point air",
  "Total air",
  "Total air + child drinking water",
  "Cumulative HI"
)

four_scenario_labels <- c(
  "Point air" =
    "Point air",

  "Total air" =
    "Total air",

  "Total air + child drinking water" =
    "Total air + Ingestion (Child)",

  "Cumulative HI" =
    "Cumulative HI (Child)"
)
```


# Distribution-summary function
```{r}
summarise_distribution <- function(
  data,
  value,
  ...
) {

  data %>%
    group_by(...) %>%
    summarise(
      sample_size = sum(
        !is.na({{ value }})
      ),

      mean = mean(
        {{ value }},
        na.rm = TRUE
      ),

      median = median(
        {{ value }},
        na.rm = TRUE
      ),

      minimum = min(
        {{ value }},
        na.rm = TRUE
      ),

      maximum = max(
        {{ value }},
        na.rm = TRUE
      ),

      sd = sd(
        {{ value }},
        na.rm = TRUE
      ),

      se = sd(
        {{ value }},
        na.rm = TRUE
      ) /
        sqrt(
          sum(
            !is.na({{ value }})
          )
        ),

      .groups = "drop"
    )
}
```


#pull benzene HQ data
```{r}
# ============================================================
# 1. Pull Benzene air HQs
#
# SELECT DISTINCT GEOID prevents duplicate rows in the
# filtering table from multiplying the HQ records.
# ============================================================

SQL_benzene_air_hq <- "
SELECT
  h.block AS geoid20,
  h.pollutant_name,
  h.endpoint,
  h.HQ_pt,
  h.HQ_tot
FROM `edf-aq-data.TSCA.BTEXSN_ATS_HQs` h
JOIN (
  SELECT DISTINCT
    GEOID
  FROM `edf-aq-data.TSCA.filtblocks_BTEXSN`
) f
  ON h.block = f.GEOID
WHERE UPPER(h.pollutant_name) = 'BENZENE'
  AND LOWER(h.endpoint) IN (
    'hematological',
    'reproductive'
  )
"


benzene_air_hq <- bq_project_query(
  project_id,
  SQL_benzene_air_hq
) %>%
  bq_table_download() %>%
  transmute(
    geoid20 = as.character(
      geoid20
    ),

    endpoint = str_to_lower(
      str_trim(endpoint)
    ),

    HQ_pt = as.numeric(
      HQ_pt
    ),

    HQ_tot = as.numeric(
      HQ_tot
    )
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    endpoint %in% selected_endpoints
  )


# ============================================================
# 2. Check Benzene air HQ uniqueness
#
# Expected unique unit:
# geoid20 + endpoint
# ============================================================

benzene_air_hq_duplicate_check <-
  benzene_air_hq %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

benzene_air_hq_duplicate_check


# ============================================================
# 3. Pull Benzene child drinking-water HQs
# ============================================================

SQL_benzene_water_hq <- "
SELECT
  p.geoid20,
  p.endpoint,
  p.pop,
  p.chem,
  p.hq AS HQ_water
FROM `edf-aq-data.TSCA.BTEXS_PWS_HQs_half` p
JOIN (
  SELECT DISTINCT
    GEOID
  FROM `edf-aq-data.TSCA.filtblocks_BTEXSN`
) f
  ON p.geoid20 = f.GEOID
WHERE LOWER(p.chem) = 'benzene'
  AND p.pop = 'ch'
  AND LOWER(p.endpoint) IN (
    'hematological',
    'reproductive'
  )
"


benzene_water_hq <- bq_project_query(
  project_id,
  SQL_benzene_water_hq
) %>%
  bq_table_download() %>%
  transmute(
    geoid20 = as.character(
      geoid20
    ),

    endpoint = str_to_lower(
      str_trim(endpoint)
    ),

    HQ_water = as.numeric(
      HQ_water
    )
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    endpoint %in% selected_endpoints
  )


# ============================================================
# 4. Check Benzene water HQ uniqueness
#
# Expected unique unit:
# geoid20 + endpoint
# ============================================================

benzene_water_hq_duplicate_check <-
  benzene_water_hq %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

benzene_water_hq_duplicate_check


# ============================================================
# 5. Create Benzene HQ data for the three HQ scenarios
#
# The inner join means all three HQ scenarios use the same
# block-endpoint population: blocks with both air and child
# drinking-water HQ data.
# ============================================================

benzene_combined_hq <- benzene_air_hq %>%
  inner_join(
    benzene_water_hq,
    by = c(
      "geoid20",
      "endpoint"
    )
  ) %>%
  mutate(
    HQ_combined =
      HQ_tot + HQ_water
  )


# ============================================================
# 6. Check final combined HQ uniqueness
# ============================================================

benzene_combined_hq_duplicate_check <-
  benzene_combined_hq %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

benzene_combined_hq_duplicate_check


# ============================================================
# 7. Convert Benzene HQ data to three-scenario long format
# ============================================================

benzene_hq_scenario_df <- benzene_combined_hq %>%
  transmute(
    geoid20,
    endpoint,

    `Point air` =
      HQ_pt,

    `Total air` =
      HQ_tot,

    `Total air + child drinking water` =
      HQ_combined
  ) %>%
  pivot_longer(
    cols = c(
      `Point air`,
      `Total air`,
      `Total air + child drinking water`
    ),

    names_to = "scenario",
    values_to = "value"
  ) %>%
  filter(
    !is.na(value)
  ) %>%
  mutate(
    metric = "Benzene HQ"
  )


# ============================================================
# 8. Check resulting scenario data
# ============================================================

benzene_hq_scenario_df %>%
  count(
    endpoint,
    scenario,
    name = "n_records"
  )


benzene_hq_scenario_df %>%
  group_by(
    endpoint,
    scenario
  ) %>%
  summarise(
    total_blocks = n_distinct(
      geoid20
    ),
    .groups = "drop"
  )
```



# Pull cumulative HIs
```{r}
# ============================================================
# Pull cumulative HIs only for the two publication endpoints
#
# Population priority is handled in BigQuery:
# 1. Use child ("ch") HI when available
# 2. Otherwise retain population-unspecified ("NA") HI
#
# Only one row per block-endpoint is downloaded to R.
# ============================================================

SQL_selected_endpoint_hi <- "
WITH eligible_hi AS (
  SELECT
    h.geoid20,
    LOWER(TRIM(h.endpoint)) AS endpoint,
    LOWER(TRIM(h.pop)) AS pop,
    h.HI_tot_comb,

    ROW_NUMBER() OVER (
      PARTITION BY
        h.geoid20,
        LOWER(TRIM(h.endpoint))

      ORDER BY
        CASE
          WHEN LOWER(TRIM(h.pop)) = 'ch' THEN 1
          WHEN LOWER(TRIM(h.pop)) = 'na' THEN 2
          ELSE 3
        END
    ) AS population_rank

  FROM
    `edf-aq-data.TSCA.BTEXSN_combined_HIs` h

  JOIN (
    SELECT DISTINCT
      GEOID
    FROM
      `edf-aq-data.TSCA.filtblocks_BTEXSN`
  ) f
    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'na'
    )

    AND LOWER(TRIM(h.endpoint)) IN (
      'hematological',
      'reproductive'
    )

    AND h.HI_tot_comb IS NOT NULL
)

SELECT
  geoid20,
  endpoint,
  pop AS selected_population,
  HI_tot_comb

FROM
  eligible_hi

WHERE
  population_rank = 1
"


selected_endpoint_hi_raw <- bq_project_query(
  project_id,
  SQL_selected_endpoint_hi
) %>%
  bq_table_download() %>%
  transmute(
    geoid20 = as.character(
      geoid20
    ),

    endpoint = str_to_lower(
      str_trim(endpoint)
    ),

    selected_population = str_to_lower(
      str_trim(selected_population)
    ),

    HI_tot_comb = as.numeric(
      HI_tot_comb
    )
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    endpoint %in% selected_endpoints,
    !is.na(HI_tot_comb)
  )


# ============================================================
# Check that one HI was returned per block-endpoint
#
# This should return zero rows.
# ============================================================

selected_endpoint_hi_duplicate_check <-
  selected_endpoint_hi_raw %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

selected_endpoint_hi_duplicate_check


# ============================================================
# Report how many selected records came from each population
# ============================================================

selected_endpoint_hi_population_summary <-
  selected_endpoint_hi_raw %>%
  count(
    selected_population,
    name = "selected_block_endpoint_records"
  )

selected_endpoint_hi_population_summary


# ============================================================
# Check block counts by endpoint
# ============================================================

selected_endpoint_hi_block_summary <-
  selected_endpoint_hi_raw %>%
  group_by(
    endpoint
  ) %>%
  summarise(
    total_records = n(),

    total_blocks = n_distinct(
      geoid20
    ),

    .groups = "drop"
  )

selected_endpoint_hi_block_summary


# ============================================================
# Prepare cumulative HI records for the final figure
# ============================================================

selected_endpoint_hi_df <- selected_endpoint_hi_raw %>%
  transmute(
    geoid20,
    endpoint,
    scenario = "Cumulative HI",
    metric = "Cumulative HI",
    value = HI_tot_comb
  )
```



#prepare data for figure
```{r}
# ============================================================
# Prepare selected cumulative HI records for the two endpoints
#
# selected_endpoint_hi_raw already contains:
# - only hematological and reproductive endpoints
# - one selected HI per block-endpoint
# - child HI preferred over NA-population HI
# ============================================================

selected_endpoint_hi_df <- selected_endpoint_hi_raw %>%
  transmute(
    geoid20,
    endpoint,
    scenario = "Cumulative HI",
    metric = "Cumulative HI",
    value = HI_tot_comb
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    !is.na(value)
  )


# ============================================================
# Combine HQ and HI records for the final analysis dataset
#
# This dataset retains zero values and is used for:
# - the distribution-summary table
# - the block threshold table
#
# A separate positive-value-only dataset is created below
# for the log-scale figure.
# ============================================================

combined_endpoint_analysis_df <- bind_rows(
  benzene_hq_scenario_df,
  selected_endpoint_hi_df
) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    !is.na(scenario),
    !is.na(metric),
    !is.na(value)
  ) %>%
  mutate(
    endpoint = str_to_title(
      endpoint
    ),

    endpoint = factor(
      endpoint,
      levels = endpoint_levels
    ),

    scenario = factor(
      scenario,
      levels = four_scenario_levels
    ),

    metric = factor(
      metric,
      levels = c(
        "Benzene HQ",
        "Cumulative HI"
      )
    )
  )


# ============================================================
# Restrict analysis to blocks represented in every endpoint
# and every scenario
#
# This ensures that every plotted distribution and every
# corresponding table uses the exact same Census blocks.
# ============================================================

required_endpoint_scenario_combinations <-
  length(endpoint_levels) *
  length(four_scenario_levels)


# Identify blocks with all eight endpoint-scenario combinations
common_analysis_blocks <-
  combined_endpoint_analysis_df %>%
  distinct(
    geoid20,
    endpoint,
    scenario
  ) %>%
  count(
    geoid20,
    name = "n_endpoint_scenario_combinations"
  ) %>%
  filter(
    n_endpoint_scenario_combinations ==
      required_endpoint_scenario_combinations
  ) %>%
  select(
    geoid20
  )


# Number of blocks retained
nrow(common_analysis_blocks)


# Restrict the full analysis dataset to the common blocks
combined_endpoint_analysis_df <-
  combined_endpoint_analysis_df %>%
  semi_join(
    common_analysis_blocks,
    by = "geoid20"
  )


# ============================================================
# Final analytical duplicate check
#
# Each block should occur once within each:
# endpoint + scenario
#
# This should return zero rows.
# ============================================================

combined_endpoint_duplicate_check <-
  combined_endpoint_analysis_df %>%
  count(
    geoid20,
    endpoint,
    scenario,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

combined_endpoint_duplicate_check


# ============================================================
# Check block counts by endpoint and scenario
#
# total_records and total_blocks should be identical if there
# is exactly one record per block-endpoint-scenario.
# ============================================================

combined_endpoint_sample_sizes <-
  combined_endpoint_analysis_df %>%
  group_by(
    endpoint,
    scenario,
    metric
  ) %>%
  summarise(
    total_records = n(),

    total_blocks = n_distinct(
      geoid20
    ),

    .groups = "drop"
  )

combined_endpoint_sample_sizes


# ============================================================
# Create the plot dataset
#
# Logarithmic axes cannot display zero or negative values.
#
# Zero values remain in combined_endpoint_analysis_df and are
# therefore retained in all table denominators.
# ============================================================

combined_endpoint_plot_df <-
  combined_endpoint_analysis_df %>%
  filter(
    value > 0
  )


# ============================================================
# Check the number of zero values omitted from the graph
# ============================================================

combined_endpoint_zero_summary <-
  combined_endpoint_analysis_df %>%
  group_by(
    endpoint,
    scenario
  ) %>%
  summarise(
    total_blocks = n_distinct(
      geoid20
    ),

    zero_value_blocks = n_distinct(
      geoid20[
        value == 0
      ]
    ),

    positive_value_blocks = n_distinct(
      geoid20[
        value > 0
      ]
    ),

    .groups = "drop"
  )

combined_endpoint_zero_summary


# ============================================================
# Create the figure distribution-summary table
#
# This table uses the full analysis dataset, including zeros.
# ============================================================

combined_endpoint_distribution_summary <-
  summarise_distribution(
    data = combined_endpoint_analysis_df,
    value = value,
    endpoint,
    scenario,
    metric
  ) %>%
  mutate(
    endpoint = as.character(
      endpoint
    ),

    scenario = as.character(
      scenario
    ),

    metric = as.character(
      metric
    )
  ) %>%
  arrange(
    factor(
      endpoint,
      levels = endpoint_levels
    ),

    factor(
      scenario,
      levels = four_scenario_levels
    )
  ) %>%
  select(
    endpoint,
    scenario,
    metric,
    sample_size,
    mean,
    median,
    minimum,
    maximum,
    sd,
    se
  )


combined_endpoint_distribution_summary


# ============================================================
# Save the distribution-summary table (report in Supplemental!)
# ============================================================

write.csv(
  combined_endpoint_distribution_summary,
  file = file.path(
    output_dir,
    paste0(
      "BTEXSN_benzene_hematological_reproductive_",
      "distribution_summary.csv"
    )
  ),
  row.names = FALSE
)
```





#plot figure
```{r}
# ============================================================
# Determine log-scale range
# ============================================================

combined_log_lower_exponent <- floor(
  log10(
    min(
      combined_endpoint_plot_df$value,
      na.rm = TRUE
    )
  )
)

combined_log_upper_exponent <- ceiling(
  log10(
    max(
      combined_endpoint_plot_df$value,
      na.rm = TRUE
    )
  )
)


combined_log_major_breaks <- 10^seq(
  combined_log_lower_exponent,
  combined_log_upper_exponent,
  by = 1
)


# ============================================================
# Define log-axis labels
#
# Keep 0.1 and 1 in decimal form.
# Use scientific notation for other powers of ten.
# ============================================================

combined_log_major_labels <- function(x) {

  ifelse(
    dplyr::near(
      x,
      0.1
    ),
    "0.1",

    ifelse(
      dplyr::near(
        x,
        1
      ),
      "1",

      format(
        x,
        scientific = TRUE,
        trim = TRUE,
        digits = 1
      )
    )
  )
}


# ============================================================
# Create the final publication figure
# ============================================================

benzene_combined_endpoints_beeswarm_log <- ggplot(
  combined_endpoint_plot_df,
  aes(
    x = scenario,
    y = value,
    color = endpoint
  )
) +

  geom_quasirandom(
    aes(
      group = interaction(
        scenario,
        endpoint
      )
    ),
    dodge.width = 0.65,
    width = 0.18,
    alpha = 0.35,
    size = 0.4
  ) +

  geom_hline(
    yintercept = c(
      0.1,
      1
    ),
    linetype = "dotted",
    linewidth = 0.8,
    color = "grey40"
  ) +

  scale_y_log10(
    breaks = combined_log_major_breaks,
    minor_breaks = NULL,
    labels = combined_log_major_labels,
    expand = expansion(
      mult = c(
        0.03,
        0.05
      )
    )
  ) +

  scale_color_manual(
    values = c(
      "Hematological" = "#0072B2",
      "Reproductive" = "#D55E00"
    ),
    labels = c(
      "Hematological" = "Hematological (most sensitive)",
      "Reproductive" = "Reproductive"
    ),
    guide = guide_legend(
      override.aes = list(
        alpha = 1,
        size = 3
      )
    )
  ) +

  scale_x_discrete(
    labels = four_scenario_labels
  ) +

  labs(
    x = NULL,
    y = "HQ or HI (log scale)",
    color = "Endpoint"
  ) +

  theme_bw() +

  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    ),

    panel.grid.major.x =
      element_blank(),

    panel.grid.minor =
      element_blank(),

    panel.grid.major.y =
      element_line(
        color = "grey85",
        linewidth = 0.4
      ),

    legend.position = "bottom"
  )


benzene_combined_endpoints_beeswarm_log


# ============================================================
# Save the final publication figure
# ============================================================

ggsave(
  filename = paste0(
    "BTEXSN_benzene_hematological_reproductive_",
    "aggregate_cumulative_log.png"
  ),
  plot = benzene_combined_endpoints_beeswarm_log,
  path = output_dir,
  width = 8,
  height = 5.5,
  dpi = 600
)
```



#create table for block threshold numbers/percentages that goes with the figure
```{r}
# ============================================================
# Create the raw eight-row block threshold table
#
# Thresholds:
# value >= 0.1
# value >= 1
#
# Percentages use all blocks with nonmissing values within
# each endpoint and scenario, including blocks with value zero.
# ============================================================

combined_endpoint_exceedance_table_raw <-
  combined_endpoint_analysis_df %>%
  group_by(
    endpoint,
    scenario,
    metric
  ) %>%
  summarise(
    total_blocks = n_distinct(
      geoid20
    ),

    blocks_ge_0_1 = n_distinct(
      geoid20[value >= 0.1]
    ),

    percent_ge_0_1 =
      100 *
      blocks_ge_0_1 /
      total_blocks,

    blocks_ge_1 = n_distinct(
      geoid20[value >= 1]
    ),

    percent_ge_1 =
      100 *
      blocks_ge_1 /
      total_blocks,

    .groups = "drop"
  ) %>%
  mutate(
    endpoint = as.character(
      endpoint
    ),

    scenario = as.character(
      scenario
    ),

    metric = as.character(
      metric
    )
  ) %>%
  arrange(
    factor(
      endpoint,
      levels = endpoint_levels
    ),

    factor(
      scenario,
      levels = four_scenario_levels
    )
  )


combined_endpoint_exceedance_table_raw


# ============================================================
# Validate the eight-row threshold table
# ============================================================

stopifnot(
  nrow(
    combined_endpoint_exceedance_table_raw
  ) == 8
)


combined_endpoint_exceedance_table_raw %>%
  summarise(
    any_ge_0_1_above_total =
      any(
        blocks_ge_0_1 >
          total_blocks
      ),

    any_ge_1_above_ge_0_1 =
      any(
        blocks_ge_1 >
          blocks_ge_0_1
      ),

    any_percent_ge_0_1_above_100 =
      any(
        percent_ge_0_1 >
          100
      ),

    any_percent_ge_1_above_100 =
      any(
        percent_ge_1 >
          100
      )
  )


# ============================================================
# Create the formatted eight-row threshold table
#
# Format:
# n (percentage%)
# ============================================================

combined_endpoint_exceedance_table <-
  combined_endpoint_exceedance_table_raw %>%
  mutate(
    total_blocks_formatted = format(
      total_blocks,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),

    blocks_ge_0_1_n_percent = paste0(
      format(
        blocks_ge_0_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_0_1,
        format = "f",
        digits = 2
      ),
      "%)"
    ),

    blocks_ge_1_n_percent = paste0(
      format(
        blocks_ge_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_1,
        format = "f",
        digits = 2
      ),
      "%)"
    )
  ) %>%
  transmute(
    Endpoint = endpoint,
    Scenario = scenario,
    Metric = metric,

    `Total blocks` =
      total_blocks_formatted,

    `Blocks >= 0.1, n (%)` =
      blocks_ge_0_1_n_percent,

    `Blocks >= 1, n (%)` =
      blocks_ge_1_n_percent
  )


combined_endpoint_exceedance_table


# Save formatted eight-row table
write.csv(
  combined_endpoint_exceedance_table,
  file = file.path(
    output_dir,
    paste0(
      "BTEXSN_benzene_hematological_reproductive_",
      "block_threshold_table.csv"
    )
  ),
  row.names = FALSE
)


# Save numeric version for analysis/checking
write.csv(
  combined_endpoint_exceedance_table_raw,
  file = file.path(
    output_dir,
    paste0(
      "BTEXSN_benzene_hematological_reproductive_",
      "block_threshold_table_raw.csv"
    )
  ),
  row.names = FALSE
)
```



#for supplemental: block exceedance numbers/percentages for HIs for each endpoint
```{r}
# ============================================================
# SUPPLEMENTAL TABLE:
#
# Population priority is handled in BigQuery:
# 1. Use child ("ch") HI when available
# 2. Otherwise retain population-unspecified ("NA") HI
#
# The calculations are performed in BigQuery so only one
# summarized row per endpoint is downloaded to R.
#
# Threshold definitions:
# HI >= 0.1
# HI >= 1
# ============================================================


SQL_all_endpoint_hi_thresholds <- "
WITH eligible_hi AS (

  SELECT
    h.geoid20,

    LOWER(
      TRIM(h.endpoint)
    ) AS endpoint,

    LOWER(
      TRIM(h.pop)
    ) AS pop,

    h.HI_tot_comb,

    ROW_NUMBER() OVER (
      PARTITION BY
        h.geoid20,
        LOWER(TRIM(h.endpoint))

      ORDER BY
        CASE
          WHEN LOWER(TRIM(h.pop)) = 'ch' THEN 1
          WHEN LOWER(TRIM(h.pop)) = 'na' THEN 2
          ELSE 3
        END
    ) AS population_rank

  FROM
    `edf-aq-data.TSCA.BTEXSN_combined_HIs` h

  JOIN (
    SELECT DISTINCT
      GEOID
    FROM
      `edf-aq-data.TSCA.filtblocks_BTEXSN`
  ) f
    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'na'
    )

    AND h.endpoint IS NOT NULL

    AND h.HI_tot_comb IS NOT NULL
),


selected_hi AS (

  SELECT
    geoid20,
    endpoint,
    pop AS selected_population,
    HI_tot_comb

  FROM
    eligible_hi

  WHERE
    population_rank = 1
)


SELECT
  endpoint,

  COUNT(
    DISTINCT geoid20
  ) AS total_blocks,

  COUNT(
    DISTINCT IF(
      HI_tot_comb >= 0.1,
      geoid20,
      NULL
    )
  ) AS blocks_ge_0_1,

  SAFE_MULTIPLY(
    SAFE_DIVIDE(
      COUNT(
        DISTINCT IF(
          HI_tot_comb >= 0.1,
          geoid20,
          NULL
        )
      ),
      COUNT(
        DISTINCT geoid20
      )
    ),
    100
  ) AS percent_ge_0_1,

  COUNT(
    DISTINCT IF(
      HI_tot_comb >= 1,
      geoid20,
      NULL
    )
  ) AS blocks_ge_1,

  SAFE_MULTIPLY(
    SAFE_DIVIDE(
      COUNT(
        DISTINCT IF(
          HI_tot_comb >= 1,
          geoid20,
          NULL
        )
      ),
      COUNT(
        DISTINCT geoid20
      )
    ),
    100
  ) AS percent_ge_1

FROM
  selected_hi

GROUP BY
  endpoint

ORDER BY
  endpoint
"


# ============================================================
# Download the summarized endpoint table
# ============================================================

cumulative_hi_threshold_table_raw <- bq_project_query(
  project_id,
  SQL_all_endpoint_hi_thresholds
) %>%
  bq_table_download() %>%
  transmute(
    endpoint = str_to_title(
      endpoint
    ),

    total_blocks = as.numeric(
      total_blocks
    ),

    blocks_ge_0_1 = as.numeric(
      blocks_ge_0_1
    ),

    percent_ge_0_1 = as.numeric(
      percent_ge_0_1
    ),

    blocks_ge_1 = as.numeric(
      blocks_ge_1
    ),

    percent_ge_1 = as.numeric(
      percent_ge_1
    )
  ) %>%
  arrange(
    endpoint
  )


cumulative_hi_threshold_table_raw


# ============================================================
# Validate supplemental HI table
# ============================================================

cumulative_hi_threshold_validation <-
  cumulative_hi_threshold_table_raw %>%
  summarise(
    any_ge_0_1_above_total =
      any(
        blocks_ge_0_1 >
          total_blocks
      ),

    any_ge_1_above_ge_0_1 =
      any(
        blocks_ge_1 >
          blocks_ge_0_1
      ),

    any_percent_ge_0_1_above_100 =
      any(
        percent_ge_0_1 >
          100
      ),

    any_percent_ge_1_above_100 =
      any(
        percent_ge_1 >
          100
      )
  )


cumulative_hi_threshold_validation


# ============================================================
# Create the formatted four-column supplemental table
#
# Counts and percentages are formatted as:
# n (percentage%)
# ============================================================

cumulative_hi_threshold_table <-
  cumulative_hi_threshold_table_raw %>%
  mutate(
    total_blocks_formatted = format(
      total_blocks,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),

    blocks_ge_0_1_n_percent = paste0(
      format(
        blocks_ge_0_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_0_1,
        format = "f",
        digits = 2
      ),
      "%)"
    ),

    blocks_ge_1_n_percent = paste0(
      format(
        blocks_ge_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_1,
        format = "f",
        digits = 2
      ),
      "%)"
    )
  ) %>%
  transmute(
    Endpoint = endpoint,

    `Total blocks` =
      total_blocks_formatted,

    `Blocks >= 0.1, n (%)` =
      blocks_ge_0_1_n_percent,

    `Blocks >= 1, n (%)` =
      blocks_ge_1_n_percent
  )


cumulative_hi_threshold_table


# ============================================================
# Save formatted supplemental HI table
# ============================================================

write.csv(
  cumulative_hi_threshold_table,
  file = file.path(
    output_dir,
    "BTEXSN_endpoint_cumulative_HI_threshold_table.csv"
  ),
  row.names = FALSE
)


# ============================================================
# Save raw numeric supplemental HI table
# ============================================================

write.csv(
  cumulative_hi_threshold_table_raw,
  file = file.path(
    output_dir,
    "BTEXSN_endpoint_cumulative_HI_threshold_table_raw.csv"
  ),
  row.names = FALSE
)
```



#for supplemental: table of summary information (n, min, median, mean, max, sd, se) for each each HI for each endpoint and population (ch, ad, NA)
```{r}
# ============================================================
# SUPPLEMENTAL TABLE:
# Combined HI summary statistics for every endpoint and
# cumulative HI type
#
# Cumulative HI types:
# - Child (air + drinking water)
# - Adult (air + drinking water)
# - Air only
#
# Summary statistics:
# n, minimum, median, mean, maximum, SD, SE
#
# All calculations are performed in BigQuery.
# Only the summarized results are downloaded to R.
# ============================================================


SQL_combined_hi_type_summary <- "
WITH filtered_blocks AS (

  SELECT DISTINCT
    GEOID

  FROM
    `edf-aq-data.TSCA.filtblocks_BTEXSN`
),


-- ------------------------------------------------------------
-- Prepare eligible combined-HI records
-- ------------------------------------------------------------

combined_hi AS (

  SELECT
    h.geoid20,

    LOWER(
      TRIM(h.endpoint)
    ) AS endpoint,

    LOWER(
      TRIM(h.pop)
    ) AS population_code,

    h.HI_tot_comb AS value

  FROM
    `edf-aq-data.TSCA.BTEXSN_combined_HIs` h

  INNER JOIN
    filtered_blocks f

    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'ad',
      'na'
    )

    AND h.endpoint IS NOT NULL
    AND h.HI_tot_comb IS NOT NULL
),


-- ------------------------------------------------------------
-- Calculate n, minimum, mean, maximum, and SD
-- ------------------------------------------------------------

aggregate_stats AS (

  SELECT
    endpoint,
    population_code,

    COUNT(value) AS n,

    MIN(value) AS minimum,

    AVG(value) AS mean,

    MAX(value) AS maximum,

    STDDEV_SAMP(value) AS sd

  FROM
    combined_hi

  GROUP BY
    endpoint,
    population_code
),


-- ------------------------------------------------------------
-- Calculate exact median
-- ------------------------------------------------------------

median_stats AS (

  SELECT DISTINCT
    endpoint,
    population_code,

    PERCENTILE_CONT(
      value,
      0.5
    ) OVER (
      PARTITION BY
        endpoint,
        population_code
    ) AS median

  FROM
    combined_hi
)


-- ------------------------------------------------------------
-- Final summarized table
-- ------------------------------------------------------------

SELECT
  a.endpoint,
  a.population_code,
  a.n,
  a.minimum,
  m.median,
  a.mean,
  a.maximum,
  a.sd,

  SAFE_DIVIDE(
    a.sd,
    SQRT(a.n)
  ) AS se

FROM
  aggregate_stats a

LEFT JOIN
  median_stats m

USING (
  endpoint,
  population_code
)

ORDER BY
  endpoint,

  CASE population_code
    WHEN 'ch' THEN 1
    WHEN 'ad' THEN 2
    WHEN 'na' THEN 3
    ELSE 4
  END
"


# ============================================================
# Download and format the summarized table
# ============================================================

combined_hi_type_summary_table <-
  bq_project_query(
    project_id,
    SQL_combined_hi_type_summary
  ) %>%
  bq_table_download() %>%
  transmute(
    Endpoint = str_to_title(
      endpoint
    ),

    `Cumulative HI type` = case_when(
      population_code == "ch" ~
        "Child (air + drinking water)",

      population_code == "ad" ~
        "Adult (air + drinking water)",

      population_code == "na" ~
        "Air only",

      TRUE ~
        population_code
    ),

    n = as.numeric(
      n
    ),

    Minimum = as.numeric(
      minimum
    ),

    Median = as.numeric(
      median
    ),

    Mean = as.numeric(
      mean
    ),

    Maximum = as.numeric(
      maximum
    ),

    SD = as.numeric(
      sd
    ),

    SE = as.numeric(
      se
    )
  ) %>%
  mutate(
    `Cumulative HI type` = factor(
      `Cumulative HI type`,
      levels = c(
        "Child (air + drinking water)",
        "Adult (air + drinking water)",
        "Air only"
      )
    )
  ) %>%
  arrange(
    Endpoint,
    `Cumulative HI type`
  ) %>%
  mutate(
    `Cumulative HI type` = as.character(
      `Cumulative HI type`
    )
  )


# View final table
combined_hi_type_summary_table


# ============================================================
# Validate the resulting table
#
# All values should return FALSE.
# ============================================================

combined_hi_type_summary_validation <-
  combined_hi_type_summary_table %>%
  summarise(
    any_n_below_1 =
      any(
        n < 1
      ),

    any_minimum_above_median =
      any(
        Minimum > Median
      ),

    any_median_above_maximum =
      any(
        Median > Maximum
      ),

    any_mean_below_minimum =
      any(
        Mean < Minimum
      ),

    any_mean_above_maximum =
      any(
        Mean > Maximum
      ),

    any_negative_sd =
      any(
        SD < 0,
        na.rm = TRUE
      ),

    any_negative_se =
      any(
        SE < 0,
        na.rm = TRUE
      )
  )


combined_hi_type_summary_validation


# ============================================================
# Save the final table
# ============================================================

write.csv(
  combined_hi_type_summary_table,
  file = file.path(
    output_dir,
    "BTEXSN_combined_HI_summary_by_endpoint_and_type.csv"
  ),
  row.names = FALSE
)
```



#table for supplemental on how many census blocks have HI of 0.1 or 1 exceedances for more than one endpoint
```{r}
# ============================================================
# 1. Query multi-endpoint exceedance patterns
# ============================================================

SQL_multi_endpoint_exceedance <- "

WITH filtered_blocks AS (

  SELECT DISTINCT
    GEOID

  FROM
    `edf-aq-data.TSCA.filtblocks_BTEXSN`
),


-- ------------------------------------------------------------
-- Select one cumulative HI per block-endpoint
--
-- Population priority:
-- 1. child
-- 2. NA / air-only
-- ------------------------------------------------------------

eligible_hi AS (

  SELECT
    h.geoid20,

    LOWER(
      TRIM(h.endpoint)
    ) AS endpoint,

    LOWER(
      TRIM(h.pop)
    ) AS pop,

    h.HI_tot_comb,

    ROW_NUMBER() OVER (

      PARTITION BY
        h.geoid20,
        LOWER(TRIM(h.endpoint))

      ORDER BY
        CASE

          WHEN LOWER(TRIM(h.pop)) = 'ch'
            THEN 1

          WHEN LOWER(TRIM(h.pop)) = 'na'
            THEN 2

          ELSE 3

        END

    ) AS population_rank

  FROM
    `edf-aq-data.TSCA.BTEXSN_combined_HIs` h

  INNER JOIN
    filtered_blocks f

    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'na'
    )

    AND h.endpoint IS NOT NULL

    AND h.HI_tot_comb IS NOT NULL
),


-- ------------------------------------------------------------
-- Keep selected population record
-- ------------------------------------------------------------

selected_hi AS (

  SELECT
    geoid20,
    endpoint,
    pop AS selected_population,
    HI_tot_comb

  FROM
    eligible_hi

  WHERE
    population_rank = 1
),


-- ------------------------------------------------------------
-- Define HI thresholds
--
-- threshold_id is INT64 and is used for grouping/windowing.
-- threshold_value is the actual HI threshold.
-- ------------------------------------------------------------

thresholds AS (

  SELECT
    1 AS threshold_id,
    0.1 AS threshold_value

  UNION ALL

  SELECT
    2 AS threshold_id,
    1.0 AS threshold_value
),


-- ------------------------------------------------------------
-- Identify every block-endpoint exceeding each threshold
-- ------------------------------------------------------------

exceeded_endpoints AS (

  SELECT
    h.geoid20,
    t.threshold_id,
    t.threshold_value,
    h.endpoint

  FROM
    selected_hi h

  CROSS JOIN
    thresholds t

  WHERE
    h.HI_tot_comb >= t.threshold_value
),


-- ------------------------------------------------------------
-- Determine total number and full combination of endpoints
-- exceeded within each block
-- ------------------------------------------------------------

block_endpoint_patterns AS (

  SELECT
    geoid20,
    threshold_id,
    threshold_value,

    COUNT(
      DISTINCT endpoint
    ) AS n_endpoints_exceeded,

    STRING_AGG(
      DISTINCT endpoint,
      ', '
      ORDER BY endpoint
    ) AS endpoint_combination

  FROM
    exceeded_endpoints

  GROUP BY
    geoid20,
    threshold_id,
    threshold_value
),


-- ------------------------------------------------------------
-- Expand each block into at-least-N categories
--
-- Example:
-- If n_endpoints_exceeded = 4,
-- create rows for:
-- at_least_n = 1
-- at_least_n = 2
-- at_least_n = 3
-- at_least_n = 4
-- ------------------------------------------------------------

at_least_endpoint_patterns AS (

  SELECT
    geoid20,
    threshold_id,
    threshold_value,
    n_endpoints_exceeded,
    endpoint_combination,
    at_least_n

  FROM
    block_endpoint_patterns,

  UNNEST(
    GENERATE_ARRAY(
      1,
      n_endpoints_exceeded
    )
  ) AS at_least_n
),


-- ------------------------------------------------------------
-- Count total blocks exceeding AT LEAST N endpoints
-- ------------------------------------------------------------

endpoint_number_totals AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,

    COUNT(
      DISTINCT geoid20
    ) AS n_blocks

  FROM
    at_least_endpoint_patterns

  GROUP BY
    threshold_id,
    threshold_value,
    at_least_n
),


-- ------------------------------------------------------------
-- Count full endpoint combinations among blocks in each
-- at-least-N category
-- ------------------------------------------------------------

combination_counts AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,
    endpoint_combination,

    COUNT(
      DISTINCT geoid20
    ) AS n_blocks_combination

  FROM
    at_least_endpoint_patterns

  GROUP BY
    threshold_id,
    threshold_value,
    at_least_n,
    endpoint_combination
),


-- ------------------------------------------------------------
-- Rank endpoint combinations within each threshold and
-- at-least-N category
--
-- threshold_id is used rather than threshold_value because
-- BigQuery does not allow FLOAT64 in PARTITION BY.
-- ------------------------------------------------------------

ranked_combinations AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,
    endpoint_combination,
    n_blocks_combination,

    DENSE_RANK() OVER (

      PARTITION BY
        threshold_id,
        at_least_n

      ORDER BY
        n_blocks_combination DESC

    ) AS combination_rank

  FROM
    combination_counts
),


-- ------------------------------------------------------------
-- Retain the most common endpoint combination
--
-- If multiple combinations are tied for most common,
-- all tied combinations are retained and separated by |
-- ------------------------------------------------------------

most_common_combinations AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,

    STRING_AGG(
      endpoint_combination,
      ' | '
      ORDER BY endpoint_combination
    ) AS most_common_endpoints,

    MAX(
      n_blocks_combination
    ) AS n_blocks_most_common

  FROM
    ranked_combinations

  WHERE
    combination_rank = 1

  GROUP BY
    threshold_id,
    threshold_value,
    at_least_n
),


-- ------------------------------------------------------------
-- Total number of distinct Census blocks evaluated
--
-- Includes all blocks represented in selected_hi,
-- regardless of whether they exceed either threshold.
-- ------------------------------------------------------------

total_blocks AS (

  SELECT

    COUNT(
      DISTINCT geoid20
    ) AS total_blocks

  FROM
    selected_hi
)


-- ------------------------------------------------------------
-- Final query output
-- ------------------------------------------------------------

SELECT
  t.threshold_value AS threshold,

  t.at_least_n AS n_endpoints_exceeded,

  b.total_blocks,

  t.n_blocks,

  m.most_common_endpoints,

  m.n_blocks_most_common,

  SAFE_MULTIPLY(
    SAFE_DIVIDE(
      m.n_blocks_most_common,
      t.n_blocks
    ),
    100
  ) AS percent_blocks_most_common

FROM
  endpoint_number_totals t

LEFT JOIN
  most_common_combinations m

  ON
    t.threshold_id =
      m.threshold_id

    AND t.at_least_n =
      m.at_least_n

CROSS JOIN
  total_blocks b

ORDER BY
  t.threshold_id,
  n_endpoints_exceeded
"


# ============================================================
# 2. Run query and download raw long-format results
# ============================================================

multi_endpoint_exceedance_table_raw <-
  bq_project_query(
    project_id,
    SQL_multi_endpoint_exceedance
  ) %>%
  bq_table_download() %>%
  transmute(

    threshold =
      as.numeric(
        threshold
      ),

    n_endpoints_exceeded =
      as.numeric(
        n_endpoints_exceeded
      ),

    total_blocks =
      as.numeric(
        total_blocks
      ),

    n_blocks =
      as.numeric(
        n_blocks
      ),

    most_common_endpoints =
      as.character(
        most_common_endpoints
      ),

    n_blocks_most_common =
      as.numeric(
        n_blocks_most_common
      ),

    percent_blocks_most_common =
      as.numeric(
        percent_blocks_most_common
      )
  ) %>%
  arrange(
    threshold,
    n_endpoints_exceeded
  )


# View raw long-format results
multi_endpoint_exceedance_table_raw


# ============================================================
# 3. Convert threshold results to wide format
# ============================================================

multi_endpoint_exceedance_table_wide_raw <-
  multi_endpoint_exceedance_table_raw %>%

  mutate(

    threshold_label =
      case_when(

        dplyr::near(
          threshold,
          0.1
        ) ~ "0_1",

        dplyr::near(
          threshold,
          1
        ) ~ "1",

        TRUE ~
          as.character(
            threshold
          )
      )
  ) %>%

  select(
    n_endpoints_exceeded,
    total_blocks,
    threshold_label,
    n_blocks,
    most_common_endpoints,
    n_blocks_most_common,
    percent_blocks_most_common
  ) %>%

  pivot_wider(

    names_from =
      threshold_label,

    values_from = c(
      n_blocks,
      most_common_endpoints,
      n_blocks_most_common,
      percent_blocks_most_common
    ),

    names_glue =
      "{.value}_{threshold_label}"
  ) %>%

  arrange(
    n_endpoints_exceeded
  )


multi_endpoint_exceedance_table_wide_raw

# ============================================================
# 4. Fill missing threshold combinations
#
# Example:
# If blocks exist for at least 8 endpoints at HI >= 0.1
# but none exist at HI >= 1, the HI >= 1 columns are
# converted from NA to zero / None.
# ============================================================

multi_endpoint_exceedance_table_wide_raw <-
  multi_endpoint_exceedance_table_wide_raw %>%

  mutate(

    n_blocks_0_1 =
      replace_na(
        n_blocks_0_1,
        0
      ),

    n_blocks_1 =
      replace_na(
        n_blocks_1,
        0
      ),


    most_common_endpoints_0_1 =
      replace_na(
        most_common_endpoints_0_1,
        "None"
      ),

    most_common_endpoints_1 =
      replace_na(
        most_common_endpoints_1,
        "None"
      ),


    n_blocks_most_common_0_1 =
      replace_na(
        n_blocks_most_common_0_1,
        0
      ),

    n_blocks_most_common_1 =
      replace_na(
        n_blocks_most_common_1,
        0
      ),


    percent_blocks_most_common_0_1 =
      replace_na(
        percent_blocks_most_common_0_1,
        0
      ),

    percent_blocks_most_common_1 =
      replace_na(
        percent_blocks_most_common_1,
        0
      )
  )


multi_endpoint_exceedance_table_wide_raw

# ============================================================
# 5. Create formatted supplemental table
# ============================================================

multi_endpoint_exceedance_table <-
  multi_endpoint_exceedance_table_wide_raw %>%

  transmute(

    `Chemical group` =
      "VOCs",


    `Total number of Census blocks` =
      format(
        total_blocks,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),


    `Number of endpoints exceeded (at least)` =
      n_endpoints_exceeded,


    `Number of Census blocks with HI >= 0.1` =
      format(
        n_blocks_0_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),


    `Most common endpoints with HI >= 0.1` =
      case_when(

        most_common_endpoints_0_1 == "None" ~
          "None",

        TRUE ~
          str_to_title(
            most_common_endpoints_0_1
          )
      ),


    `Blocks with most common HI >= 0.1 combination, n (%)` =
      case_when(

        n_blocks_0_1 == 0 ~
          "0 (0.00%)",

        TRUE ~
          paste0(

            format(
              n_blocks_most_common_0_1,
              big.mark = ",",
              scientific = FALSE,
              trim = TRUE
            ),

            " (",

            formatC(
              percent_blocks_most_common_0_1,
              format = "f",
              digits = 2
            ),

            "%)"
          )
      ),


    `Number of Census blocks with HI >= 1` =
      format(
        n_blocks_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),


    `Most common endpoints with HI >= 1` =
      case_when(

        most_common_endpoints_1 == "None" ~
          "None",

        TRUE ~
          str_to_title(
            most_common_endpoints_1
          )
      ),


    `Blocks with most common HI >= 1 combination, n (%)` =
      case_when(

        n_blocks_1 == 0 ~
          "0 (0.00%)",

        TRUE ~
          paste0(

            format(
              n_blocks_most_common_1,
              big.mark = ",",
              scientific = FALSE,
              trim = TRUE
            ),

            " (",

            formatC(
              percent_blocks_most_common_1,
              format = "f",
              digits = 2
            ),

            "%)"
          )
      )
  )


multi_endpoint_exceedance_table

# ============================================================
# 6. Validate results
#
# Because these are AT LEAST N counts:
#
# at least 2 must be <= at least 1
# at least 3 must be <= at least 2
# etc.
#
# All checks below should return FALSE.
# ============================================================

multi_endpoint_exceedance_validation <-
  multi_endpoint_exceedance_table_wide_raw %>%

  arrange(
    n_endpoints_exceeded
  ) %>%

  summarise(

    any_0_1_above_total =
      any(
        n_blocks_0_1 >
          total_blocks
      ),

    any_1_above_total =
      any(
        n_blocks_1 >
          total_blocks
      ),


    any_0_1_count_increases =
      any(
        diff(
          n_blocks_0_1
        ) > 0
      ),

    any_1_count_increases =
      any(
        diff(
          n_blocks_1
        ) > 0
      ),


    any_common_0_1_above_group_total =
      any(
        n_blocks_most_common_0_1 >
          n_blocks_0_1
      ),

    any_common_1_above_group_total =
      any(
        n_blocks_most_common_1 >
          n_blocks_1
      ),


    any_common_percent_0_1_above_100 =
      any(
        percent_blocks_most_common_0_1 >
          100
      ),

    any_common_percent_1_above_100 =
      any(
        percent_blocks_most_common_1 >
          100
      )
  )


multi_endpoint_exceedance_validation

# ============================================================
# 7. Save formatted supplemental table
# ============================================================

write.csv(
  multi_endpoint_exceedance_table,
  file = file.path(
    output_dir,
    "BTEXSN_multi_endpoint_HI_exceedance_table_wide.csv"
  ),
  row.names = FALSE
)


# ============================================================
# 8. Save raw numeric version
# ============================================================

write.csv(
  multi_endpoint_exceedance_table_wide_raw,
  file = file.path(
    output_dir,
    "BTEXSN_multi_endpoint_HI_exceedance_table_wide_raw.csv"
  ),
  row.names = FALSE
)
```







