---
title: "Drinking water analysis for chlorinated solvents"
author: "Paige Varner"
date: "2025-07-08"
output: html_document
---

#load libraries
```{r}
library("dplyr")
library("tidyr")
library("stringr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("car")
library("arrow")
library("progressr")
library("purrr")
library("readr")
library("haven")
library("tidycensus")
detach(package:Rmisc)
detach(package:plyr)

filtblocks = read.csv("~/TSCA/CRA/metdb_tsca/data/filtblocks_chlor.csv")
```


#database connection
```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "airtoxscreen",)

bq_auth(email = NA)
```


## note: running DW analysis after creating a filtblocks_chlor table in BQ during the ATS analysis phase so this table already exists, and subsequent analysis is filtering only for those blocks ##


#summary table for assuming nd is zero (low) and nd is the LOD (high) overall and by child/adult populations (note: the N in these summaries is for block + endpoint combinations)
```{r}
# --- Overall low/high summary for PWS, filtered by filtblocks_chlor ---
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = f.GEOID
)
SELECT 
  low_up,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_HIs
GROUP BY low_up
ORDER BY low_up
"

chlor_HI_summary_low_up <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_summary_low_up, 
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HIsummary_low_up.csv",
  row.names = FALSE
)

# --- Low/high summary by population for PWS ---
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = f.GEOID
)
SELECT 
  low_up,
  pop,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_HIs
GROUP BY low_up, pop
ORDER BY low_up, pop
"

chlor_HI_summary_low_up_pop <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_summary_low_up_pop, 
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HIsummary_low_up_pop.csv",
  row.names = FALSE
)

```


## moving forward - analysis conducted for assuming non-detects are 1/2 the LOD; this pulls from BQ views and tables that end in _half ##


#summary tables (note: the N in these summaries is for block + endpoint combinations, except when broken out by endpoint)
```{r}
# --- Overall summary ---
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
"

chlor_HI_summary <- dbGetQuery(con, SQL)

write.csv(chlor_HI_summary, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HIsummary.csv")

# --- By population ---
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  pop,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
GROUP BY pop
ORDER BY pop
"

chlor_HI_summary_pop <- dbGetQuery(con, SQL)

write.csv(chlor_HI_summary_pop, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HIsummary_pop.csv")

# --- By endpoint ---
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  endpoint,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
GROUP BY endpoint
ORDER BY endpoint
"

chlor_HI_summary_endpoint <- dbGetQuery(con, SQL)

write.csv(chlor_HI_summary_endpoint, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HIsummary_by_endpoint.csv")


# --- By population and endpoint ---
SQL <- "
WITH filtered_blocks AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = CAST(f.GEOID AS STRING)
)
SELECT 
  pop,
  endpoint,
  MIN(HI) AS min,
  MAX(HI) AS max,
  AVG(HI) AS mean,
  APPROX_QUANTILES(HI, 2)[OFFSET(1)] AS median,
  STDDEV(HI) AS std_dev,
  STDDEV(HI) / SQRT(COUNT(HI)) AS std_error,
  COUNT(HI) AS n
FROM filtered_blocks
GROUP BY pop, endpoint
ORDER BY pop, endpoint
"

chlor_HI_summary_pop_endpoint <- dbGetQuery(con, SQL)

write.csv(chlor_HI_summary_pop_endpoint, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HIsummary_pop_endpoint.csv")

```


#delta for each pollutant
```{r}
# --- Overall deltas ---
SQL <- "
WITH filtered_blocks AS (
  SELECT hq.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HQs_half` hq
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON CAST(hq.geoid20 AS STRING) = f.GEOID
)
SELECT
  hq.chem,
  COUNT(*) AS n,
  AVG(hq.hq) AS avg_HQ,
  AVG(hi.HI) AS avg_HI,
  SAFE_DIVIDE(AVG(hi.HI) - AVG(hq.hq), 1) AS delta_avg
FROM filtered_blocks hq
JOIN `edf-aq-data.TSCA.chlor_PWS_HIs_half` hi
  ON hq.geoid20 = hi.geoid20 AND hq.endpoint = hi.endpoint
GROUP BY hq.chem
ORDER BY hq.chem
"

chlor_chem_avg_deltas <- dbGetQuery(con, SQL)
write.csv(chlor_chem_avg_deltas, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/avg_deltas.csv")

# --- Deltas for HI >= 1 ---
SQL <- "
WITH filtered_blocks AS (
  SELECT hq.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HQs_half` hq
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON CAST(hq.geoid20 AS STRING) = f.GEOID
)
SELECT
  hq.chem,
  COUNT(*) AS n,
  AVG(hq.hq) AS avg_HQ,
  AVG(hi.HI) AS avg_HI,
  SAFE_DIVIDE(AVG(hi.HI) - AVG(hq.hq), 1) AS delta_avg
FROM filtered_blocks hq
JOIN `edf-aq-data.TSCA.chlor_PWS_HIs_half` hi
  ON hq.geoid20 = hi.geoid20 AND hq.endpoint = hi.endpoint
WHERE hi.HI >= 1
GROUP BY hq.chem
ORDER BY hq.chem
"

chlor_chem_avg_deltas_1 <- dbGetQuery(con, SQL)
write.csv(chlor_chem_avg_deltas_1, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/avg_deltas_HI1.csv")

# --- Deltas for HI >= 0.1 ---
SQL <- "
WITH filtered_blocks AS (
  SELECT hq.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HQs_half` hq
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON CAST(hq.geoid20 AS STRING) = f.GEOID
)
SELECT
  hq.chem,
  COUNT(*) AS n,
  AVG(hq.hq) AS avg_HQ,
  AVG(hi.HI) AS avg_HI,
  SAFE_DIVIDE(AVG(hi.HI) - AVG(hq.hq), 1) AS delta_avg
FROM filtered_blocks hq
JOIN `edf-aq-data.TSCA.chlor_PWS_HIs_half` hi
  ON hq.geoid20 = hi.geoid20 AND hq.endpoint = hi.endpoint
WHERE hi.HI >= 0.1
GROUP BY hq.chem
ORDER BY hq.chem
"

chlor_chem_avg_deltas_01 <- dbGetQuery(con, SQL)
write.csv(chlor_chem_avg_deltas_01, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/avg_deltas_HI01.csv")

```


#CDFs for children and adults
```{r}
# --- 1. Pull the data ---
project_id <- "edf-aq-data"

SQL <- "
SELECT
  t.geoid20,
  t.endpoint,
  t.pop,
  t.HI
FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
"

chlor_PWS_HIs_filtered <- bq_project_query(project_id, SQL) %>%
  bq_table_download()

# --- 2. Pre-compute CDF by population ---
chlor_cdf <- chlor_PWS_HIs_filtered %>%
  group_by(pop) %>%
  arrange(HI) %>%
  mutate(cdf = row_number() / n()) %>%
  ungroup()

# Optional: thin points (every 10th to speed up plotting)
#chlor_cdf_sample <- chlor_cdf %>%
 # group_by(pop) %>%
  #slice(seq(1, n(), by = 10)) %>%
  #ungroup()

# --- 3. Plot ---
cdf_plot <- ggplot(chlor_cdf, aes(x = HI, y = cdf)) +
  geom_line() +
  # vertical reference lines for thresholds
  geom_vline(xintercept = c(0.1, 1), linetype = "dotted", color = "black") +
  scale_x_log10() +
  labs(
    title = "CDF of PWS HI (Children vs Adults)",
    x = "HI value (log scale)",
    y = "P(HI ≤ x)"
  ) +
  facet_wrap(~pop, scales = "free_y") +
  theme_minimal(base_size = 14)

cdf_plot

ggsave(
  filename = "chlor_PWS_cdf.png",
  device = "png",
  plot = cdf_plot,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/",
  width = 7, height = 5
)

```




#######################################

### analysis with HI threshold of 1

#######################################


# Count of blocks with HIs above 1 for further analysis and HQs for those blocks
```{r}
#overall
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON CAST(t.geoid20 AS STRING) = f.GEOID
)

SELECT
  'HI' AS HI_type,
  COUNT(DISTINCT CASE WHEN HI > 1 THEN geoid20 END) AS unique_blocks_above_1,
  COUNT(DISTINCT geoid20) AS total_unique_blocks,
  COUNT(CASE WHEN HI > 1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs;
"

chlor_blocks_HIsabove1 <- dbGetQuery(con, SQL)

write.csv(chlor_blocks_HIsabove1, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/blocks_HIsabove1_summary.csv",
          row.names = FALSE)


#for each population
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON CAST(t.geoid20 AS STRING) = f.GEOID
)
SELECT
  pop,
  COUNT(DISTINCT CASE WHEN HI > 1 THEN geoid20 END) AS unique_blocks_above_1,
  COUNT(DISTINCT geoid20) AS total_unique_blocks,
  COUNT(CASE WHEN HI > 1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs
GROUP BY pop
ORDER BY pop;
"

chlor_blocks_HIsabove1_by_pop <- dbGetQuery(con, SQL)

write.csv(chlor_blocks_HIsabove1_by_pop, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/blocks_HIsabove1_bypop_summary.csv",
          row.names = FALSE)

```


# Pull HI and HQ dfs of HIs > 1
```{r}
# --- pull blocks with HI > 1 filtered by filtblocks_chlor ---
SQL_hi <- "
SELECT t.*
FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON CAST(t.geoid20 AS STRING) = f.GEOID
WHERE HI > 1
"

chlor_exceedances <- dbGetQuery(con, SQL_hi)

# --- pull HQ data for those blocks and endpoints ---
SQL_hq <- "
SELECT hq.*
FROM `edf-aq-data.TSCA.chlor_PWS_HQs_half` hq
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON CAST(hq.geoid20 AS STRING) = f.GEOID
JOIN (
  SELECT geoid20, endpoint
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half`
  WHERE HI > 1
) t
  ON hq.geoid20 = t.geoid20 AND hq.endpoint = t.endpoint
"

chlor_HQs_HIexceedances <- dbGetQuery(con, SQL_hq)

```


# Blocks exceeding more than one endpoint
```{r}
#overall
blocks_mult_endpoints <- chlor_exceedances %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints <- chlor_exceedances %>%
  semi_join(blocks_mult_endpoints, by = "geoid20") %>%
  left_join(blocks_mult_endpoints, by = "geoid20")

write.csv(blocks_mult_endpoints, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/blocks_mult_endpoints.csv", row.names = FALSE)

#adult
blocks_mult_endpoints_ad <- chlor_exceedances %>%
  filter(pop == "ad") %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints_ad <- chlor_exceedances %>%
  filter(pop == "ad") %>%
  semi_join(blocks_mult_endpoints_ad, by = "geoid20") %>%
  left_join(blocks_mult_endpoints_ad, by = "geoid20")

write.csv(blocks_mult_endpoints_ad, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/blocks_mult_endpoints_ad.csv", row.names = FALSE)

#child
blocks_mult_endpoints_ch <- chlor_exceedances %>%
  filter(pop == "ch") %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints_ch <- chlor_exceedances %>%
  filter(pop == "ch") %>%
  semi_join(blocks_mult_endpoints, by = "geoid20") %>%
  left_join(blocks_mult_endpoints, by = "geoid20")

write.csv(blocks_mult_endpoints_ch, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/blocks_mult_endpoints_ch.csv", row.names = FALSE)

```


#graph average HQs for each endpoint that was exceeded in any block group to see which chemicals contribute 
```{r}
# Join HI and HQ datasets together 
chlor_exceedances_with_HQs <- chlor_exceedances %>%
  left_join(
    chlor_HQs_HIexceedances %>%
      select(geoid20, endpoint, pop, hq, chem),
    by = c("geoid20", "endpoint", "pop")
  )

# HQ averages by endpoint, chemical, and population
HQ_averages <- chlor_exceedances_with_HQs %>%
  group_by(endpoint, chem, pop) %>%
  summarise(
    n = sum(!is.na(hq)),
    avg_HQ = mean(hq, na.rm = TRUE),
    .groups = "drop"
  )

write.csv(HQ_averages, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/HQ_averages.csv")

# stacked bar plot faceted by pop
HQ_averages <- HQ_averages %>%
  mutate(pop = dplyr::recode(pop,
                             "ad" = "Adult",
                             "ch" = "Child"))

chlor_HQtot_bar <- ggplot(HQ_averages, aes(x = endpoint, y = avg_HQ, fill = chem)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average HQ by Endpoint, Pollutant, and Population Group",
    x = "Endpoint",
    y = "Average HQ",
    fill = "Pollutant"
  ) +
  facet_wrap(~ pop) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

chlor_HQtot_bar

ggsave(
  filename = "bar_avgHQ_endpoints_by_pop.png",
  device = "png",
  plot = chlor_HQtot_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/",
  width = 7,
  height = 5
)

```


# Chemicals contributing to HI exceedances for adults (returning 0, so no graph)
```{r}
blocks_HI_only <- chlor_exceedances_with_HQs %>%
  group_by(geoid20, endpoint) %>%
  mutate(max_HQ_block = max(hq, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(pop == "ad" & HI > 1 & (is.na(max_HQ_block) | max_HQ_block <= 1))
```


# Chemicals contributing to HI exceedances for children (returning 0, so no graph)
```{r}
# 1. Join HI and HQ datasets (child population only)
chlor_exceedances_with_HQs <- chlor_exceedances %>%
  filter(pop == "ch") %>%
  left_join(
    chlor_HQs_HIexceedances %>%
      filter(pop == "ch") %>%
      select(geoid20, endpoint, hq, chem),
    by = c("geoid20", "endpoint")
  )

# 2. Compute max HI and max HQ per block/endpoint
blocks_HI_only <- chlor_exceedances_with_HQs %>%
  group_by(geoid20, endpoint) %>%
  summarise(
    HI_block = max(HI, na.rm = TRUE),
    HQ_block_max = max(hq, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(HI_block > 1 & (is.na(HQ_block_max) | HQ_block_max <= 1))

```


#population counts for exceedances
```{r}
# Step 1: Pad geoid20s to 15 characters and extract state/county codes
chlor_exceedances <- chlor_exceedances %>%
  mutate(
    geoid20 = str_pad(geoid20, 15, pad = "0"),
    statefp = substr(geoid20, 1, 2),
    countyfp = substr(geoid20, 3, 5)
  )

# Step 2: Get distinct state-county combinations
state_county_pairs <- chlor_exceedances %>%
  distinct(statefp, countyfp)

# Step 3: For each state-county, get block-level population from Census
get_block_pop <- function(statefp, countyfp) {
  tryCatch({
    get_decennial(
      geography = "block",
      variables = "P1_001N",   # Total population
      year = 2020,
      sumfile = "pl",
      state = statefp,
      county = countyfp,
      geometry = FALSE,
      output = "wide"
    ) %>%
      rename(census_population = P1_001N) %>%
      mutate(GEOID = str_pad(GEOID, 15, pad = "0"))
  }, error = function(e) {
    message(paste("Error in", statefp, countyfp, ":", e$message))
    return(NULL)
  })
}

# Step 4: Pull Census populations for all counties in exceedances
census_pop_all <- map2_dfr(state_county_pairs$statefp,
                           state_county_pairs$countyfp,
                           get_block_pop)

# Step 5: Join population with exceedances
chlor_with_pop <- chlor_exceedances %>%
  left_join(census_pop_all, by = c("geoid20" = "GEOID"))

# Step 6: Reduce to unique blocks (avoid double-counting when multiple endpoints/chems per block)
chlor_unique_blocks <- chlor_with_pop %>%
  filter(HI > 1) %>%
  distinct(geoid20, pop, census_population)

# Step 7: Summarize population by group (adults vs children)
pop_summary <- chlor_unique_blocks %>%
  group_by(pop) %>%  # assuming "pop" is "ad" or "ch"
  summarise(
    total_census_population = sum(census_population, na.rm = TRUE),
    avg_census_population   = mean(census_population, na.rm = TRUE),
    n_blocks                = n_distinct(geoid20),
    .groups = "drop"
  )

# Step 8: Write output to CSV
write_csv(
  pop_summary,
  "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI1/population_summary.csv"
)
```




#######################################

### analysis with HI threshold of 0.1

#######################################



# Count of blocks with HIs above 0.1 for further analysis and HQs for those blocks
```{r}
#overall
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON CAST(t.geoid20 AS STRING) = f.GEOID
)

SELECT
  'HI' AS HI_type,
  COUNT(DISTINCT CASE WHEN HI > 0.1 THEN geoid20 END) AS unique_blocks_above_01,
  COUNT(DISTINCT geoid20) AS total_unique_blocks,
  COUNT(CASE WHEN HI > 0.1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs;
"

chlor_blocks_HIsabove01 <- dbGetQuery(con, SQL)

write.csv(chlor_blocks_HIsabove01, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/blocks_HIsabove01_summary.csv",
          row.names = FALSE)


#for each population
SQL <- "
WITH filtered_HIs AS (
  SELECT t.*
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON CAST(t.geoid20 AS STRING) = f.GEOID
)
SELECT
  pop,
  COUNT(DISTINCT CASE WHEN HI > 0.1 THEN geoid20 END) AS unique_blocks_above_01,
  COUNT(DISTINCT geoid20) AS total_unique_blocks,
  COUNT(CASE WHEN HI > 0.1 THEN 1 END) AS total_exceedances,
  COUNT(*) AS total_block_endpoint_combos
FROM filtered_HIs
GROUP BY pop
ORDER BY pop;
"

chlor_blocks_HIsabove01_by_pop <- dbGetQuery(con, SQL)

write.csv(chlor_blocks_HIsabove01_by_pop, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/blocks_HIsabove01_bypop_summary.csv",
          row.names = FALSE)

```


# Pull HI and HQ dfs of HIs > 0.1
```{r}
# --- pull blocks with HI > 0.1 filtered by filtblocks_chlor ---
SQL_hi <- "
SELECT t.*
FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON CAST(t.geoid20 AS STRING) = f.GEOID
WHERE HI > 0.1
"

chlor_exceedances <- dbGetQuery(con, SQL_hi)

# --- pull HQ data for those blocks and endpoints ---
SQL_hq <- "
SELECT hq.*
FROM `edf-aq-data.TSCA.chlor_PWS_HQs_half` hq
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON CAST(hq.geoid20 AS STRING) = f.GEOID
JOIN (
  SELECT geoid20, endpoint
  FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half`
  WHERE HI > 0.1
) t
  ON hq.geoid20 = t.geoid20 AND hq.endpoint = t.endpoint
"

chlor_HQs_HIexceedances <- dbGetQuery(con, SQL_hq)

```


# Blocks exceeding more than one endpoint
```{r}
#overall
blocks_mult_endpoints <- chlor_exceedances %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints <- chlor_exceedances %>%
  semi_join(blocks_mult_endpoints, by = "geoid20") %>%
  left_join(blocks_mult_endpoints, by = "geoid20")

write.csv(blocks_mult_endpoints, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/blocks_mult_endpoints.csv", row.names = FALSE)

#adult
blocks_mult_endpoints_ad <- chlor_exceedances %>%
  filter(pop == "ad") %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints_ad <- chlor_exceedances %>%
  filter(pop == "ad") %>%
  semi_join(blocks_mult_endpoints_ad, by = "geoid20") %>%
  left_join(blocks_mult_endpoints_ad, by = "geoid20")

write.csv(blocks_mult_endpoints_ad, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/blocks_mult_endpoints_ad.csv", row.names = FALSE)

#child
blocks_mult_endpoints_ch <- chlor_exceedances %>%
  filter(pop == "ch") %>%
  group_by(geoid20) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(sort(unique(endpoint)), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints_ch <- chlor_exceedances %>%
  filter(pop == "ch") %>%
  semi_join(blocks_mult_endpoints, by = "geoid20") %>%
  left_join(blocks_mult_endpoints, by = "geoid20")

write.csv(blocks_mult_endpoints_ch, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/blocks_mult_endpoints_ch.csv", row.names = FALSE)

```


#graph average HQs for each endpoint that was exceeded in any block group to see which chemicals contribute 
```{r}
# Join HI and HQ datasets together 
chlor_exceedances_with_HQs <- chlor_exceedances %>%
  left_join(
    chlor_HQs_HIexceedances %>%
      select(geoid20, endpoint, pop, hq, chem),
    by = c("geoid20", "endpoint", "pop")
  )

# HQ averages by endpoint, chemical, and population
HQ_averages <- chlor_exceedances_with_HQs %>%
  group_by(endpoint, chem, pop) %>%
  summarise(
    n = sum(!is.na(hq)),
    avg_HQ = mean(hq, na.rm = TRUE),
    .groups = "drop"
  )

write.csv(HQ_averages, file = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/HQ_averages.csv")

# stacked bar plot faceted by pop
HQ_averages <- HQ_averages %>%
  mutate(pop = dplyr::recode(pop,
                             "ad" = "Adult",
                             "ch" = "Child"))

chlor_HQtot_bar <- ggplot(HQ_averages, aes(x = endpoint, y = avg_HQ, fill = chem)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average HQ by Endpoint, Pollutant, and Population Group",
    x = "Endpoint",
    y = "Average HQ",
    fill = "Pollutant"
  ) +
  facet_wrap(~ pop) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

chlor_HQtot_bar

ggsave(
  filename = "bar_avgHQ_endpoints_by_pop.png",
  device = "png",
  plot = chlor_HQtot_bar,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/",
  width = 7,
  height = 5
)

```


# Chemicals contributing to HI exceedances for adults (returning 0, no exceedance of this type) and children
```{r}
# 1. Join HI and HQ datasets (child population only)
chlor_exceedances_with_HQs <- chlor_exceedances %>%
  filter(pop == "ch") %>%
  left_join(
    chlor_HQs_HIexceedances %>%
      filter(pop == "ch") %>%
      select(geoid20, endpoint, hq, chem),
    by = c("geoid20", "endpoint")
  )

# 2. Compute max HI and max HQ per block/endpoint
blocks_HI_only <- chlor_exceedances_with_HQs %>%
  group_by(geoid20, endpoint) %>%
  summarise(
    HI_block = max(HI, na.rm = TRUE),
    HQ_block_max = max(hq, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(HI_block > 0.1 & (is.na(HQ_block_max) | HQ_block_max <= 0.1))

# 3. Keep all HQ rows for blocks meeting the condition
chlor_exceedances_HI_only <- chlor_exceedances_with_HQs %>%
  semi_join(blocks_HI_only, by = c("geoid20", "endpoint"))

# 4. Optional: Create block labels per endpoint for plotting
chlor_exceedances_HI_only <- chlor_exceedances_HI_only %>%
  group_by(endpoint) %>%
  mutate(block_label = paste0("Block ", dense_rank(desc(HI)))) %>%
  ungroup()

# 5. Plot HQ by pollutant for all selected blocks
chlor_HQblocks_bar <- ggplot(chlor_exceedances_HI_only, 
                             aes(x = geoid20, y = hq, fill = chem)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ endpoint, scales = "free") +
  labs(
    title = "HQ by Pollutant and Blocks per Endpoint (HI > 0.1, Children, HQ ≤ 0.1)",
    x = "Block",
    y = "HQ",
    fill = "Pollutant"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10, face = "bold")
  )

chlor_HQblocks_bar

ggsave(filename = "bar_HQ_blocks_child.png", 
       device = "png", 
       plot = chlor_HQblocks_bar, 
       path = "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/", 
       width = 7, height = 5)

```


#population counts for exceedances 
```{r}
# Step 1: Pad geoid20s to 15 characters and extract state/county codes
chlor_exceedances <- chlor_exceedances %>%
  mutate(
    geoid20 = str_pad(geoid20, 15, pad = "0"),
    statefp = substr(geoid20, 1, 2),
    countyfp = substr(geoid20, 3, 5)
  )

# Step 2: Get distinct state-county combinations
state_county_pairs <- chlor_exceedances %>%
  distinct(statefp, countyfp)

# Step 3: For each state-county, get block-level population from Census
get_block_pop <- function(statefp, countyfp) {
  tryCatch({
    get_decennial(
      geography = "block",
      variables = "P1_001N",   # Total population
      year = 2020,
      sumfile = "pl",
      state = statefp,
      county = countyfp,
      geometry = FALSE,
      output = "wide"
    ) %>%
      rename(census_population = P1_001N) %>%
      mutate(GEOID = str_pad(GEOID, 15, pad = "0"))
  }, error = function(e) {
    message(paste("Error in", statefp, countyfp, ":", e$message))
    return(NULL)
  })
}

# Step 4: Pull Census populations for all counties in exceedances
census_pop_all <- map2_dfr(state_county_pairs$statefp,
                           state_county_pairs$countyfp,
                           get_block_pop)

# Step 5: Join population with exceedances
chlor_with_pop <- chlor_exceedances %>%
  left_join(census_pop_all, by = c("geoid20" = "GEOID"))

# Step 6: Reduce to unique blocks (avoid double-counting when multiple endpoints/chems per block)
chlor_unique_blocks <- chlor_with_pop %>%
  filter(HI > 0.1) %>%
  distinct(geoid20, pop, census_population)

# Step 7: Summarize population by group (adults vs children)
pop_summary <- chlor_unique_blocks %>%
  group_by(pop) %>%  # assuming "pop" is "ad" or "ch"
  summarise(
    total_census_population = sum(census_population, na.rm = TRUE),
    avg_census_population   = mean(census_population, na.rm = TRUE),
    n_blocks                = n_distinct(geoid20),
    .groups = "drop"
  )

# Step 8: Write output to CSV
write_csv(
  pop_summary,
  "~/TSCA/CRA/metdb_tsca/output/chlor/Water/HI01/population_summary.csv"
)
```


