---
title: "Combined analysis for chlorinated solvents"
author: "Paige Varner"
date: "2025-07-08"
output: html_document
---



#load libraries
```{r}
library("dplyr")
library("tidyr")
library("stringr")
library("bigrquery")
library("DBI")
library("pheatmap")
library("RColorBrewer")
library("Rmisc")
library("stats")
library("rstatix")
library("ggplot2")
library("car")
library("arrow")
library("progressr")
library("purrr")
library("readr")
library("haven")
library("ggokabeito")
library("ggbeeswarm")
library("patchwork")
library(tidycensus)
detach(package:Rmisc)
detach(package:plyr)

```


#database connection
```{r}
bq_auth(email = NA)

con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "TSCA",)

```



#important sensitivity analysis on combination statistics: 15,166 blocks are included in the PWS dataset but not the ATS dataset. need to check the population from the Census data on these blocks to see if they weren't included in ATS because ATS doesn't model exposures for blocks with population = 0. It makes sense otherwise that there are lots of blocks that have ATS data and not PWS data because millions are not on PWS.
```{r}
# Step 1: Query unmatched PWS blocks from BigQuery
SQL <- "
SELECT DISTINCT
  LPAD(CAST(pws.geoid20 AS STRING), 15, '0') AS geoid20
FROM `edf-aq-data.TSCA.chlor_PWS_HQs_half` AS pws
WHERE LPAD(CAST(pws.geoid20 AS STRING), 15, '0') NOT IN (
  SELECT DISTINCT LPAD(CAST(block AS STRING), 15, '0')
  FROM `edf-aq-data.TSCA.chlor_ATS_HIs`
)
"
unmatched_PWS <- dbGetQuery(con, SQL)

# Step 2: Get state FIPS codes from unmatched blocks
unmatched_PWS <- unmatched_PWS %>%
  mutate(state_fips = substr(geoid20, 1, 2))

states_needed <- unique(unmatched_PWS$state_fips)

# Step 4: Loop through those states to pull 2020 block-level population
pop_blocks_all <- data.frame()

for (fips in states_needed) {
  message("Getting population for state FIPS: ", fips)
  
  tryCatch({
    state_blocks <- get_decennial(
      geography = "block",
      variables = "P1_001N",  # total population 2020 PL data
      year = 2020,
      state = fips,
      geometry = FALSE,
      output = "wide",
      dataset = "pl"
    )
    
    pop_blocks_all <- bind_rows(pop_blocks_all, state_blocks)
  }, error = function(e) {
    message("Error for state ", fips, ": ", e$message)
  })
}

# Step 5: Join population to unmatched PWS blocks
pws_blocks_with_pop <- unmatched_PWS %>%
  left_join(pop_blocks_all, by = c("geoid20" = "GEOID"))

# Step 6: Summarize
summary_table <- pws_blocks_with_pop %>%
  mutate(pop = P1_001N) %>%
  summarise(
    total_blocks = n(),
    blocks_with_population = sum(!is.na(pop) & pop > 0),
    blocks_with_zero_or_na = sum(is.na(pop) | pop == 0),
    total_population = sum(pop, na.rm = TRUE)
  )

# Step 7: View results
print(summary_table)

# all 15,166 blocks included in PWS and not in ATS have a population of zero or NA in the 2020 Census - this is great news, which means that the joining of the ATS and PWS HIs are joined properly and there are no issues with mismatching geoids, for example
```



#pull summary info on combined HIs - whether it is truly combined or if there's only ATS or PWS data (note: the N in these summaries is for block + endpoint combinations, except when broken by endpoint)
```{r}
# Overall summary table with sample size
SQL <- "
-- Summary for HI_tot_comb
SELECT 
  'HI_tot_comb' AS HI_type,
  COUNT(HI_tot_comb) AS n,
  MIN(HI_tot_comb) AS min,
  MAX(HI_tot_comb) AS max,
  AVG(HI_tot_comb) AS mean,
  APPROX_QUANTILES(HI_tot_comb, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot_comb) AS std_dev,
  STDDEV(HI_tot_comb) / SQRT(COUNT(HI_tot_comb)) AS std_error
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID

UNION ALL

-- Summary for HI_pt_comb
SELECT
  'HI_pt_comb' AS HI_type,
  COUNT(HI_pt_comb),
  MIN(HI_pt_comb),
  MAX(HI_pt_comb),
  AVG(HI_pt_comb),
  APPROX_QUANTILES(HI_pt_comb, 2)[OFFSET(1)],
  STDDEV(HI_pt_comb),
  STDDEV(HI_pt_comb) / SQRT(COUNT(HI_pt_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID

UNION ALL

-- Summary for HI_other_comb
SELECT
  'HI_other_comb' AS HI_type,
  COUNT(HI_other_comb),
  MIN(HI_other_comb),
  MAX(HI_other_comb),
  AVG(HI_other_comb),
  APPROX_QUANTILES(HI_other_comb, 2)[OFFSET(1)],
  STDDEV(HI_other_comb),
  STDDEV(HI_other_comb) / SQRT(COUNT(HI_other_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
"

chlor_HI_comb_summary <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_comb_summary,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI_combined_summary.csv",
  row.names = FALSE
)

# Population summary with sample size
SQL <- "
-- Summary for HI_tot_comb by pop
SELECT 
  pop,
  'HI_tot_comb' AS HI_type,
  COUNT(HI_tot_comb) AS n,
  MIN(HI_tot_comb) AS min,
  MAX(HI_tot_comb) AS max,
  AVG(HI_tot_comb) AS mean,
  APPROX_QUANTILES(HI_tot_comb, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot_comb) AS std_dev,
  STDDEV(HI_tot_comb) / SQRT(COUNT(HI_tot_comb)) AS std_error
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
GROUP BY pop

UNION ALL

-- Summary for HI_pt_comb by pop
SELECT 
  pop,
  'HI_pt_comb' AS HI_type,
  COUNT(HI_pt_comb),
  MIN(HI_pt_comb),
  MAX(HI_pt_comb),
  AVG(HI_pt_comb),
  APPROX_QUANTILES(HI_pt_comb, 2)[OFFSET(1)],
  STDDEV(HI_pt_comb),
  STDDEV(HI_pt_comb) / SQRT(COUNT(HI_pt_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
GROUP BY pop

UNION ALL

-- Summary for HI_other_comb by pop
SELECT 
  pop,
  'HI_other_comb' AS HI_type,
  COUNT(HI_other_comb),
  MIN(HI_other_comb),
  MAX(HI_other_comb),
  AVG(HI_other_comb),
  APPROX_QUANTILES(HI_other_comb, 2)[OFFSET(1)],
  STDDEV(HI_other_comb),
  STDDEV(HI_other_comb) / SQRT(COUNT(HI_other_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
GROUP BY pop
"

chlor_HI_comb_summary_by_pop <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_comb_summary_by_pop,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI_combined_summary_by_pop.csv",
  row.names = FALSE
)

# Population by endpoint summary with sample size
SQL <- "
-- Summary for HI_tot_comb by pop and endpoint
SELECT 
  pop,
  endpoint,
  'HI_tot_comb' AS HI_type,
  COUNT(HI_tot_comb) AS n,
  MIN(HI_tot_comb) AS min,
  MAX(HI_tot_comb) AS max,
  AVG(HI_tot_comb) AS mean,
  APPROX_QUANTILES(HI_tot_comb, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot_comb) AS std_dev,
  STDDEV(HI_tot_comb) / SQRT(COUNT(HI_tot_comb)) AS std_error
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
GROUP BY pop, endpoint

UNION ALL

-- Summary for HI_pt_comb by pop and endpoint
SELECT 
  pop,
  endpoint,
  'HI_pt_comb' AS HI_type,
  COUNT(HI_pt_comb),
  MIN(HI_pt_comb),
  MAX(HI_pt_comb),
  AVG(HI_pt_comb),
  APPROX_QUANTILES(HI_pt_comb, 2)[OFFSET(1)],
  STDDEV(HI_pt_comb),
  STDDEV(HI_pt_comb) / SQRT(COUNT(HI_pt_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
GROUP BY pop, endpoint

UNION ALL

-- Summary for HI_other_comb by pop and endpoint
SELECT 
  pop,
  endpoint,
  'HI_other_comb' AS HI_type,
  COUNT(HI_other_comb),
  MIN(HI_other_comb),
  MAX(HI_other_comb),
  AVG(HI_other_comb),
  APPROX_QUANTILES(HI_other_comb, 2)[OFFSET(1)],
  STDDEV(HI_other_comb),
  STDDEV(HI_other_comb) / SQRT(COUNT(HI_other_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
GROUP BY pop, endpoint
"

chlor_HI_comb_summary_by_pop_endpoint <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_comb_summary_by_pop_endpoint,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI_combined_summary_by_pop_endpoint.csv",
  row.names = FALSE
)


SQL <- "
SELECT COUNT(DISTINCT t.geoid20) AS n_blocks_in_filtblocks
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE t.data_source = 'both'
"

dbGetQuery(con, SQL)

```


#pull summary info on combined HIs for blocks only where there are truly combined HIs (note: the N in these summaries is for block + endpoint combinations, except when broken by endpoint)
```{r}
# Overall summary table with sample size
SQL <- "
-- Summary for HI_tot_comb
SELECT 
  'HI_tot_comb' AS HI_type,
  COUNT(HI_tot_comb) AS n,
  MIN(HI_tot_comb) AS min,
  MAX(HI_tot_comb) AS max,
  AVG(HI_tot_comb) AS mean,
  APPROX_QUANTILES(HI_tot_comb, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot_comb) AS std_dev,
  STDDEV(HI_tot_comb) / SQRT(COUNT(HI_tot_comb)) AS std_error
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'

UNION ALL

-- Summary for HI_pt_comb
SELECT
  'HI_pt_comb',
  COUNT(HI_pt_comb),
  MIN(HI_pt_comb),
  MAX(HI_pt_comb),
  AVG(HI_pt_comb),
  APPROX_QUANTILES(HI_pt_comb, 2)[OFFSET(1)],
  STDDEV(HI_pt_comb),
  STDDEV(HI_pt_comb) / SQRT(COUNT(HI_pt_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'

UNION ALL

-- Summary for HI_other_comb
SELECT
  'HI_other_comb',
  COUNT(HI_other_comb),
  MIN(HI_other_comb),
  MAX(HI_other_comb),
  AVG(HI_other_comb),
  APPROX_QUANTILES(HI_other_comb, 2)[OFFSET(1)],
  STDDEV(HI_other_comb),
  STDDEV(HI_other_comb) / SQRT(COUNT(HI_other_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'
"

chlor_HI_combonly_summary <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_combonly_summary,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI_combinedonly_summary.csv",
  row.names = FALSE
)

# Summary by population with sample size
SQL <- "
-- Summary for HI_tot_comb by pop
SELECT 
  pop,
  'HI_tot_comb' AS HI_type,
  COUNT(HI_tot_comb) AS n,
  MIN(HI_tot_comb) AS min,
  MAX(HI_tot_comb) AS max,
  AVG(HI_tot_comb) AS mean,
  APPROX_QUANTILES(HI_tot_comb, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot_comb) AS std_dev,
  STDDEV(HI_tot_comb) / SQRT(COUNT(HI_tot_comb)) AS std_error
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'
GROUP BY pop

UNION ALL

-- Summary for HI_pt_comb by pop
SELECT 
  pop,
  'HI_pt_comb',
  COUNT(HI_pt_comb),
  MIN(HI_pt_comb),
  MAX(HI_pt_comb),
  AVG(HI_pt_comb),
  APPROX_QUANTILES(HI_pt_comb, 2)[OFFSET(1)],
  STDDEV(HI_pt_comb),
  STDDEV(HI_pt_comb) / SQRT(COUNT(HI_pt_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'
GROUP BY pop

UNION ALL

-- Summary for HI_other_comb by pop
SELECT 
  pop,
  'HI_other_comb',
  COUNT(HI_other_comb),
  MIN(HI_other_comb),
  MAX(HI_other_comb),
  AVG(HI_other_comb),
  APPROX_QUANTILES(HI_other_comb, 2)[OFFSET(1)],
  STDDEV(HI_other_comb),
  STDDEV(HI_other_comb) / SQRT(COUNT(HI_other_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'
GROUP BY pop
"

chlor_HI_combonly_summary_by_pop <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_combonly_summary_by_pop,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI_combinedonly_summary_by_pop.csv",
  row.names = FALSE
)

# Summary by population and endpoint with sample size
SQL <- "
-- Summary for HI_tot_comb by pop and endpoint
SELECT 
  pop,
  endpoint,
  'HI_tot_comb' AS HI_type,
  COUNT(HI_tot_comb) AS n,
  MIN(HI_tot_comb) AS min,
  MAX(HI_tot_comb) AS max,
  AVG(HI_tot_comb) AS mean,
  APPROX_QUANTILES(HI_tot_comb, 2)[OFFSET(1)] AS median,
  STDDEV(HI_tot_comb) AS std_dev,
  STDDEV(HI_tot_comb) / SQRT(COUNT(HI_tot_comb)) AS std_error
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'
GROUP BY pop, endpoint

UNION ALL

-- Summary for HI_pt_comb by pop and endpoint
SELECT 
  pop,
  endpoint,
  'HI_pt_comb',
  COUNT(HI_pt_comb),
  MIN(HI_pt_comb),
  MAX(HI_pt_comb),
  AVG(HI_pt_comb),
  APPROX_QUANTILES(HI_pt_comb, 2)[OFFSET(1)],
  STDDEV(HI_pt_comb),
  STDDEV(HI_pt_comb) / SQRT(COUNT(HI_pt_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'
GROUP BY pop, endpoint

UNION ALL

-- Summary for HI_other_comb by pop and endpoint
SELECT 
  pop,
  endpoint,
  'HI_other_comb',
  COUNT(HI_other_comb),
  MIN(HI_other_comb),
  MAX(HI_other_comb),
  AVG(HI_other_comb),
  APPROX_QUANTILES(HI_other_comb, 2)[OFFSET(1)],
  STDDEV(HI_other_comb),
  STDDEV(HI_other_comb) / SQRT(COUNT(HI_other_comb))
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE data_source = 'both'
GROUP BY pop, endpoint
"

chlor_HI_combonly_summary_by_pop_endpoint <- dbGetQuery(con, SQL)

write.csv(
  chlor_HI_combonly_summary_by_pop_endpoint,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI_combinedonly_summary_by_pop_endpoint.csv",
  row.names = FALSE
)

```


#CDFs faceted by pop only for total HI
```{r}
# --- 1. Pull the data ---
project_id <- "edf-aq-data"

SQL <- "
SELECT
  t.geoid20,
  t.endpoint,
  t.pop,
  t.HI_tot_comb
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
"

chlor_combined_filtered <- bq_project_query(project_id, SQL) %>%
  bq_table_download()

# --- 2. Pre-compute CDF by population AND endpoint ---
chlor_cdf <- chlor_combined_filtered %>%
  group_by(pop, endpoint) %>%
  arrange(HI_tot_comb) %>%
  mutate(cdf = row_number() / n()) %>%
  ungroup()

# Optional: thin points (every 10th to speed up plotting)
chlor_cdf_sample <- chlor_cdf %>%
  group_by(pop, endpoint) %>%
  slice(seq(1, n(), by = 10)) %>%
  ungroup()

# --- 3. Plot ---
cdf_plot <- ggplot(
  chlor_cdf_sample,
  aes(x = HI_tot_comb, y = cdf, group = endpoint, color = endpoint)
) +
  geom_line(linewidth = 1) +
  # vertical reference lines for thresholds
  geom_vline(xintercept = c(0.1, 1), linetype = "dotted", color = "black") +
  scale_x_log10() +
  labs(
    title = "CDF of Combined HI_tot (chlor, Children vs Adults)",
    x = "HI_tot_comb (log scale)",
    y = "P(HI ≤ x)",
    color = "Endpoint"
  ) +
  facet_wrap(~pop, scales = "free_y") +
  theme_minimal(base_size = 14)

cdf_plot

ggsave(
  filename = "chlor_combined_cdf.png",
  device = "png",
  plot = cdf_plot,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/",
  width = 7,
  height = 5
)

```



#######################################

### analysis with HI threshold of 1 

#######################################



#count of blocks with various HIs above 1 
```{r}
SQL_counts <- "
WITH counts AS (
  SELECT
    t.data_source,
    t.pop,  -- adult or child

    -- Unique block counts
    COUNT(DISTINCT CASE WHEN t.HI_tot_comb   > 1 THEN t.geoid20 END) AS unique_blocks_above_1_tot,
    COUNT(DISTINCT t.geoid20) AS total_unique_blocks_tot,

    COUNT(DISTINCT CASE WHEN t.HI_pt_comb    > 1 THEN t.geoid20 END) AS unique_blocks_above_1_pt,
    COUNT(DISTINCT t.geoid20) AS total_unique_blocks_pt,

    COUNT(DISTINCT CASE WHEN t.HI_other_comb > 1 THEN t.geoid20 END) AS unique_blocks_above_1_other,
    COUNT(DISTINCT t.geoid20) AS total_unique_blocks_other,

    -- Row (block + endpoint combo) counts
    COUNT(CASE WHEN t.HI_tot_comb   > 1 THEN 1 END) AS total_exceedances_tot,
    COUNT(t.HI_tot_comb)   AS total_combos_tot,

    COUNT(CASE WHEN t.HI_pt_comb    > 1 THEN 1 END) AS total_exceedances_pt,
    COUNT(t.HI_pt_comb)    AS total_combos_pt,

    COUNT(CASE WHEN t.HI_other_comb > 1 THEN 1 END) AS total_exceedances_other,
    COUNT(t.HI_other_comb) AS total_combos_other

  FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = f.GEOID
  GROUP BY t.data_source, t.pop
)

-- Reshape into long format
SELECT data_source, pop, 'HI_tot_comb' AS HI_type,
  unique_blocks_above_1_tot AS unique_blocks_above_1,
  total_unique_blocks_tot   AS total_unique_blocks,
  total_exceedances_tot     AS total_exceedances,
  total_combos_tot          AS total_block_endpoint_combos
FROM counts

UNION ALL

SELECT data_source, pop, 'HI_pt_comb',
  unique_blocks_above_1_pt,
  total_unique_blocks_pt,
  total_exceedances_pt,
  total_combos_pt
FROM counts

UNION ALL

SELECT data_source, pop, 'HI_other_comb',
  unique_blocks_above_1_other,
  total_unique_blocks_other,
  total_exceedances_other,
  total_combos_other
FROM counts

ORDER BY data_source, pop, HI_type;
"

chlor_blocks_HIsabove1_by_sourcepop <- dbGetQuery(con, SQL_counts)

write.csv(
  chlor_blocks_HIsabove1_by_sourcepop,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/blocks_HIsabove1_combined_by_sourcepop.csv",
  row.names = FALSE
)


# Pull df of HIs > 1 for further analysis, filtered by filtblocks_chlor
SQL_exceedances <- "
SELECT t.*
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE t.HI_tot_comb > 1 OR t.HI_pt_comb > 1 OR t.HI_other_comb > 1;
"

chlor_exceedances <- dbGetQuery(con, SQL_exceedances)
```


# how many blocks exceed more than one health endpoint threshold (grouped by block, data source, and population): 1 PWS and multiple ATS blocks exceeded multiple endpoints; all other mult endpoint exceedances are from both sources
```{r}
blocks_mult_endpoints <- chlor_exceedances %>%
  group_by(geoid20, data_source, pop) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(unique(endpoint), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints <- chlor_exceedances %>%
  semi_join(blocks_mult_endpoints, by = c("geoid20", "data_source", "pop")) %>%
  left_join(blocks_mult_endpoints, by = c("geoid20", "data_source", "pop"))

write.csv(blocks_mult_endpoints, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/blocks_mult_endpoints_by_source_pop.csv", 
          row.names = FALSE)

```


#number of blocks with combined HIs that exceeds threshold that didn't exceed from just air or water (note: these numbers are for block + endpoint combinations)
```{r}
get_hi_exceedance_summary <- function(df, hi_type, ats_col, pws_col, comb_col) {
  df %>%
    filter(.data[[comb_col]] > 1) %>%
    mutate(
      ats_below_1 = is.na(.data[[ats_col]]) | .data[[ats_col]] < 1,
      pws_below_1 = is.na(.data[[pws_col]]) | .data[[pws_col]] < 1,
      both_below_1 = ats_below_1 & pws_below_1,
      either_below_1 = ats_below_1 | pws_below_1,
      below_which = case_when(
        both_below_1 ~ "ats, pws",
        ats_below_1 ~ "ats",
        pws_below_1 ~ "pws",
        TRUE ~ "neither"
      ),
      exceeded_which = paste0(
        ifelse(!is.na(.data[[ats_col]]) & .data[[ats_col]] > 1, ats_col, ""),
        ifelse(!is.na(.data[[ats_col]]) & .data[[ats_col]] > 1 & !is.na(.data[[pws_col]]) & .data[[pws_col]] > 1, ", ", ""),
        ifelse(!is.na(.data[[pws_col]]) & .data[[pws_col]] > 1, "pws_HI", "")
      ),
      HI_type = hi_type
    ) %>%
    select(geoid20, pop, HI_type, !!comb_col, !!ats_col, !!pws_col,
           ats_below_1, pws_below_1, both_below_1, either_below_1,
           below_which, exceeded_which)
}

df_tot <- get_hi_exceedance_summary(chlor_exceedances, "HI_tot_comb", "ats_HI_tot", "pws_HI", "HI_tot_comb")
df_pt <- get_hi_exceedance_summary(chlor_exceedances, "HI_pt_comb", "ats_HI_pt", "pws_HI", "HI_pt_comb")
df_other <- get_hi_exceedance_summary(chlor_exceedances, "HI_other_comb", "ats_HI_other", "pws_HI", "HI_other_comb")

all_HI_exceedances <- bind_rows(df_tot, df_pt, df_other)

summary_table <- all_HI_exceedances %>%
  group_by(HI_type, pop) %>%
  summarise(
    total_comb_HI_over_1 = n(),
    count_both_below_1 = sum(both_below_1),
    count_ats_below_1_only = sum(ats_below_1 & !pws_below_1),
    count_pws_below_1_only = sum(pws_below_1 & !ats_below_1),
    .groups = "drop"
  )

write.csv(summary_table, "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/blocks_combined_HI_summary_by_pop.csv", row.names = FALSE)

```


#graph on fractional contributions of air vs. water for exceedances (note: these numbers are for block + endpoint combinations)
```{r}
combined_contrib <- chlor_exceedances %>%
  filter(data_source == "both") %>%
  select(pop, ats_HI_tot, ats_HI_pt, pws_HI, HI_tot_comb, HI_pt_comb) %>%
  pivot_longer(
    cols = c(ats_HI_tot, ats_HI_pt),
    names_to = "HI_type",
    names_prefix = "ats_HI_",
    values_to = "ats_HI"
  ) %>%
  mutate(
    HI_comb = case_when(
      HI_type == "tot" ~ HI_tot_comb,
      HI_type == "pt" ~ HI_pt_comb
    ),
    ats_fraction = ats_HI / HI_comb,
    pws_fraction = pws_HI / HI_comb
  ) %>%
  filter(HI_comb > 0) %>%
  group_by(pop, HI_type) %>%
  summarise(
    mean_ats = mean(ats_fraction, na.rm = TRUE),
    sd_ats = sd(ats_fraction, na.rm = TRUE),
    mean_pws = mean(pws_fraction, na.rm = TRUE),
    sd_pws = sd(pws_fraction, na.rm = TRUE),
    n_blocks = n(),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(mean_ats, mean_pws, sd_ats, sd_pws),
               names_to = c("stat", "source"),
               names_sep = "_",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(
    source = ifelse(source == "ats", "ATS", "PWS"),
    HI_type = dplyr::recode(as.character(HI_type),
                            "tot" = "HI_tot_comb",
                            "pt"  = "HI_pt_comb")
  )

write.csv(combined_contrib, 
          "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/ats_pws_frac_contributions.csv", 
          row.names = FALSE)

```


#graph on just raw mean and/or median of air vs. water HIs for exceedances where both contirubted to the HI (note: these numbers are for block + endpoint combinations)
```{r}
# Prepare long-format data for ATS and PWS HI values by type (exclude "other")
combined_raw_HIs <- chlor_exceedances %>%
  filter(data_source == "both") %>%
  select(pop, 
         ats_HI_tot, ats_HI_pt,
         pws_HI) %>%   # removed ats_HI_other
  pivot_longer(
    cols = c(ats_HI_tot, ats_HI_pt),   # only tot and pt
    names_to = "HI_type",
    names_prefix = "ats_HI_",
    values_to = "ats_HI"
  ) %>%
  mutate(
    HI_type_full = dplyr::recode(HI_type,
                                 "tot" = "HI_tot_comb",
                                 "pt"  = "HI_pt_comb")
  ) %>%
  group_by(pop, HI_type_full) %>%
  summarise(
    mean_ats = mean(ats_HI, na.rm = TRUE),
    median_ats = median(ats_HI, na.rm = TRUE),
    sd_ats = sd(ats_HI, na.rm = TRUE),
    
    mean_pws = mean(pws_HI, na.rm = TRUE),
    median_pws = median(pws_HI, na.rm = TRUE),
    sd_pws = sd(pws_HI, na.rm = TRUE),
    
    n_blocks = n(),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(mean_ats, median_ats, sd_ats, mean_pws, median_pws, sd_pws),
    names_to = c("stat", "source"),
    names_sep = "_",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(
    source = ifelse(source == "ats", "ATS", "PWS"),
    HI_type = HI_type_full
  ) %>%
  select(-HI_type_full)

write.csv(combined_raw_HIs, "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/ats_pws_raw_HIs.csv", row.names = FALSE)

# Plot only tot and pt
ats_pws_raw = ggplot(combined_raw_HIs, aes(x = HI_type, y = mean, fill = source)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2,
    color = "black"
  ) +
  facet_wrap(~pop, labeller = labeller(pop = c("ad" = "Adults", "ch" = "Children"))) +
  scale_fill_manual(values = c("ATS" = "#1f77b4", "PWS" = "#ff7f0e")) +
  labs(
    title = "Average HI for Blocks with HI > 1 & Both Sources",
    x = "HI Type",
    y = "Avg HI",
    fill = "Data Source"
  ) +
  scale_x_discrete(labels = c(
    "HI_pt_comb"  = "Point Only",
    "HI_tot_comb" = "Total"
  )) +
  theme_minimal(base_size = 14)

ats_pws_raw

ggsave(filename = "bar_ats_pws_avgHI.png", device = "png", plot = ats_pws_raw, 
       path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/", width = 7, height = 5)


```


##alternative/clean - box plot of air vs. water HIs for exceedances where both contirubted to the HI (note: these numbers are for block + endpoint combinations)
```{r}
#--------------------------------------------------
# Create raw block-level dataset
#--------------------------------------------------

combined_raw_HIs_dist <- chlor_exceedances %>%
  filter(data_source == "both") %>%
  select(
    pop,
    ats_HI_tot,
    ats_HI_pt,
    pws_HI
  ) %>%
  pivot_longer(
    cols = c(ats_HI_tot, ats_HI_pt),
    names_to = "HI_type",
    names_prefix = "ats_HI_",
    values_to = "ats_HI"
  ) %>%
  mutate(
    HI_type = dplyr::recode(
      HI_type,
      "tot" = "HI_tot_comb",
      "pt"  = "HI_pt_comb"
    )
  ) %>%
  pivot_longer(
    cols = c(ats_HI, pws_HI),
    names_to = "source",
    values_to = "HI"
  ) %>%
  mutate(
    source = dplyr::recode(
      source,
      ats_HI = "Air",
      pws_HI = "Drinking Water"
    )
  )

#--------------------------------------------------
# Boxplot
#--------------------------------------------------

ats_pws_box <- ggplot(
  combined_raw_HIs_dist,
  aes(
    x = HI_type,
    y = HI,
    fill = source
  )
) +

  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.6,
    alpha = 0.6,
    outlier.shape = NA
  ) +

  geom_jitter(
    aes(shape = source),
    position = position_jitterdodge(
      jitter.width = 0.15,
      dodge.width = 0.8
    ),
    size = 1.5,
    alpha = 0.5
  ) +

  facet_wrap(
    ~pop,
    labeller = labeller(
      pop = c(
        "ad" = "Adults",
        "ch" = "Children"
      )
    )
  ) +

  scale_fill_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    )
  ) +

  scale_shape_manual(
    values = c(
      "Air" = 16,
      "Drinking Water" = 18
    )
  ) +

  scale_x_discrete(
    labels = c(
      "HI_pt_comb" = "TSCA Sources",
      "HI_tot_comb" = "All Sources"
    )
  ) +

  labs(
    x = "Air Exposure Type",
    y = "Hazard Index (HI)",
    fill = "Exposure Source",
    shape = "Exposure Source"
  ) +

  theme_minimal(base_size = 14)

ats_pws_box

ggsave(
  filename = "boxplot_ats_pws_HI.png",
  plot = ats_pws_box,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/",
  width = 7,
  height = 5,
  dpi = 600
)

# save for downstream paneling

chlor_ats_pws_box_HI1 <- ats_pws_box

```



##alternative/clean - beeswarm
```{r}
chlor_ats_pws_beeswarm_HI1 <- ggplot(
  combined_raw_HIs_dist,
  aes(
    x = HI_type,
    y = HI,
    fill = source
  )
) +

  geom_quasirandom(
    aes(color = source),
    dodge.width = 0.8,
    alpha = 0.9,
    size = 1.5
  ) +

  stat_summary(
    aes(group = source),
    fun = median,
    geom = "crossbar",
    position = position_dodge(width = 0.8),
    width = 0.7,
    color = "black",
    linewidth = 0.4
  ) +

  facet_wrap(
    ~pop,
    labeller = labeller(
      pop = c(
        "ad" = "Adults",
        "ch" = "Children"
      )
    )
  ) +

  scale_fill_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    )
  ) +

  scale_color_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    ),
    guide = "none"
  ) +

  scale_x_discrete(
    labels = c(
      "HI_pt_comb"  = "TSCA Sources",
      "HI_tot_comb" = "All Sources"
    )
  ) +

  labs(
    x = "Air Exposure Type",
    y = "Hazard Index (HI)",
    fill = "Exposure Source"
  ) +

  theme_minimal(base_size = 14)

chlor_ats_pws_beeswarm_HI1


ggsave(
  filename = "beeswarm_chlor_ats_pws_HI.png",
  plot = chlor_ats_pws_beeswarm_HI1,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/",
  width = 7,
  height = 5,
  dpi = 600
)


chlor_ats_pws_beeswarm_HI1_obj <- chlor_ats_pws_beeswarm_HI1
```


#population counts for exceedances
```{r}
pop_summary_by_source_unique <- chlor_exceedances %>%
  pivot_longer(
    cols = c(HI_tot_comb, HI_pt_comb, HI_other_comb),
    names_to = "HI_type",
    values_to = "HI_value"
  ) %>%
  filter(HI_value > 1) %>%
  # keep only unique blocks per pop + HI_type + data_source
  distinct(geoid20, pop, HI_type, data_source, population, .keep_all = TRUE) %>%
  group_by(pop, HI_type, data_source) %>%
  summarise(
    total_pop = sum(population, na.rm = TRUE),
    avg_pop   = mean(population, na.rm = TRUE),
    n_blocks  = n_distinct(geoid20),
    .groups   = "drop"
  )

write.csv(
  pop_summary_by_source_unique,
  "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI1/population_summary_unique_blocks.csv",
  row.names = FALSE
)
```



#pull in df of blocks that have both PWS and ATS sources and match the TRI filter, and also have any HI above 1
```{r}
project_id <- "edf-aq-data"


SQL <- "
WITH blocks_with_hi_above1 AS (
  SELECT DISTINCT geoid20
  FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = f.GEOID
  WHERE t.block_has_both_sources = TRUE
    AND (t.HI_tot_comb > 1 OR t.HI_pt_comb > 1 OR t.HI_other_comb > 1)
)

SELECT t.*
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `blocks_with_hi_above1` b
  ON t.geoid20 = b.geoid20
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
"

chlor_HI_any_above1 <- bq_project_query(project_id, SQL) %>%
  bq_table_download()
```



#which blocks had more than 1 endpoint exceeding an HI of 1, and what were the sources? were there any that had HI exceedances from more than one source (both ATS and PWS?) --> answer: blocks that had multiple exceedances had both sources contributing to that endpoint's HI exceedances; for example, there were no blocks that exceeded for hematological from ATS and exceeded endocrine from PWS
```{r}
block_hi_long <- chlor_HI_any_above1 %>%
  pivot_longer(
    cols = c(HI_tot_comb, HI_pt_comb, HI_other_comb),
    names_to = "hi_type",
    values_to = "hi_value"
  ) %>%
  filter(hi_value > 1) %>%
  select(geoid20, endpoint, data_source, pop, hi_type, hi_value) %>%
  arrange(geoid20, endpoint, pop, hi_type)
```




#######################################

### analysis with HI threshold of 0.1 

#######################################





#count of blocks with various HIs above 0.1 
```{r}
SQL_counts <- "
WITH counts AS (
  SELECT
    t.data_source,
    t.pop,  -- adult or child

    -- Unique block counts
    COUNT(DISTINCT CASE WHEN t.HI_tot_comb   > 0.1 THEN t.geoid20 END) AS unique_blocks_above_01_tot,
    COUNT(DISTINCT t.geoid20) AS total_unique_blocks_tot,

    COUNT(DISTINCT CASE WHEN t.HI_pt_comb    > 0.1 THEN t.geoid20 END) AS unique_blocks_above_01_pt,
    COUNT(DISTINCT t.geoid20) AS total_unique_blocks_pt,

    COUNT(DISTINCT CASE WHEN t.HI_other_comb > 0.1 THEN t.geoid20 END) AS unique_blocks_above_01_other,
    COUNT(DISTINCT t.geoid20) AS total_unique_blocks_other,

    -- Row (block + endpoint combo) counts
    COUNT(CASE WHEN t.HI_tot_comb   > 0.1 THEN 1 END) AS total_exceedances_tot,
    COUNT(t.HI_tot_comb)   AS total_combos_tot,

    COUNT(CASE WHEN t.HI_pt_comb    > 0.1 THEN 1 END) AS total_exceedances_pt,
    COUNT(t.HI_pt_comb)    AS total_combos_pt,

    COUNT(CASE WHEN t.HI_other_comb > 0.1 THEN 1 END) AS total_exceedances_other,
    COUNT(t.HI_other_comb) AS total_combos_other

  FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = f.GEOID
  GROUP BY t.data_source, t.pop
)

-- Reshape into long format
SELECT data_source, pop, 'HI_tot_comb' AS HI_type,
  unique_blocks_above_01_tot AS unique_blocks_above_01,
  total_unique_blocks_tot    AS total_unique_blocks,
  total_exceedances_tot      AS total_exceedances,
  total_combos_tot           AS total_block_endpoint_combos
FROM counts

UNION ALL

SELECT data_source, pop, 'HI_pt_comb',
  unique_blocks_above_01_pt,
  total_unique_blocks_pt,
  total_exceedances_pt,
  total_combos_pt
FROM counts

UNION ALL

SELECT data_source, pop, 'HI_other_comb',
  unique_blocks_above_01_other,
  total_unique_blocks_other,
  total_exceedances_other,
  total_combos_other
FROM counts

ORDER BY data_source, pop, HI_type;
"

chlor_blocks_HIsabove01_by_sourcepop <- dbGetQuery(con, SQL_counts)

write.csv(
  chlor_blocks_HIsabove01_by_sourcepop,
  file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/blocks_HIsabove01_combined_by_sourcepop.csv",
  row.names = FALSE
)


# Pull df of HIs > 0.1 for further analysis, filtered by filtblocks_chlor
SQL_exceedances <- "
SELECT t.*
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
WHERE t.HI_tot_comb > 0.1 OR t.HI_pt_comb > 0.1 OR t.HI_other_comb > 0.1;
"

chlor_exceedances <- dbGetQuery(con, SQL_exceedances)
```


# how many blocks exceed more than one health endpoint threshold (grouped by block, data source, and population): 1 PWS and multiple ATS blocks exceeded multiple endpoints; all other mult endpoint exceedances are from both sources
```{r}
blocks_mult_endpoints <- chlor_exceedances %>%
  group_by(geoid20, data_source, pop) %>%
  summarise(
    n_endpoints_exceeding = n_distinct(endpoint),
    endpoints = paste(unique(endpoint), collapse = ", "),
    .groups = "drop"
  ) %>%
  filter(n_endpoints_exceeding > 1)

chlor_mult_endpoints <- chlor_exceedances %>%
  semi_join(blocks_mult_endpoints, by = c("geoid20", "data_source", "pop")) %>%
  left_join(blocks_mult_endpoints, by = c("geoid20", "data_source", "pop"))

write.csv(blocks_mult_endpoints, 
          file = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/blocks_mult_endpoints_by_source_pop.csv", 
          row.names = FALSE)

```


#number of blocks with combined HIs that exceeds threshold that didn't exceed from just air or water (note: these numbers are for block + endpoint combinations)
```{r}
get_hi_exceedance_summary <- function(df, hi_type, ats_col, pws_col, comb_col) {
  df %>%
    filter(.data[[comb_col]] > 0.1) %>%
    mutate(
      ats_below_01 = is.na(.data[[ats_col]]) | .data[[ats_col]] < 0.1,
      pws_below_01 = is.na(.data[[pws_col]]) | .data[[pws_col]] < 0.1,
      both_below_01 = ats_below_01 & pws_below_01,
      either_below_01 = ats_below_01 | pws_below_01,
      below_which = case_when(
        both_below_01 ~ "ats, pws",
        ats_below_01 ~ "ats",
        pws_below_01 ~ "pws",
        TRUE ~ "neither"
      ),
      exceeded_which = paste0(
        ifelse(!is.na(.data[[ats_col]]) & .data[[ats_col]] > 0.1, ats_col, ""),
        ifelse(!is.na(.data[[ats_col]]) & .data[[ats_col]] > 0.1 & !is.na(.data[[pws_col]]) & .data[[pws_col]] > 0.1, ", ", ""),
        ifelse(!is.na(.data[[pws_col]]) & .data[[pws_col]] > 0.1, "pws_HI", "")
      ),
      HI_type = hi_type
    ) %>%
    select(geoid20, pop, HI_type, !!comb_col, !!ats_col, !!pws_col,
           ats_below_01, pws_below_01, both_below_01, either_below_01,
           below_which, exceeded_which)
}

df_tot <- get_hi_exceedance_summary(chlor_exceedances, "HI_tot_comb", "ats_HI_tot", "pws_HI", "HI_tot_comb")
df_pt <- get_hi_exceedance_summary(chlor_exceedances, "HI_pt_comb", "ats_HI_pt", "pws_HI", "HI_pt_comb")
df_other <- get_hi_exceedance_summary(chlor_exceedances, "HI_other_comb", "ats_HI_other", "pws_HI", "HI_other_comb")

all_HI_exceedances <- bind_rows(df_tot, df_pt, df_other)

summary_table <- all_HI_exceedances %>%
  group_by(HI_type, pop) %>%
  summarise(
    total_comb_HI_over_01 = n(),
    count_both_below_01 = sum(both_below_01),
    count_ats_below_01_only = sum(ats_below_01 & !pws_below_01),
    count_pws_below_01_only = sum(pws_below_01 & !ats_below_01),
    .groups = "drop"
  )

write.csv(summary_table, "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/blocks_combined_HI_summary_by_pop.csv", row.names = FALSE)

```


#fractional contributions of air vs. water for exceedances (note: these numbers are for block + endpoint combinations)
```{r}
combined_contrib <- chlor_exceedances %>%
  filter(data_source == "both") %>%
  select(pop, ats_HI_tot, ats_HI_pt, pws_HI, HI_tot_comb, HI_pt_comb) %>%
  pivot_longer(
    cols = c(ats_HI_tot, ats_HI_pt),
    names_to = "HI_type",
    names_prefix = "ats_HI_",
    values_to = "ats_HI"
  ) %>%
  mutate(
    HI_comb = case_when(
      HI_type == "tot" ~ HI_tot_comb,
      HI_type == "pt" ~ HI_pt_comb
    ),
    ats_fraction = ats_HI / HI_comb,
    pws_fraction = pws_HI / HI_comb
  ) %>%
  filter(HI_comb > 0) %>%
  group_by(pop, HI_type) %>%
  summarise(
    mean_ats = mean(ats_fraction, na.rm = TRUE),
    sd_ats = sd(ats_fraction, na.rm = TRUE),
    mean_pws = mean(pws_fraction, na.rm = TRUE),
    sd_pws = sd(pws_fraction, na.rm = TRUE),
    n_blocks = n(),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(mean_ats, mean_pws, sd_ats, sd_pws),
               names_to = c("stat", "source"),
               names_sep = "_",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(
    source = ifelse(source == "ats", "ATS", "PWS"),
    HI_type = dplyr::recode(as.character(HI_type),
                            "tot" = "HI_tot_comb",
                            "pt"  = "HI_pt_comb")
  )

write.csv(combined_contrib, 
          "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/ats_pws_frac_contributions.csv", 
          row.names = FALSE)

```


#graph on just raw mean and/or median of air vs. water HIs for exceedances where both contirubted to the HI (note: these numbers are for block + endpoint combinations)
```{r}
# Prepare long-format data for ATS and PWS HI values by type (exclude "other")
combined_raw_HIs <- chlor_exceedances %>%
  filter(data_source == "both") %>%
  select(pop, 
         ats_HI_tot, ats_HI_pt,
         pws_HI) %>%   # removed ats_HI_other
  pivot_longer(
    cols = c(ats_HI_tot, ats_HI_pt),   # only tot and pt
    names_to = "HI_type",
    names_prefix = "ats_HI_",
    values_to = "ats_HI"
  ) %>%
  mutate(
    HI_type_full = dplyr::recode(HI_type,
                                 "tot" = "HI_tot_comb",
                                 "pt"  = "HI_pt_comb")
  ) %>%
  group_by(pop, HI_type_full) %>%
  summarise(
    mean_ats = mean(ats_HI, na.rm = TRUE),
    median_ats = median(ats_HI, na.rm = TRUE),
    sd_ats = sd(ats_HI, na.rm = TRUE),
    
    mean_pws = mean(pws_HI, na.rm = TRUE),
    median_pws = median(pws_HI, na.rm = TRUE),
    sd_pws = sd(pws_HI, na.rm = TRUE),
    
    n_blocks = n(),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(mean_ats, median_ats, sd_ats, mean_pws, median_pws, sd_pws),
    names_to = c("stat", "source"),
    names_sep = "_",
    values_to = "value"
  ) %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(
    source = ifelse(source == "ats", "ATS", "PWS"),
    HI_type = HI_type_full
  ) %>%
  select(-HI_type_full)

write.csv(combined_raw_HIs, "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/ats_pws_raw_HIs.csv", row.names = FALSE)

# Plot only tot and pt
ats_pws_raw = ggplot(combined_raw_HIs, aes(x = HI_type, y = mean, fill = source)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd),
    position = position_dodge(width = 0.9),
    width = 0.2,
    color = "black"
  ) +
  facet_wrap(~pop, labeller = labeller(pop = c("ad" = "Adults", "ch" = "Children"))) +
  scale_fill_manual(values = c("ATS" = "#1f77b4", "PWS" = "#ff7f0e")) +
  labs(
    title = "Average HI for Blocks with HI > 0.1 & Both Sources",
    x = "HI Type",
    y = "Avg HI",
    fill = "Data Source"
  ) +
  scale_x_discrete(labels = c(
    "HI_pt_comb"  = "Point Only",
    "HI_tot_comb" = "Total"
  )) +
  theme_minimal(base_size = 14)

ats_pws_raw

ggsave(filename = "bar_ats_pws_avgHI.png", device = "png", plot = ats_pws_raw, 
       path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/", width = 7, height = 5)


```


##alternative/clean - box plot of air vs. water HIs for exceedances where both contirubted to the HI (note: these numbers are for block + endpoint combinations) - going to use the boxplot!
```{r}
#--------------------------------------------------
# Prepare long-format HI values 
#--------------------------------------------------

combined_raw_HIs <- chlor_exceedances %>%
  filter(data_source == "both") %>%
  select(
    pop,
    ats_HI_tot, ats_HI_pt,
    pws_HI
  ) %>%
  pivot_longer(
    cols = c(ats_HI_tot, ats_HI_pt),
    names_to = "HI_type",
    names_prefix = "ats_HI_",
    values_to = "ats_HI"
  ) %>%
  mutate(
    HI_type = dplyr::recode(
      HI_type,
      "tot" = "HI_tot_comb",
      "pt"  = "HI_pt_comb"
    )
  ) %>%
  pivot_longer(
    cols = c(ats_HI, pws_HI),
    names_to = "source",
    values_to = "HI"
  ) %>%
  mutate(
    source = dplyr::recode(
      source,
      ats_HI = "Air",
      pws_HI = "Drinking Water"
    )
  )

#--------------------------------------------------
# Save raw data
#--------------------------------------------------

write.csv(
  combined_raw_HIs,
  "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/ats_pws_raw_HIs.csv",
  row.names = FALSE
)

#--------------------------------------------------
# Box plot 
#--------------------------------------------------

chlor_ats_pws_box_HI01 <- ggplot(
  combined_raw_HIs,
  aes(x = HI_type, y = HI, fill = source)
) +

  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.6,
    alpha = 0.6,
    outlier.shape = NA
  ) +

  facet_wrap(
    ~pop,
    labeller = labeller(
      pop = c(
        "ad" = "Adults",
        "ch" = "Children"
      )
    )
  ) +

  scale_fill_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    )
  ) +

  scale_x_discrete(
    labels = c(
      "HI_pt_comb" = "TSCA Sources",
      "HI_tot_comb" = "All Sources"
    )
  ) +

  coord_cartesian(ylim = c(0, 1)) +

  labs(
    x = "Air Exposure Type",
    y = "Hazard Index (HI)",
    fill = "Exposure Source"
  ) +

  theme_minimal(base_size = 14)

chlor_ats_pws_box_HI01

#--------------------------------------------------
# Save plot
#--------------------------------------------------

ggsave(
  filename = "boxplot_chlor_ats_pws_HI.png",
  plot = chlor_ats_pws_box_HI01,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/",
  width = 7,
  height = 5,
  dpi = 600
)


#--------------------------------------------------
# Sample sizes
#--------------------------------------------------

combined_raw_HIs %>%
  group_by(pop, HI_type, source) %>%
  summarise(
    n = n(),
    .groups = "drop"
  )
```


#beeswarm plot (using this paneled with HI 1 plot)
```{r}
chlor_ats_pws_beeswarm_median_HI01 <- ggplot(
  combined_raw_HIs,
  aes(
    x = HI_type,
    y = HI,
    fill = source
  )
) +

  geom_quasirandom(
    aes(color = source),
    dodge.width = 0.8,
    alpha = 0.15,
    size = 0.5
  ) +

  stat_summary(
    aes(group = source),
    fun = median,
    geom = "crossbar",
    position = position_dodge(width = 0.8),
    width = 0.7,
    color = "black",
    linewidth = 0.4
  ) +

  facet_wrap(
    ~pop,
    labeller = labeller(
      pop = c(
        "ad" = "Adults",
        "ch" = "Children"
      )
    )
  ) +

  scale_fill_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    )
  ) +

  scale_color_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    ),
    guide = "none"
  ) +
  
  coord_cartesian(ylim = c(0, 1.5)) +
  
  scale_x_discrete(
    labels = c(
      "HI_pt_comb"  = "TSCA Sources",
      "HI_tot_comb" = "All Sources"
    )
  ) +

  labs(
    x = "Air Exposure Type",
    y = "Hazard Index (HI)",
    fill = "Exposure Source"
  ) +

  theme_minimal(base_size = 14)

chlor_ats_pws_beeswarm_median_HI01


ggsave(
  filename = "beeswarm_chlor_ats_pws_HI.png",
  plot = chlor_ats_pws_beeswarm_median_HI01,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/",
  width = 7,
  height = 5,
  dpi = 600
)

```


#trying different versions of boxplots - boxplot with 5% sample of all observations (saving this to present in presentation as another example of how to show this data)
```{r}
set.seed(123)

combined_raw_HIs_sample <- combined_raw_HIs %>%
  group_by(pop, HI_type, source) %>%
  slice_sample(prop = 0.05) %>%
  ungroup()

chlor_ats_pws_box_sample_HI01 <- ggplot(
  combined_raw_HIs,
  aes(x = HI_type, y = HI, fill = source)
) +

  geom_point(
    data = combined_raw_HIs_sample,
    aes(color = source),
    position = position_jitterdodge(
      jitter.width = 0.15,
      dodge.width = 0.8
    ),
    alpha = 0.4,
    size = 0.8
  ) +

  geom_boxplot(
    position = position_dodge(width = 0.8),
    width = 0.6,
    alpha = 0.6,
    outlier.alpha = 0.25,
    outlier.size = 0.8
  ) +

  facet_wrap(
    ~pop,
    labeller = labeller(
      pop = c(
        "ad" = "Adults",
        "ch" = "Children"
      )
    )
  ) +

  scale_fill_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    )
  ) +

  scale_color_manual(
    values = c(
      "Air" = "#0072B2",
      "Drinking Water" = "#E69F00"
    ),
    guide = "none"
  ) +

  scale_x_discrete(
    labels = c(
      "HI_pt_comb" = "TSCA Sources",
      "HI_tot_comb" = "All Sources"
    )
  ) +

  labs(
    x = "Air Exposure Type",
    y = "Hazard Index (HI)",
    fill = "Exposure Source"
  ) +

  theme_minimal(base_size = 14)

chlor_ats_pws_box_sample_HI01

ggsave(
  filename = "boxplot5_chlor_ats_pws_HI.png",
  plot = chlor_ats_pws_box_sample_HI01,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/",
  width = 7,
  height = 5,
  dpi = 600
)

```



#population counts for exceedances
```{r}
pop_summary_by_source_unique <- chlor_exceedances %>%
  pivot_longer(
    cols = c(HI_tot_comb, HI_pt_comb, HI_other_comb),
    names_to = "HI_type",
    values_to = "HI_value"
  ) %>%
  filter(HI_value > 0.1) %>%
  # keep only unique blocks per pop + HI_type + data_source
  distinct(geoid20, pop, HI_type, data_source, population, .keep_all = TRUE) %>%
  group_by(pop, HI_type, data_source) %>%
  summarise(
    total_pop = sum(population, na.rm = TRUE),
    avg_pop   = mean(population, na.rm = TRUE),
    n_blocks  = n_distinct(geoid20),
    .groups   = "drop"
  )

write.csv(
  pop_summary_by_source_unique,
  "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/HI01/population_summary_unique_blocks.csv",
  row.names = FALSE
)
```



#pull in df of blocks that have both PWS and ATS sources and match the TRI filter, and also have any HI above 0.1
```{r}
project_id <- "edf-aq-data"


SQL <- "
WITH blocks_with_hi_above01 AS (
  SELECT DISTINCT geoid20
  FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
  JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
    ON t.geoid20 = f.GEOID
  WHERE t.block_has_both_sources = TRUE
    AND (t.HI_tot_comb > 0.1 OR t.HI_pt_comb > 0.1 OR t.HI_other_comb > 0.1)
)

SELECT t.*
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `blocks_with_hi_above01` b
  ON t.geoid20 = b.geoid20
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
"

chlor_HI_any_above01 <- bq_project_query(project_id, SQL) %>%
  bq_table_download()
```



#which blocks had more than 1 endpoint exceeding an HI of 0.1, and what were the sources? were there any that had HI exceedances from more than one source (both ATS and PWS?) --> answer: blocks that had multiple exceedances had both sources contributing to that endpoint's HI exceedances; for example, there were no blocks that exceeded for hematological from ATS and exceeded endocrine from PWS

#occular, respiratory, and cardiological had 'ats_only' source but that is because there's no RfDs for DW exposure for these endpoints
```{r}
block_hi_long <- chlor_HI_any_above01 %>%
  pivot_longer(
    cols = c(HI_tot_comb, HI_pt_comb, HI_other_comb),
    names_to = "hi_type",
    values_to = "hi_value"
  ) %>%
  filter(hi_value > 0.1) %>%
  select(geoid20, endpoint, data_source, pop, hi_type, hi_value) %>%
  arrange(geoid20, endpoint, pop, hi_type)
```





### panel together beeswarm plots for manuscript ###

```{r}
chlor_ats_pws_beeswarm_median_HI1 <- chlor_ats_pws_beeswarm_HI1 +
  guides(
    color = "none",
    fill = guide_legend()
  )

chlor_ats_pws_beeswarm_HI01 <- chlor_ats_pws_beeswarm_median_HI01 + 
  guides(
    color = "none",
    fill = "none"
  )


combined_panel <- chlor_ats_pws_beeswarm_median_HI1 +
  chlor_ats_pws_beeswarm_HI01 +
  plot_layout(ncol = 2, guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.justification = "center"
  )

combined_panel


ggsave(
  filename = "beeswarm_chlor_combined_HI.png",
  plot = combined_panel,
  path = "~/TSCA/CRA/metdb_tsca/output/chlor/Combined/",
  width = 11,
  height = 5,
  dpi = 600
)
```



### combined CDFs that show air, water, and combined data for each endpoint ###



#load data needed
```{r}
project_id <- "edf-aq-data"

# air
air_sql <- "
SELECT
  t.block AS geoid,
  t.endpoint,
  t.HI_tot,
  t.HI_pt
FROM `edf-aq-data.TSCA.chlor_ATS_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.block = f.GEOID
"

air_df <- bq_project_query(project_id, air_sql) %>%
  bq_table_download()

# water
water_sql <- "
SELECT
  t.geoid20 AS geoid,
  t.endpoint,
  t.pop,
  t.HI
FROM `edf-aq-data.TSCA.chlor_PWS_HIs_half` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
"

water_df <- bq_project_query(project_id, water_sql) %>%
  bq_table_download()

# combined
combined_sql <- "
SELECT
  t.geoid20 AS geoid,
  t.endpoint,
  t.pop,
  t.HI_tot_comb,
  t.HI_pt_comb
FROM `edf-aq-data.TSCA.chlor_combined_HIs` t
JOIN `edf-aq-data.TSCA.filtblocks_chlor` f
  ON t.geoid20 = f.GEOID
"

combined_df <- bq_project_query(project_id, combined_sql) %>%
  bq_table_download()
```


#convert data to long format and combine
```{r}
# air
air_long <- air_df %>%
  pivot_longer(
    cols = c(HI_tot, HI_pt),
    names_to = "scenario",
    values_to = "HI"
  ) %>%
  mutate(
    scenario = case_when(
      scenario == "HI_tot" ~ "Air Total",
      scenario == "HI_pt"  ~ "Air Point"
    ),
    source = "Air"
  ) %>%
  select(endpoint, scenario, source, HI)

# water
water_long <- water_df %>%
  mutate(
    scenario = case_when(
      pop == "ad" ~ "Water Adult",
      pop == "ch" ~ "Water Child"
    ),
    source = "Water"
  ) %>%
  select(endpoint, scenario, source, HI)

# combined
combined_long <- combined_df %>%
  pivot_longer(
    cols = c(HI_tot_comb, HI_pt_comb),
    names_to = "comb_type",
    values_to = "HI"
  ) %>%
  mutate(
    scenario = case_when(
      pop == "ad" & comb_type == "HI_pt_comb"  ~ "Combined Adult Point",
      pop == "ad" & comb_type == "HI_tot_comb" ~ "Combined Adult Total",
      pop == "ch" & comb_type == "HI_pt_comb"  ~ "Combined Child Point",
      pop == "ch" & comb_type == "HI_tot_comb" ~ "Combined Child Total"
    ),
    source = "Combined"
  ) %>%
  select(endpoint, scenario, source, HI)

#put tables together
cdf_input <- bind_rows(
  air_long,
  water_long,
  combined_long
)
```


#calculate and graph CDFs
```{r}
#calculate CDFs
cdf_df <- cdf_input %>%
  filter(!is.na(HI), HI > 0) %>%
  group_by(endpoint, scenario) %>%
  arrange(HI, .by_group = TRUE) %>%
  mutate(
    cdf = row_number() / n()
  ) %>%
  ungroup()

#optional thinning, if needed
cdf_sample <- cdf_df %>%
  group_by(endpoint, scenario) %>%
  slice(seq(1, n(), by = 10)) %>%
  ungroup()

#coloring and line types for CDFs
scenario_colors <- c(
  "Air Point" = "#9ecae1",
  "Air Total" = "#08519c",

  "Water Adult" = "#a1d99b",
  "Water Child" = "#238b45",

  "Combined Adult Point" = "#fdd0a2",
  "Combined Adult Total" = "#e6550d",

  "Combined Child Point" = "#fcae91",
  "Combined Child Total" = "#cb181d"
)

scenario_linetypes <- c(
  "Air Point" = "dashed",
  "Air Total" = "solid",

  "Water Adult" = "solid",
  "Water Child" = "dashed",

  "Combined Adult Point" = "dashed",
  "Combined Adult Total" = "solid",

  "Combined Child Point" = "dashed",
  "Combined Child Total" = "solid"
)

#one plot per endpoint
endpoint_list <- sort(unique(cdf_sample$endpoint))

plots_by_endpoint <- list()

for(ep in endpoint_list){

  plot_df <- cdf_sample %>%
    filter(endpoint == ep)

  p <- ggplot(
    plot_df,
    aes(
      x = HI,
      y = cdf,
      color = scenario,
      linetype = scenario
    )
  ) +
    geom_line(linewidth = 1) +
    geom_vline(
      xintercept = c(0.1, 1),
      color = "black",
      linewidth = 1,
      linetype = "dashed"
    ) +
    scale_x_log10() +
    coord_cartesian(
      xlim = c(1e-8, max(plot_df$HI, na.rm = TRUE))
    ) +
    scale_color_manual(values = scenario_colors) +
    scale_linetype_manual(values = scenario_linetypes) +
    labs(
      title = ep,
      x = "Hazard Index",
      y = "Cumulative Probability"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position = "bottom",
      plot.title = element_text(face = "bold")
    )

  plots_by_endpoint[[ep]] <- p

  print(p)
}

#save
output_dir <- "~/TSCA/CRA/metdb_tsca/output/chlor/Combined"

for(ep in names(plots_by_endpoint)){

  ggsave(
    filename = paste0(
      "CDF_",
      gsub("[^A-Za-z0-9]", "_", ep),
      ".png"
    ),
    plot = plots_by_endpoint[[ep]],
    path = output_dir,
    width = 8,
    height = 6,
    dpi = 300
  )
}
```








