---
title: "Chlor Visualization for MS publication - paneled plots"
output: html_document
date: "2026-07-01"
---


# Load libraries
```{r}
library(bigrquery)
library(DBI)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
library(stringr)
```


# Connect to BigQuery
```{r}
bq_auth(
  email = NA
)

con <- dbConnect(
  bigrquery::bigquery(),
  project = "edf-aq-data",
  dataset = "airtoxscreen"
)

project_id <- "edf-aq-data"
```


#important sensitivity analysis on combination statistics: 15,166 blocks are included in the PWS dataset but not the ATS dataset. need to check the population from the Census data on these blocks to see if they weren't included in ATS because ATS doesn't model exposures for blocks with population = 0. It makes sense otherwise that there are lots of blocks that have ATS data and not PWS data because millions are not on PWS.
```{r}
# Step 1: Query unmatched PWS blocks from BigQuery
SQL <- "
SELECT DISTINCT
  LPAD(CAST(pws.geoid20 AS STRING), 15, '0') AS geoid20
FROM `edf-aq-data.TSCA.chlor_PWS_HQs_half` AS pws
WHERE LPAD(CAST(pws.geoid20 AS STRING), 15, '0') NOT IN (
  SELECT DISTINCT LPAD(CAST(block AS STRING), 15, '0')
  FROM `edf-aq-data.TSCA.chlor_ATS_HIs`
)
"
unmatched_PWS <- dbGetQuery(con, SQL)

# Step 2: Get state FIPS codes from unmatched blocks
unmatched_PWS <- unmatched_PWS %>%
  mutate(state_fips = substr(geoid20, 1, 2))

states_needed <- unique(unmatched_PWS$state_fips)

# Step 4: Loop through those states to pull 2020 block-level population
pop_blocks_all <- data.frame()

for (fips in states_needed) {
  message("Getting population for state FIPS: ", fips)
  
  tryCatch({
    state_blocks <- get_decennial(
      geography = "block",
      variables = "P1_001N",  # total population 2020 PL data
      year = 2020,
      state = fips,
      geometry = FALSE,
      output = "wide",
      dataset = "pl"
    )
    
    pop_blocks_all <- bind_rows(pop_blocks_all, state_blocks)
  }, error = function(e) {
    message("Error for state ", fips, ": ", e$message)
  })
}

# Step 5: Join population to unmatched PWS blocks
pws_blocks_with_pop <- unmatched_PWS %>%
  left_join(pop_blocks_all, by = c("geoid20" = "GEOID"))

# Step 6: Summarize
summary_table <- pws_blocks_with_pop %>%
  mutate(pop = P1_001N) %>%
  summarise(
    total_blocks = n(),
    blocks_with_population = sum(!is.na(pop) & pop > 0),
    blocks_with_zero_or_na = sum(is.na(pop) | pop == 0),
    total_population = sum(pop, na.rm = TRUE)
  )

# Step 7: View results
print(summary_table)

# all 15,166 blocks included in PWS and not in ATS have a population of zero or NA in the 2020 Census - this is great news, which means that the joining of the ATS and PWS HIs are joined properly and there are no issues with mismatching geoids, for example
```



# Define output directory
```{r}
output_dir <- "~/TSCA/CRA/metdb_tsca/output/Chlor/"

dir.create(
  output_dir,
  recursive = TRUE,
  showWarnings = FALSE
)
```


# Define endpoint and scenario settings
```{r}

selected_endpoints <- c(
  "developmental",
  "renal"
)

endpoint_levels <- c(
  "Developmental",
  "Renal"
)

four_scenario_levels <- c(
  "Point air",
  "Total air",
  "Total air + child drinking water",
  "Cumulative HI"
)

four_scenario_labels <- c(
  "Point air" =
    "Point air",

  "Total air" =
    "Total air",

  "Total air + child drinking water" =
    "Total air + Ingestion (Child)",

  "Cumulative HI" =
    "Cumulative HI (Child)"
)
```


# Distribution-summary function
```{r}
summarise_distribution <- function(
  data,
  value,
  ...
) {

  data %>%
    group_by(...) %>%
    summarise(
      sample_size = sum(
        !is.na({{ value }})
      ),

      mean = mean(
        {{ value }},
        na.rm = TRUE
      ),

      median = median(
        {{ value }},
        na.rm = TRUE
      ),

      minimum = min(
        {{ value }},
        na.rm = TRUE
      ),

      maximum = max(
        {{ value }},
        na.rm = TRUE
      ),

      sd = sd(
        {{ value }},
        na.rm = TRUE
      ),

      se = sd(
        {{ value }},
        na.rm = TRUE
      ) /
        sqrt(
          sum(
            !is.na({{ value }})
          )
        ),

      .groups = "drop"
    )
}
```


#pull TCE HQ data
```{r}
# ============================================================
# 1. Pull TCE air HQs
#
# GEOIDs are standardized to 15 characters before joining.
# Only the two publication endpoints are downloaded.
# ============================================================

SQL_tce_air_hq <- "
SELECT
  LPAD(
    CAST(h.block AS STRING),
    15,
    '0'
  ) AS geoid20,

  h.pollutant_name,
  LOWER(TRIM(h.endpoint)) AS endpoint,
  h.HQ_pt,
  h.HQ_tot

FROM
  `edf-aq-data.TSCA.chlor_ATS_HQs` h

JOIN (
  SELECT DISTINCT
    GEOID

  FROM
    `edf-aq-data.TSCA.filtblocks_chlor`
) f

  ON LPAD(
    CAST(h.block AS STRING),
    15,
    '0'
  ) = f.GEOID

WHERE
  UPPER(TRIM(h.pollutant_name)) =
    'TRICHLOROETHYLENE'

  AND LOWER(TRIM(h.endpoint)) IN (
    'developmental',
    'renal'
  )
"


tce_air_hq <- bq_project_query(
  project_id,
  SQL_tce_air_hq
) %>%
  bq_table_download() %>%
  transmute(
    geoid20 = as.character(
      geoid20
    ),

    endpoint = str_to_lower(
      str_trim(endpoint)
    ),

    HQ_pt = as.numeric(
      HQ_pt
    ),

    HQ_tot = as.numeric(
      HQ_tot
    )
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    endpoint %in% selected_endpoints
  )


# ============================================================
# 2. Check TCE air HQ uniqueness
#
# Expected unique unit:
# geoid20 + endpoint
# ============================================================

tce_air_hq_duplicate_check <-
  tce_air_hq %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

tce_air_hq_duplicate_check


# ============================================================
# 3. Pull TCE child drinking-water HQs
#
# The PWS table uses the chemical code "tce".
# ============================================================

SQL_tce_water_hq <- "
SELECT
  LPAD(
    CAST(p.geoid20 AS STRING),
    15,
    '0'
  ) AS geoid20,

  LOWER(TRIM(p.endpoint)) AS endpoint,
  p.pop,
  p.chem,
  p.hq AS HQ_water

FROM
  `edf-aq-data.TSCA.chlor_PWS_HQs_half` p

JOIN (
  SELECT DISTINCT
    GEOID

  FROM
    `edf-aq-data.TSCA.filtblocks_chlor`
) f

  ON LPAD(
    CAST(p.geoid20 AS STRING),
    15,
    '0'
  ) = f.GEOID

WHERE
  LOWER(TRIM(p.chem)) = 'tce'

  AND LOWER(TRIM(p.pop)) = 'ch'

  AND LOWER(TRIM(p.endpoint)) IN (
    'developmental',
    'renal'
  )
"


tce_water_hq <- bq_project_query(
  project_id,
  SQL_tce_water_hq
) %>%
  bq_table_download() %>%
  transmute(
    geoid20 = as.character(
      geoid20
    ),

    endpoint = str_to_lower(
      str_trim(endpoint)
    ),

    HQ_water = as.numeric(
      HQ_water
    )
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    endpoint %in% selected_endpoints
  )


# ============================================================
# 4. Check TCE water HQ uniqueness
# ============================================================

tce_water_hq_duplicate_check <-
  tce_water_hq %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

tce_water_hq_duplicate_check


# ============================================================
# 5. Create TCE HQ data for the three HQ scenarios
#
# The inner join ensures that all three HQ scenarios initially
# use block-endpoint combinations with both air and child
# drinking-water HQ information.
# ============================================================

tce_combined_hq <- tce_air_hq %>%
  inner_join(
    tce_water_hq,
    by = c(
      "geoid20",
      "endpoint"
    )
  ) %>%
  mutate(
    HQ_combined =
      HQ_tot + HQ_water
  )


# ============================================================
# 6. Check final combined TCE HQ uniqueness
# ============================================================

tce_combined_hq_duplicate_check <-
  tce_combined_hq %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

tce_combined_hq_duplicate_check


# ============================================================
# 7. Convert TCE HQ data to three-scenario long format
# ============================================================

tce_hq_scenario_df <- tce_combined_hq %>%
  transmute(
    geoid20,
    endpoint,

    `Point air` =
      HQ_pt,

    `Total air` =
      HQ_tot,

    `Total air + child drinking water` =
      HQ_combined
  ) %>%
  pivot_longer(
    cols = c(
      `Point air`,
      `Total air`,
      `Total air + child drinking water`
    ),

    names_to = "scenario",
    values_to = "value"
  ) %>%
  filter(
    !is.na(value)
  ) %>%
  mutate(
    metric = "TCE HQ"
  )


# ============================================================
# 8. Check resulting TCE scenario data
# ============================================================

tce_hq_scenario_df %>%
  count(
    endpoint,
    scenario,
    name = "n_records"
  )


tce_hq_scenario_df %>%
  group_by(
    endpoint,
    scenario
  ) %>%
  summarise(
    total_blocks = n_distinct(
      geoid20
    ),

    .groups = "drop"
  )

```



# Pull cumulative HIs
```{r}
# ============================================================
# Pull cumulative HIs only for Developmental and Renal
#
# Population priority is handled in BigQuery:
# 1. Use child ("ch") HI when available
# 2. Otherwise use population-unspecified ("NA") HI
#
# Only one row per block-endpoint is downloaded.
# ============================================================

SQL_selected_endpoint_hi <- "
WITH eligible_hi AS (

  SELECT
    h.geoid20,

    LOWER(
      TRIM(h.endpoint)
    ) AS endpoint,

    LOWER(
      TRIM(h.pop)
    ) AS pop,

    h.HI_tot_comb,

    ROW_NUMBER() OVER (
      PARTITION BY
        h.geoid20,
        LOWER(TRIM(h.endpoint))

      ORDER BY
        CASE
          WHEN LOWER(TRIM(h.pop)) = 'ch'
            THEN 1

          WHEN LOWER(TRIM(h.pop)) = 'na'
            THEN 2

          ELSE 3
        END
    ) AS population_rank

  FROM
    `edf-aq-data.TSCA.chlor_combined_HIs` h

  JOIN (
    SELECT DISTINCT
      GEOID

    FROM
      `edf-aq-data.TSCA.filtblocks_chlor`
  ) f

    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'na'
    )

    AND LOWER(TRIM(h.endpoint)) IN (
      'developmental',
      'renal'
    )

    AND h.HI_tot_comb IS NOT NULL
)

SELECT
  geoid20,
  endpoint,
  pop AS selected_population,
  HI_tot_comb

FROM
  eligible_hi

WHERE
  population_rank = 1
"


selected_endpoint_hi_raw <- bq_project_query(
  project_id,
  SQL_selected_endpoint_hi
) %>%
  bq_table_download() %>%
  transmute(
    geoid20 = as.character(
      geoid20
    ),

    endpoint = str_to_lower(
      str_trim(endpoint)
    ),

    selected_population = str_to_lower(
      str_trim(selected_population)
    ),

    HI_tot_comb = as.numeric(
      HI_tot_comb
    )
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    endpoint %in% selected_endpoints,
    !is.na(HI_tot_comb)
  )


# ============================================================
# Check that one selected HI exists per block-endpoint
#
# This should return zero rows.
# ============================================================

selected_endpoint_hi_duplicate_check <-
  selected_endpoint_hi_raw %>%
  count(
    geoid20,
    endpoint,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

selected_endpoint_hi_duplicate_check


# ============================================================
# Report selected HI population sources
# ============================================================

selected_endpoint_hi_population_summary <-
  selected_endpoint_hi_raw %>%
  count(
    selected_population,
    name = "selected_block_endpoint_records"
  )

selected_endpoint_hi_population_summary


# ============================================================
# Check selected HI block counts by endpoint
# ============================================================

selected_endpoint_hi_block_summary <-
  selected_endpoint_hi_raw %>%
  group_by(
    endpoint
  ) %>%
  summarise(
    total_records = n(),

    total_blocks = n_distinct(
      geoid20
    ),

    .groups = "drop"
  )

selected_endpoint_hi_block_summary
```



#prepare data for figure
```{r}

# ============================================================
# Prepare cumulative HI records for the final figure
# ============================================================

selected_endpoint_hi_df <- selected_endpoint_hi_raw %>%
  transmute(
    geoid20,
    endpoint,
    scenario = "Cumulative HI",
    metric = "Cumulative HI",
    value = HI_tot_comb
  ) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    !is.na(value)
  )


# ============================================================
# Combine TCE HQ and cumulative HI records
#
# This full dataset retains zeros and is used for tables.
# ============================================================

combined_endpoint_analysis_df <- bind_rows(
  tce_hq_scenario_df,
  selected_endpoint_hi_df
) %>%
  filter(
    !is.na(geoid20),
    !is.na(endpoint),
    !is.na(scenario),
    !is.na(metric),
    !is.na(value)
  ) %>%
  mutate(
    endpoint = str_to_title(
      endpoint
    ),

    endpoint = factor(
      endpoint,
      levels = endpoint_levels
    ),

    scenario = factor(
      scenario,
      levels = four_scenario_levels
    ),

    metric = factor(
      metric,
      levels = c(
        "TCE HQ",
        "Cumulative HI"
      )
    )
  )


# ============================================================
# Restrict to blocks present in both endpoints and all
#     four scenarios
#
# Required combinations:
# 2 endpoints x 4 scenarios = 8 combinations per block.
# ============================================================

required_endpoint_scenario_combinations <-
  length(endpoint_levels) *
  length(four_scenario_levels)


common_analysis_blocks <-
  combined_endpoint_analysis_df %>%
  distinct(
    geoid20,
    endpoint,
    scenario
  ) %>%
  count(
    geoid20,
    name = "n_endpoint_scenario_combinations"
  ) %>%
  filter(
    n_endpoint_scenario_combinations ==
      required_endpoint_scenario_combinations
  ) %>%
  select(
    geoid20
  )


# Number of common blocks retained
nrow(
  common_analysis_blocks
)


combined_endpoint_analysis_df <-
  combined_endpoint_analysis_df %>%
  semi_join(
    common_analysis_blocks,
    by = "geoid20"
  )


# ============================================================
# Final analytical duplicate check
#
# This should return zero rows.
# ============================================================

combined_endpoint_duplicate_check <-
  combined_endpoint_analysis_df %>%
  count(
    geoid20,
    endpoint,
    scenario,
    name = "n_records"
  ) %>%
  filter(
    n_records > 1
  )

combined_endpoint_duplicate_check


# ============================================================
# Confirm identical sample sizes across all eight groups
# ============================================================

combined_endpoint_sample_sizes <-
  combined_endpoint_analysis_df %>%
  group_by(
    endpoint,
    scenario,
    metric
  ) %>%
  summarise(
    total_records = n(),

    total_blocks = n_distinct(
      geoid20
    ),

    .groups = "drop"
  )

combined_endpoint_sample_sizes


stopifnot(
  n_distinct(
    combined_endpoint_sample_sizes$total_blocks
  ) == 1
)

stopifnot(
  all(
    combined_endpoint_sample_sizes$total_records ==
      combined_endpoint_sample_sizes$total_blocks
  )
)


# ============================================================
# Create positive-value-only plot dataset
#
# Zero values remain in combined_endpoint_analysis_df for
# summary tables and denominator calculations.
# ============================================================

combined_endpoint_plot_df <-
  combined_endpoint_analysis_df %>%
  filter(
    value > 0
  )


# ============================================================
# Check zero values omitted from the log-scale figure
# ============================================================

combined_endpoint_zero_summary <-
  combined_endpoint_analysis_df %>%
  group_by(
    endpoint,
    scenario
  ) %>%
  summarise(
    total_blocks = n_distinct(
      geoid20
    ),

    zero_value_blocks = n_distinct(
      geoid20[
        value == 0
      ]
    ),

    positive_value_blocks = n_distinct(
      geoid20[
        value > 0
      ]
    ),

    .groups = "drop"
  )

combined_endpoint_zero_summary


# ============================================================
# Create distribution summary table for figure data
#
# Uses the full common-block dataset, including zero values.
# ============================================================

combined_endpoint_distribution_summary <-
  summarise_distribution(
    data = combined_endpoint_analysis_df,
    value = value,
    endpoint,
    scenario,
    metric
  ) %>%
  mutate(
    endpoint = as.character(
      endpoint
    ),

    scenario = as.character(
      scenario
    ),

    metric = as.character(
      metric
    )
  ) %>%
  arrange(
    factor(
      endpoint,
      levels = endpoint_levels
    ),

    factor(
      scenario,
      levels = four_scenario_levels
    )
  ) %>%
  select(
    endpoint,
    scenario,
    metric,
    sample_size,
    mean,
    median,
    minimum,
    maximum,
    sd,
    se
  )


combined_endpoint_distribution_summary


write.csv(
  combined_endpoint_distribution_summary,
  file = file.path(
    output_dir,
    paste0(
      "chlor_TCE_developmental_renal_",
      "distribution_summary.csv"
    )
  ),
  row.names = FALSE
)
```




#plot figure
```{r}
# ============================================================
# Determine log-scale range
# ============================================================

combined_log_lower_exponent <- floor(
  log10(
    min(
      combined_endpoint_plot_df$value,
      na.rm = TRUE
    )
  )
)

combined_log_upper_exponent <- ceiling(
  log10(
    max(
      combined_endpoint_plot_df$value,
      na.rm = TRUE
    )
  )
)


combined_log_major_breaks <- 10^seq(
  combined_log_lower_exponent,
  combined_log_upper_exponent,
  by = 1
)


# ============================================================
# Define log-axis labels
# ============================================================

combined_log_major_labels <- function(x) {

  ifelse(
    dplyr::near(
      x,
      0.1
    ),
    "0.1",

    ifelse(
      dplyr::near(
        x,
        1
      ),
      "1",

      format(
        x,
        scientific = TRUE,
        trim = TRUE,
        digits = 1
      )
    )
  )
}


# ============================================================
# Create final TCE publication figure
# ============================================================

tce_combined_endpoints_beeswarm_log <- ggplot(
  combined_endpoint_plot_df,
  aes(
    x = scenario,
    y = value,
    color = endpoint
  )
) +

  geom_quasirandom(
    aes(
      group = interaction(
        scenario,
        endpoint
      )
    ),
    dodge.width = 0.65,
    width = 0.18,
    alpha = 0.35,
    size = 0.4
  ) +

  geom_hline(
    yintercept = c(
      0.1,
      1
    ),
    linetype = "dotted",
    linewidth = 0.8,
    color = "grey40"
  ) +

  scale_y_log10(
    breaks = combined_log_major_breaks,
    minor_breaks = NULL,
    labels = combined_log_major_labels,
    expand = expansion(
      mult = c(
        0.03,
        0.05
      )
    )
  ) +

  scale_color_manual(
    values = c(
      "Developmental" = "#0072B2",
      "Renal" = "#D55E00"
    ),
    labels = c(
      "Developmental" = "Developmental (most sensitive)",
      "Renal" = "Renal"
    ),
    guide = guide_legend(
      override.aes = list(
        alpha = 1,
        size = 3
      )
    )
  ) +

  scale_x_discrete(
    labels = four_scenario_labels
  ) +

  labs(
    x = NULL,
    y = "HQ or HI (log scale)",
    color = "Endpoint"
  ) +

  theme_bw() +

  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    ),

    panel.grid.major.x =
      element_blank(),

    panel.grid.minor =
      element_blank(),

    panel.grid.major.y =
      element_line(
        color = "grey85",
        linewidth = 0.4
      ),

    legend.position = "bottom"
  )


tce_combined_endpoints_beeswarm_log


# ============================================================
# Save final figure
# ============================================================

ggsave(
  filename = paste0(
    "chlor_TCE_developmental_renal_",
    "aggregate_cumulative_log.png"
  ),
  plot = tce_combined_endpoints_beeswarm_log,
  path = output_dir,
  width = 8,
  height = 5.5,
  dpi = 600
)
```



#create table for block threshold numbers/percentages that goes with the figure
```{r}

# ============================================================
# Create raw eight-row threshold table
#
# Thresholds:
# value >= 0.1
# value >= 1
# ============================================================

combined_endpoint_exceedance_table_raw <-
  combined_endpoint_analysis_df %>%
  group_by(
    endpoint,
    scenario,
    metric
  ) %>%
  summarise(
    total_blocks = n_distinct(
      geoid20
    ),

    blocks_ge_0_1 = n_distinct(
      geoid20[
        value >= 0.1
      ]
    ),

    percent_ge_0_1 =
      100 *
      blocks_ge_0_1 /
      total_blocks,

    blocks_ge_1 = n_distinct(
      geoid20[
        value >= 1
      ]
    ),

    percent_ge_1 =
      100 *
      blocks_ge_1 /
      total_blocks,

    .groups = "drop"
  ) %>%
  mutate(
    endpoint = as.character(
      endpoint
    ),

    scenario = as.character(
      scenario
    ),

    metric = as.character(
      metric
    )
  ) %>%
  arrange(
    factor(
      endpoint,
      levels = endpoint_levels
    ),

    factor(
      scenario,
      levels = four_scenario_levels
    )
  )


combined_endpoint_exceedance_table_raw


stopifnot(
  nrow(
    combined_endpoint_exceedance_table_raw
  ) == 8
)


combined_endpoint_exceedance_validation <-
  combined_endpoint_exceedance_table_raw %>%
  summarise(
    any_ge_0_1_above_total =
      any(
        blocks_ge_0_1 >
          total_blocks
      ),

    any_ge_1_above_ge_0_1 =
      any(
        blocks_ge_1 >
          blocks_ge_0_1
      ),

    any_percent_ge_0_1_above_100 =
      any(
        percent_ge_0_1 >
          100
      ),

    any_percent_ge_1_above_100 =
      any(
        percent_ge_1 >
          100
      )
  )

combined_endpoint_exceedance_validation


# ============================================================
# Format eight-row threshold table
# ============================================================

combined_endpoint_exceedance_table <-
  combined_endpoint_exceedance_table_raw %>%
  mutate(
    total_blocks_formatted = format(
      total_blocks,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),

    blocks_ge_0_1_n_percent = paste0(
      format(
        blocks_ge_0_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_0_1,
        format = "f",
        digits = 2
      ),
      "%)"
    ),

    blocks_ge_1_n_percent = paste0(
      format(
        blocks_ge_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_1,
        format = "f",
        digits = 2
      ),
      "%)"
    )
  ) %>%
  transmute(
    Endpoint = endpoint,
    Scenario = scenario,
    Metric = metric,

    `Total blocks` =
      total_blocks_formatted,

    `Blocks >= 0.1, n (%)` =
      blocks_ge_0_1_n_percent,

    `Blocks >= 1, n (%)` =
      blocks_ge_1_n_percent
  )


combined_endpoint_exceedance_table


write.csv(
  combined_endpoint_exceedance_table,
  file = file.path(
    output_dir,
    paste0(
      "chlor_TCE_developmental_renal_",
      "block_threshold_table.csv"
    )
  ),
  row.names = FALSE
)


write.csv(
  combined_endpoint_exceedance_table_raw,
  file = file.path(
    output_dir,
    paste0(
      "chlor_TCE_developmental_renal_",
      "block_threshold_table_raw.csv"
    )
  ),
  row.names = FALSE
)

```



#for supplemental: block exceedance numbers/percentages for HIs for each endpoint
```{r}

# ============================================================
# Calculate all-endpoint threshold results in BigQuery
#
# Population priority:
# 1. Child HI
# 2. Air-only/NA HI if child is unavailable
# ============================================================

SQL_all_endpoint_hi_thresholds <- "
WITH eligible_hi AS (

  SELECT
    h.geoid20,

    LOWER(
      TRIM(h.endpoint)
    ) AS endpoint,

    LOWER(
      TRIM(h.pop)
    ) AS pop,

    h.HI_tot_comb,

    ROW_NUMBER() OVER (
      PARTITION BY
        h.geoid20,
        LOWER(TRIM(h.endpoint))

      ORDER BY
        CASE
          WHEN LOWER(TRIM(h.pop)) = 'ch'
            THEN 1

          WHEN LOWER(TRIM(h.pop)) = 'na'
            THEN 2

          ELSE 3
        END
    ) AS population_rank

  FROM
    `edf-aq-data.TSCA.chlor_combined_HIs` h

  JOIN (
    SELECT DISTINCT
      GEOID

    FROM
      `edf-aq-data.TSCA.filtblocks_chlor`
  ) f

    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'na'
    )

    AND h.endpoint IS NOT NULL
    AND h.HI_tot_comb IS NOT NULL
),

selected_hi AS (

  SELECT
    geoid20,
    endpoint,
    HI_tot_comb

  FROM
    eligible_hi

  WHERE
    population_rank = 1
)

SELECT
  endpoint,

  COUNT(
    DISTINCT geoid20
  ) AS total_blocks,

  COUNT(
    DISTINCT IF(
      HI_tot_comb >= 0.1,
      geoid20,
      NULL
    )
  ) AS blocks_ge_0_1,

  SAFE_MULTIPLY(
    SAFE_DIVIDE(
      COUNT(
        DISTINCT IF(
          HI_tot_comb >= 0.1,
          geoid20,
          NULL
        )
      ),

      COUNT(
        DISTINCT geoid20
      )
    ),

    100
  ) AS percent_ge_0_1,

  COUNT(
    DISTINCT IF(
      HI_tot_comb >= 1,
      geoid20,
      NULL
    )
  ) AS blocks_ge_1,

  SAFE_MULTIPLY(
    SAFE_DIVIDE(
      COUNT(
        DISTINCT IF(
          HI_tot_comb >= 1,
          geoid20,
          NULL
        )
      ),

      COUNT(
        DISTINCT geoid20
      )
    ),

    100
  ) AS percent_ge_1

FROM
  selected_hi

GROUP BY
  endpoint

ORDER BY
  endpoint
"


cumulative_hi_threshold_table_raw <- bq_project_query(
  project_id,
  SQL_all_endpoint_hi_thresholds
) %>%
  bq_table_download() %>%
  transmute(
    endpoint = str_to_title(
      endpoint
    ),

    total_blocks = as.numeric(
      total_blocks
    ),

    blocks_ge_0_1 = as.numeric(
      blocks_ge_0_1
    ),

    percent_ge_0_1 = as.numeric(
      percent_ge_0_1
    ),

    blocks_ge_1 = as.numeric(
      blocks_ge_1
    ),

    percent_ge_1 = as.numeric(
      percent_ge_1
    )
  ) %>%
  arrange(
    endpoint
  )


cumulative_hi_threshold_table_raw


cumulative_hi_threshold_validation <-
  cumulative_hi_threshold_table_raw %>%
  summarise(
    any_ge_0_1_above_total =
      any(
        blocks_ge_0_1 >
          total_blocks
      ),

    any_ge_1_above_ge_0_1 =
      any(
        blocks_ge_1 >
          blocks_ge_0_1
      ),

    any_percent_ge_0_1_above_100 =
      any(
        percent_ge_0_1 >
          100
      ),

    any_percent_ge_1_above_100 =
      any(
        percent_ge_1 >
          100
      )
  )

cumulative_hi_threshold_validation


# ============================================================
# Format all-endpoint threshold table
# ============================================================

cumulative_hi_threshold_table <-
  cumulative_hi_threshold_table_raw %>%
  mutate(
    total_blocks_formatted = format(
      total_blocks,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),

    blocks_ge_0_1_n_percent = paste0(
      format(
        blocks_ge_0_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_0_1,
        format = "f",
        digits = 2
      ),
      "%)"
    ),

    blocks_ge_1_n_percent = paste0(
      format(
        blocks_ge_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),
      " (",
      formatC(
        percent_ge_1,
        format = "f",
        digits = 2
      ),
      "%)"
    )
  ) %>%
  transmute(
    Endpoint = endpoint,

    `Total blocks` =
      total_blocks_formatted,

    `Blocks >= 0.1, n (%)` =
      blocks_ge_0_1_n_percent,

    `Blocks >= 1, n (%)` =
      blocks_ge_1_n_percent
  )


cumulative_hi_threshold_table


write.csv(
  cumulative_hi_threshold_table,
  file = file.path(
    output_dir,
    "chlor_endpoint_cumulative_HI_threshold_table.csv"
  ),
  row.names = FALSE
)


write.csv(
  cumulative_hi_threshold_table_raw,
  file = file.path(
    output_dir,
    "chlor_endpoint_cumulative_HI_threshold_table_raw.csv"
  ),
  row.names = FALSE
)

```



#for supplemental: table of summary information (n, min, median, mean, max, sd, se) for each chemical HQ for each HI for each endpoint and population (ch, ad, NA)
```{r}

# ============================================================
# Calculate combined-HI summaries by endpoint and type
#
# Types:
# - Child (air + drinking water)
# - Adult (air + drinking water)
# - Air only
#
# These categories are summarized independently.
# No child-over-NA prioritization is applied here because the
# purpose is to report each cumulative HI type separately.
# ============================================================

SQL_combined_hi_type_summary <- "
WITH filtered_blocks AS (

  SELECT DISTINCT
    GEOID

  FROM
    `edf-aq-data.TSCA.filtblocks_chlor`
),

combined_hi AS (

  SELECT
    h.geoid20,

    LOWER(
      TRIM(h.endpoint)
    ) AS endpoint,

    LOWER(
      TRIM(h.pop)
    ) AS population_code,

    h.HI_tot_comb AS value

  FROM
    `edf-aq-data.TSCA.chlor_combined_HIs` h

  INNER JOIN
    filtered_blocks f

    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'ad',
      'na'
    )

    AND h.endpoint IS NOT NULL
    AND h.HI_tot_comb IS NOT NULL
),

aggregate_stats AS (

  SELECT
    endpoint,
    population_code,

    COUNT(value) AS n,

    MIN(value) AS minimum,

    AVG(value) AS mean,

    MAX(value) AS maximum,

    STDDEV_SAMP(value) AS sd

  FROM
    combined_hi

  GROUP BY
    endpoint,
    population_code
),

median_stats AS (

  SELECT DISTINCT
    endpoint,
    population_code,

    PERCENTILE_CONT(
      value,
      0.5
    ) OVER (
      PARTITION BY
        endpoint,
        population_code
    ) AS median

  FROM
    combined_hi
)

SELECT
  a.endpoint,
  a.population_code,
  a.n,
  a.minimum,
  m.median,
  a.mean,
  a.maximum,
  a.sd,

  SAFE_DIVIDE(
    a.sd,
    SQRT(a.n)
  ) AS se

FROM
  aggregate_stats a

LEFT JOIN
  median_stats m

USING (
  endpoint,
  population_code
)

ORDER BY
  endpoint,

  CASE population_code
    WHEN 'ch' THEN 1
    WHEN 'ad' THEN 2
    WHEN 'na' THEN 3
    ELSE 4
  END
"


combined_hi_type_summary_table <-
  bq_project_query(
    project_id,
    SQL_combined_hi_type_summary
  ) %>%
  bq_table_download() %>%
  transmute(
    Endpoint = str_to_title(
      endpoint
    ),

    `Cumulative HI type` = case_when(
      population_code == "ch" ~
        "Child (air + drinking water)",

      population_code == "ad" ~
        "Adult (air + drinking water)",

      population_code == "na" ~
        "Air only",

      TRUE ~
        population_code
    ),

    n = as.numeric(
      n
    ),

    Minimum = as.numeric(
      minimum
    ),

    Median = as.numeric(
      median
    ),

    Mean = as.numeric(
      mean
    ),

    Maximum = as.numeric(
      maximum
    ),

    SD = as.numeric(
      sd
    ),

    SE = as.numeric(
      se
    )
  ) %>%
  mutate(
    `Cumulative HI type` = factor(
      `Cumulative HI type`,
      levels = c(
        "Child (air + drinking water)",
        "Adult (air + drinking water)",
        "Air only"
      )
    )
  ) %>%
  arrange(
    Endpoint,
    `Cumulative HI type`
  ) %>%
  mutate(
    `Cumulative HI type` = as.character(
      `Cumulative HI type`
    )
  )


combined_hi_type_summary_table


# ============================================================
# Validate combined-HI type summary
#
# All values should return FALSE.
# ============================================================

combined_hi_type_summary_validation <-
  combined_hi_type_summary_table %>%
  summarise(
    any_n_below_1 =
      any(
        n < 1
      ),

    any_minimum_above_median =
      any(
        Minimum > Median
      ),

    any_median_above_maximum =
      any(
        Median > Maximum
      ),

    any_mean_below_minimum =
      any(
        Mean < Minimum
      ),

    any_mean_above_maximum =
      any(
        Mean > Maximum
      ),

    any_negative_sd =
      any(
        SD < 0,
        na.rm = TRUE
      ),

    any_negative_se =
      any(
        SE < 0,
        na.rm = TRUE
      )
  )


combined_hi_type_summary_validation


write.csv(
  combined_hi_type_summary_table,
  file = file.path(
    output_dir,
    "chlor_combined_HI_summary_by_endpoint_and_type.csv"
  ),
  row.names = FALSE
)
```



#table for supplemental on how many census blocks have HI of 0.1 or 1 exceedances for more than one endpoint
```{r}

# ============================================================
# 1. Query multi-endpoint exceedance patterns
# ============================================================

SQL_multi_endpoint_exceedance_chlor <- "

WITH filtered_blocks AS (

  SELECT DISTINCT
    GEOID

  FROM
    `edf-aq-data.TSCA.filtblocks_chlor`
),


-- ------------------------------------------------------------
-- Select one cumulative HI per block-endpoint
--
-- Population priority:
-- 1. child
-- 2. NA / air-only
-- ------------------------------------------------------------

eligible_hi AS (

  SELECT
    h.geoid20,

    LOWER(
      TRIM(h.endpoint)
    ) AS endpoint,

    LOWER(
      TRIM(h.pop)
    ) AS pop,

    h.HI_tot_comb,

    ROW_NUMBER() OVER (

      PARTITION BY
        h.geoid20,
        LOWER(TRIM(h.endpoint))

      ORDER BY
        CASE

          WHEN LOWER(TRIM(h.pop)) = 'ch'
            THEN 1

          WHEN LOWER(TRIM(h.pop)) = 'na'
            THEN 2

          ELSE 3

        END

    ) AS population_rank

  FROM
    `edf-aq-data.TSCA.chlor_combined_HIs` h

  INNER JOIN
    filtered_blocks f

    ON h.geoid20 = f.GEOID

  WHERE
    LOWER(TRIM(h.pop)) IN (
      'ch',
      'na'
    )

    AND h.endpoint IS NOT NULL

    AND h.HI_tot_comb IS NOT NULL
),


-- ------------------------------------------------------------
-- Keep selected population record
-- ------------------------------------------------------------

selected_hi AS (

  SELECT
    geoid20,
    endpoint,
    pop AS selected_population,
    HI_tot_comb

  FROM
    eligible_hi

  WHERE
    population_rank = 1
),


-- ------------------------------------------------------------
-- Define HI thresholds
--
-- threshold_id is INT64 and is used for grouping/windowing.
-- threshold_value is the actual HI threshold.
--
-- Using threshold_id avoids BigQuery FLOAT64 PARTITION BY
-- errors.
-- ------------------------------------------------------------

thresholds AS (

  SELECT
    1 AS threshold_id,
    0.1 AS threshold_value

  UNION ALL

  SELECT
    2 AS threshold_id,
    1.0 AS threshold_value
),


-- ------------------------------------------------------------
-- Identify every block-endpoint exceeding each threshold
-- ------------------------------------------------------------

exceeded_endpoints AS (

  SELECT
    h.geoid20,
    t.threshold_id,
    t.threshold_value,
    h.endpoint

  FROM
    selected_hi h

  CROSS JOIN
    thresholds t

  WHERE
    h.HI_tot_comb >= t.threshold_value
),


-- ------------------------------------------------------------
-- Determine total number and full combination of endpoints
-- exceeded within each block
-- ------------------------------------------------------------

block_endpoint_patterns AS (

  SELECT
    geoid20,
    threshold_id,
    threshold_value,

    COUNT(
      DISTINCT endpoint
    ) AS n_endpoints_exceeded,

    STRING_AGG(
      DISTINCT endpoint,
      ', '
      ORDER BY endpoint
    ) AS endpoint_combination

  FROM
    exceeded_endpoints

  GROUP BY
    geoid20,
    threshold_id,
    threshold_value
),


-- ------------------------------------------------------------
-- Expand each block into at-least-N categories
--
-- Example:
-- If n_endpoints_exceeded = 4,
-- create rows for:
-- at_least_n = 1
-- at_least_n = 2
-- at_least_n = 3
-- at_least_n = 4
-- ------------------------------------------------------------

at_least_endpoint_patterns AS (

  SELECT
    geoid20,
    threshold_id,
    threshold_value,
    n_endpoints_exceeded,
    endpoint_combination,
    at_least_n

  FROM
    block_endpoint_patterns,

  UNNEST(
    GENERATE_ARRAY(
      1,
      n_endpoints_exceeded
    )
  ) AS at_least_n
),


-- ------------------------------------------------------------
-- Count total blocks exceeding AT LEAST N endpoints
-- ------------------------------------------------------------

endpoint_number_totals AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,

    COUNT(
      DISTINCT geoid20
    ) AS n_blocks

  FROM
    at_least_endpoint_patterns

  GROUP BY
    threshold_id,
    threshold_value,
    at_least_n
),


-- ------------------------------------------------------------
-- Count full endpoint combinations among blocks in each
-- at-least-N category
-- ------------------------------------------------------------

combination_counts AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,
    endpoint_combination,

    COUNT(
      DISTINCT geoid20
    ) AS n_blocks_combination

  FROM
    at_least_endpoint_patterns

  GROUP BY
    threshold_id,
    threshold_value,
    at_least_n,
    endpoint_combination
),


-- ------------------------------------------------------------
-- Rank endpoint combinations within each threshold and
-- at-least-N category
-- ------------------------------------------------------------

ranked_combinations AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,
    endpoint_combination,
    n_blocks_combination,

    DENSE_RANK() OVER (

      PARTITION BY
        threshold_id,
        at_least_n

      ORDER BY
        n_blocks_combination DESC

    ) AS combination_rank

  FROM
    combination_counts
),


-- ------------------------------------------------------------
-- Retain the most common endpoint combination
--
-- If multiple combinations are tied for most common,
-- all tied combinations are retained and separated by |
-- ------------------------------------------------------------

most_common_combinations AS (

  SELECT
    threshold_id,
    threshold_value,
    at_least_n,

    STRING_AGG(
      endpoint_combination,
      ' | '
      ORDER BY endpoint_combination
    ) AS most_common_endpoints,

    MAX(
      n_blocks_combination
    ) AS n_blocks_most_common

  FROM
    ranked_combinations

  WHERE
    combination_rank = 1

  GROUP BY
    threshold_id,
    threshold_value,
    at_least_n
),


-- ------------------------------------------------------------
-- Total number of distinct Census blocks evaluated
--
-- Includes all blocks represented in selected_hi,
-- regardless of whether they exceed either threshold.
-- ------------------------------------------------------------

total_blocks AS (

  SELECT

    COUNT(
      DISTINCT geoid20
    ) AS total_blocks

  FROM
    selected_hi
)


-- ------------------------------------------------------------
-- Final query output
-- ------------------------------------------------------------

SELECT
  t.threshold_value AS threshold,

  t.at_least_n AS n_endpoints_exceeded,

  b.total_blocks,

  t.n_blocks,

  m.most_common_endpoints,

  m.n_blocks_most_common,

  SAFE_MULTIPLY(
    SAFE_DIVIDE(
      m.n_blocks_most_common,
      t.n_blocks
    ),
    100
  ) AS percent_blocks_most_common

FROM
  endpoint_number_totals t

LEFT JOIN
  most_common_combinations m

  ON
    t.threshold_id =
      m.threshold_id

    AND t.at_least_n =
      m.at_least_n

CROSS JOIN
  total_blocks b

ORDER BY
  t.threshold_id,
  n_endpoints_exceeded
"

# ============================================================
# 2. Run query and download raw long-format results
# ============================================================

multi_endpoint_exceedance_chlor_raw <-
  bq_project_query(
    project_id,
    SQL_multi_endpoint_exceedance_chlor
  ) %>%
  bq_table_download() %>%
  transmute(

    threshold =
      as.numeric(
        threshold
      ),

    n_endpoints_exceeded =
      as.numeric(
        n_endpoints_exceeded
      ),

    total_blocks =
      as.numeric(
        total_blocks
      ),

    n_blocks =
      as.numeric(
        n_blocks
      ),

    most_common_endpoints =
      as.character(
        most_common_endpoints
      ),

    n_blocks_most_common =
      as.numeric(
        n_blocks_most_common
      ),

    percent_blocks_most_common =
      as.numeric(
        percent_blocks_most_common
      )
  ) %>%
  arrange(
    threshold,
    n_endpoints_exceeded
  )


multi_endpoint_exceedance_chlor_raw

# ============================================================
# 3. Convert threshold results to wide format
# ============================================================

multi_endpoint_exceedance_chlor_wide_raw <-
  multi_endpoint_exceedance_chlor_raw %>%

  mutate(

    threshold_label =
      case_when(

        dplyr::near(
          threshold,
          0.1
        ) ~ "0_1",

        dplyr::near(
          threshold,
          1
        ) ~ "1",

        TRUE ~
          as.character(
            threshold
          )
      )
  ) %>%

  select(
    n_endpoints_exceeded,
    total_blocks,
    threshold_label,
    n_blocks,
    most_common_endpoints,
    n_blocks_most_common,
    percent_blocks_most_common
  ) %>%

  pivot_wider(

    names_from =
      threshold_label,

    values_from = c(
      n_blocks,
      most_common_endpoints,
      n_blocks_most_common,
      percent_blocks_most_common
    ),

    names_glue =
      "{.value}_{threshold_label}"
  ) %>%

  arrange(
    n_endpoints_exceeded
  )


multi_endpoint_exceedance_chlor_wide_raw

# ============================================================
# 4. Fill missing threshold combinations
# ============================================================

multi_endpoint_exceedance_chlor_wide_raw <-
  multi_endpoint_exceedance_chlor_wide_raw %>%

  mutate(

    n_blocks_0_1 =
      replace_na(
        n_blocks_0_1,
        0
      ),

    n_blocks_1 =
      replace_na(
        n_blocks_1,
        0
      ),


    most_common_endpoints_0_1 =
      replace_na(
        most_common_endpoints_0_1,
        "None"
      ),

    most_common_endpoints_1 =
      replace_na(
        most_common_endpoints_1,
        "None"
      ),


    n_blocks_most_common_0_1 =
      replace_na(
        n_blocks_most_common_0_1,
        0
      ),

    n_blocks_most_common_1 =
      replace_na(
        n_blocks_most_common_1,
        0
      ),


    percent_blocks_most_common_0_1 =
      replace_na(
        percent_blocks_most_common_0_1,
        0
      ),

    percent_blocks_most_common_1 =
      replace_na(
        percent_blocks_most_common_1,
        0
      )
  )


multi_endpoint_exceedance_chlor_wide_raw

# ============================================================
# 5. Create formatted supplemental table
# ============================================================

multi_endpoint_exceedance_chlor <-
  multi_endpoint_exceedance_chlor_wide_raw %>%

  transmute(

    `Chemical group` =
      "Chlorinated solvents",


    `Total number of Census blocks` =
      format(
        total_blocks,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),


    `Number of endpoints exceeded (at least)` =
      n_endpoints_exceeded,


    `Number of Census blocks with HI >= 0.1` =
      format(
        n_blocks_0_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),


    `Most common endpoints with HI >= 0.1` =
      case_when(

        most_common_endpoints_0_1 == "None" ~
          "None",

        TRUE ~
          str_to_title(
            most_common_endpoints_0_1
          )
      ),


    `Blocks with most common HI >= 0.1 combination, n (%)` =
      case_when(

        n_blocks_0_1 == 0 ~
          "0 (0.00%)",

        TRUE ~
          paste0(

            format(
              n_blocks_most_common_0_1,
              big.mark = ",",
              scientific = FALSE,
              trim = TRUE
            ),

            " (",

            formatC(
              percent_blocks_most_common_0_1,
              format = "f",
              digits = 2
            ),

            "%)"
          )
      ),


    `Number of Census blocks with HI >= 1` =
      format(
        n_blocks_1,
        big.mark = ",",
        scientific = FALSE,
        trim = TRUE
      ),


    `Most common endpoints with HI >= 1` =
      case_when(

        most_common_endpoints_1 == "None" ~
          "None",

        TRUE ~
          str_to_title(
            most_common_endpoints_1
          )
      ),


    `Blocks with most common HI >= 1 combination, n (%)` =
      case_when(

        n_blocks_1 == 0 ~
          "0 (0.00%)",

        TRUE ~
          paste0(

            format(
              n_blocks_most_common_1,
              big.mark = ",",
              scientific = FALSE,
              trim = TRUE
            ),

            " (",

            formatC(
              percent_blocks_most_common_1,
              format = "f",
              digits = 2
            ),

            "%)"
          )
      )
  )


multi_endpoint_exceedance_chlor

# ============================================================
# 6. Validate results
#
# All checks below should return FALSE.
# ============================================================

multi_endpoint_exceedance_chlor_validation <-
  multi_endpoint_exceedance_chlor_wide_raw %>%

  arrange(
    n_endpoints_exceeded
  ) %>%

  summarise(

    any_0_1_above_total =
      any(
        n_blocks_0_1 >
          total_blocks
      ),

    any_1_above_total =
      any(
        n_blocks_1 >
          total_blocks
      ),


    any_0_1_count_increases =
      any(
        diff(
          n_blocks_0_1
        ) > 0
      ),

    any_1_count_increases =
      any(
        diff(
          n_blocks_1
        ) > 0
      ),


    any_common_0_1_above_group_total =
      any(
        n_blocks_most_common_0_1 >
          n_blocks_0_1
      ),

    any_common_1_above_group_total =
      any(
        n_blocks_most_common_1 >
          n_blocks_1
      ),


    any_common_percent_0_1_above_100 =
      any(
        percent_blocks_most_common_0_1 >
          100
      ),

    any_common_percent_1_above_100 =
      any(
        percent_blocks_most_common_1 >
          100
      )
  )


multi_endpoint_exceedance_chlor_validation

# ============================================================
# 7. Save formatted supplemental table
# ============================================================

write.csv(
  multi_endpoint_exceedance_chlor,
  file = file.path(
    output_dir,
    "chlor_multi_endpoint_HI_exceedance_table_wide.csv"
  ),
  row.names = FALSE
)


# ============================================================
# 8. Save raw numeric version
# ============================================================

write.csv(
  multi_endpoint_exceedance_chlor_wide_raw,
  file = file.path(
    output_dir,
    "chlor_multi_endpoint_HI_exceedance_table_wide_raw.csv"
  ),
  row.names = FALSE
)
```







